### 2022-10-06 paper.tex
Initial Overleaf Import

\documentclass{juliacon}
\setcounter{page}{1}

\begin{document}

\input{header}

\maketitle

\begin{abstract}

2 - 3 sentences

\end{abstract}

\section{Introduction}

Paragraph 1: Background on observational health research

Observational health data captures a comprehensive view of patient cohort populations, and they are often used to assess incidence rates, prevalence rates, risk factors, and prognosis associated with physical and mental health conditions and treatment interventions. By analyzing healthcare records and clinical events, observational health research can perform cohort characterization, population-estimation, and patient-level predictions to evaluate population health, target interventions and prevention measures, and ultimately transform healthcare policy.

However, current observational health and epidemiological surveys are limited by their data collection and administration methods, sample size representativeness, cohort phenotype definitions, and statistical/algorithmic methods to combine outside, disparate data sources. There is a need for timely, unbiased, accurate, and quality data regarding the mental illness and the ability detect and characterize trends in prevalence, severity, and the association with other chronic illnesses/morbidities.

This effort could help leading health officials implement better informed interventions that will increase equitable access to mental healthcare across vulnerable populations, support efforts to decrease acute mental health-related morbidities, and improve efficacy of health spending.

Paragraph 2: What is OHDSI + HADES R ecosystem

Observational Health Data Sciences and Informatics (OHDSI) is a global, collaborative, interdisciplinary, and open-science network of researchers, healthcare professionals, students, patients, other individuals and organizations committed to large-scale health data analytics and generating reliable evidence, with the goal of promoting better health decisions and better care.

Since current observational health data sources are aggregated at different levels and cannot be generalized to the overall population, the data is often incomplete, inaccurate, and prone to issues of bias, error handling, and multidimensional complexity. The goal of OHDSI is to produce reproducible, robust, and reliable evidence by developing standardized data formats and research protocol designs to enable large-scale collaborative research and improve the transparency and scalability of observational health research.

The development of the Observational Medical Outcomes Partnership (OMOP) was created in 2009 to address concerns related to data type, study design, and data privacy concerns. The OMOP Common Data Model (CDM) is a "person-centered" information model with standardized relationships and structures with other event tables to create a longitudinal view of all clinical, healthcare-related events for each person.

OHDSI also develops open-source analytical and software tools, such as HADES and ATLAS, which features a comprehensive set of functions needed to conduct network studies from assessing demographic information to generating summary statistics and interactive dashboards. The best practices set forth by OHDSI has improved the quality and reliability of evidence in observational studies, and it improved network research study collaboration, supported large-scale analytics, and standardized data structures and medical vocabulary.

Health Analytics Data-to-Evidence Suite (HADES) is a collection of 20 R packages, developed by the OHDSI team, that interact directly with the OMOP CDM to support large-scale analytics, including cohort construction (CirceR, CohortGenerator), population-level estimation (CohortMethod), patient-level prediction (PatientLevelPrediction), and return statistics, figures, and tables.

Paragraph 3: Why tooling was needed

\section{Methods}

\subsection{Tooling for Observational Health Research}

Paragraph 1: Introduction of novel packages

Paragraph 2: Development philosophy and algorithmic design

\subsection{Example Use Cases}

\subsubsection{Minimal Characterization Study}

Paragraph 1: Condensed version of tutorial: https://juliahealth.org/OMOPCDMCohortCreator.jl/dev/tutorials/

Paragraph 2: Example of using results (base off this paragraph): From here, the potential to calculate valuable metrics disease prevalence across niche populations is rich and opens up the possibility to do even more with these subpopulations one could generate (e.g. cross walking with survey panel data, area estimation, etc.). Furthermore, I have tested this but have not had an opportunity to write a tutorial on the matter, but these functionalities work with Distributed.jl to enable non-blocking asynchronous code execution (still working out exact approach/functionalities) to fully utilize Julia’s performance when working and transforming large amounts of data that is often found in observational studies.

\subsubsection{Composability within Ecosystem}

Paragraph 1 - 2: Distributed computing + Distributed.jl (maybe with Memoization) + Light benchmarking (?)

Paragraph 3: Support for other databases + FunSQL.jl + SQL output

\section{Discussion}

Paragraph 1: How it is being used

Paragraph 2: Next steps and Future - base off of this: As I was recently funded another year to continue my health disparities research (yay!), I have great ambitions to build out observational health research capacity in JuliaHealth through avenues such as:

Finalize a submission to JuliaCon Proceedings (need to finish writing tests across this portfolio and adding documentation in the coming weeks)
Partnering on potentially running studies within the Julia community using these tools
Scoping out better support for survey and panel data
Building additional tooling around observational health research (such as determining fairness metrics, easily deployable interactive results explorers, interop with OHDSI tools, etc.)
Lowering barrier to entry of prototyping large scale studies for researchers

\section{Conclusion}

Paragraph 1: Final thoughts

\section{Acknowledgements}

\section{References}

\input{bib.tex}

\end{document}

### 2023-09-30 paper.tex

@@
 
-2 - 3 sentences
+Observational health continues to be a growing field in health informatics research as electronic health records, patient medical claims, and other ancilliary patient data source become more readily computable and accessible to researchers.
+JuliaHealth is poised as an ecosystem to innovate within this area of research by bringing highly performant analytics approaches, composable solutions, and interoperable software that leverages prior state of the art.
+This paper will discuss the state of the art observational health research tools within the JuliaHealth ecosystem and how JuliaHealth is prepared to further research goals within this domain.
 
@@
 
-Paragraph 1: Background on observational health research
+Over the past $10$ years, there have been several innovations within the observational health research.
+Arguably, the most crucial innovation that has emerged is the Observational Medical Outcomes Partnership Common Data Model (OMOP CDM) which was designed specifically to standardize observational health data to enable transferable analyses.
+The importance of this model has become apparent within open science organizations like the OHDSI (Observational Health Data Science and Informatics) network where members across $100+$ countries with access to various health systems have been able to successfully map their patient data to the OMOP CDM.
+This has led to successful collaboration relationships across countries where one research group can develop a network study and then disseminate the study across collaborators leading to the idea that you can "write once, run everywhere."
 
-Observational health data captures a comprehensive view of patient cohort populations, and they are often used to assess incidence rates, prevalence rates, risk factors, and prognosis associated with physical and mental health conditions and treatment interventions.
-By analyzing healthcare records and clinical events, observational health research can perform cohort characterization, population-estimation, and patient-level predictions to evaluate population health, target interventions and prevention measures, and ultimately transform healthcare policy.
+The OMOP CDM itself does not require a specific technology to work with the data stored in this standard.
+It features a person-centric design where each domain records personal identity while prioritizing data protection through the limiting of information that could endanger patient anonymity.
+As the OMOP CDM has become a cornerstone in running these widely distributed network studies, special attention to developing tools around handling OMOP CDM data is necessary to meaningfully contribute within the observational health research space.
 
-However, current observational health and epidemiological surveys are limited by their data collection and administration methods, sample size representativeness, cohort phenotype definitions, and statistical/algorithmic methods to combine outside, disparate data sources.
-There is a need for timely, unbiased, accurate, and quality data regarding the mental illness and the ability detect and characterize trends in prevalence, severity, and the association with other chronic illnesses/morbidities.
+\subsection{What Is JuliaHealth and Its Goals?}
 
-This effort could help leading health officials implement better informed interventions that will increase equitable access to mental healthcare across vulnerable populations, support efforts to decrease acute mental health-related morbidities, and improve efficacy of health spending.
+\subsection{Status of Conducting Observational Health Research in Julia}
 
-Paragraph 2: What is OHDSI + HADES R ecosystem
+Over the history of JuliaHealth, a sub-ecosystem within the JuliaHealth ecosystem has developed which is dedicated to performing observational health research around the OMOP CDM.
+This sub-ecosystem is now mature enough to conduct internationally federated observational health research using tools housed within JuliaHealth and Julia exclusively.
+Additionally, due to the strong interoperability capabilities of Julia, rather than forgoing or completely reinventing aspects of pre-existing observational health research tools, the Julia observational health subecosystem builds on some of these innovations.
 
-Observational Health Data Sciences and Informatics (OHDSI) is a global, collaborative, interdisciplinary, and open-science network of researchers, healthcare professionals, students, patients, other individuals and organizations committed to large-scale health data analytics and generating reliable evidence, with the goal of promoting better health decisions and better care.
+As an example, OHDSI's HADES (Health Analytics Data-To-Evidence Suite) has been in existence for over $10$ years and has several rich packages.
+One of the premier tools within HADES is the ATLAS Shiny application which was developed to "support the design and execution of observational analyses to generate real world evidence from patient level observational data."
 
-Since current observational health data sources are aggregated at different levels and cannot be generalized to the overall population, the data is often incomplete, inaccurate, and prone to issues of bias, error handling, and multidimensional complexity.
-The goal of OHDSI is to produce reproducible, robust, and reliable evidence by developing standardized data formats and research protocol designs to enable large-scale collaborative research and improve the transparency and scalability of observational health research.
+\section{Describing an Observational Health Research Workflow Using JuliaHealth}
 
-The development of the Observational Medical Outcomes Partnership (OMOP) was created in 2009 to address concerns related to data type, study design, and data privacy concerns.
-The OMOP Common Data Model (CDM) is a "person-centered" information model with standardized relationships and structures with other event tables to create a longitudinal view of all clinical, healthcare-related events for each person.
+TODO: Discuss the "big idea" of what this workflow could look like
 
-OHDSI also develops open-source analytical and software tools, such as HADES and ATLAS, which features a comprehensive set of functions needed to conduct network studies from assessing demographic information to generating summary statistics and interactive dashboards.
-The best practices set forth by OHDSI has improved the quality and reliability of evidence in observational studies, and it improved network research study collaboration, supported large-scale analytics, and standardized data structures and medical vocabulary.
+\subsection{Cohort Expressions Usability}
 
-Health Analytics Data-to-Evidence Suite (HADES) is a collection of 20 R packages, developed by the OHDSI team, that interact directly with the OMOP CDM to support large-scale analytics, including cohort construction (CirceR, CohortGenerator), population-level estimation (CohortMethod), patient-level prediction (PatientLevelPrediction), and return statistics, figures, and tables.
+Broadly, ATLAS has become the de facto tool to develop patient phenotype definitions within observational health research contexts.
+These phenotype definitions have varying levels of complexity where they can be very simple (finding all patients with one health condition or diagnosis) or grow to be extremely complex (defining controlled, uncontrolled, and indeterminate patient cohorts based on medications, lab results, etc.)
+The computable phenotype definitions that ATLAS generates based on phenotype definitions are in a variety of serialization formats such as JSON or prepared SQL expressions available in multiple SQL flavors.
+(Examples of a computable phenotype definition serialized to JSON can be found in Appendix 1 and an exmaple of a serialization to PostgreSQL can be found in Appendix 2)
 
-Paragraph 3: Why tooling was needed
+The Julia package, OHDSICohortExpressions.jl, exists specifically to ingest these serialized expressions of computable phenotype definitions and create patient cohorts that match the given criterion in the definition.
+How this tool works is to 1) read and verify the serialized computable phenotype definition 2) reinterpret the serialization into the FunSQL.jl Domain Specific Language 3) return a prepared SQL statement against a given SQL database that queries an OMOP CDM database for the desired patient population(s) of interest.
+Once these prepared statements are made, researchers can execute the generated SQL against their OMOP CDM instance to generate cohorts for future analyses.
 
-\section{Methods}
+\subsection{Fine-Grain Creation of Cohorts}
 
-\subsection{Tooling for Observational Health Research}
+TODO: Discuss OMOPCDMCohortCreator.jl
 
-Paragraph 1: Introduction of novel packages
+\subsection{Working with Patient Databases}
 
-Paragraph 2: Development philosophy and algorithmic design
+TODO: Discuss DBConnector
 
-\subsection{Example Use Cases}
+\subsection{Building on Current Tools and State of the Art}
 
-\subsubsection{Minimal Characterization Study}
+TODO: Discuss OHDSIAPI.jl
 
-Paragraph 1: Condensed version of tutorial: https://juliahealth.org/OMOPCDMCohortCreator.jl/dev/tutorials/
+\section{Conducting a Small-Scale Observational Health Research Study Using JuliaHealth Tools}
 
-Paragraph 2: Example of using results (base off this paragraph): From here, the potential to calculate valuable metrics disease prevalence across niche populations is rich and opens up the possibility to do even more with these subpopulations one could generate (e.g.
-cross walking with survey panel data, area estimation, etc.).
-Furthermore, I have tested this but have not had an opportunity to write a tutorial on the matter, but these functionalities work with Distributed.jl to enable non-blocking asynchronous code execution (still working out exact approach/functionalities) to fully utilize Julia’s performance when working and transforming large amounts of data that is often found in observational studies.
-
-\subsubsection{Composability within Ecosystem}
-
-Paragraph 1 - 2: Distributed computing + Distributed.jl (maybe with Memoization) + Light benchmarking (?)
-
-Paragraph 3: Support for other databases + FunSQL.jl + SQL output
+  1. Showing DBConnection
+  2. Using ATLAS Cohort Definition
+  3. Pulling from OMOP CDM Cohort
+  4. Writing characterization code
+  5. Creating artifacts like plots and summary statistics
 
@@
 
-Paragraph 1: How it is being used
+\subsection{Advanced Features of the JuliaHealth Ecosystem}
 
-Paragraph 2: Next steps and Future - base off of this: As I was recently funded another year to continue my health disparities research (yay!), I have great ambitions to build out observational health research capacity in JuliaHealth through avenues such as:
+    1. Modularization of code and chaining
+    2. Distributed computing
+    3. FunSQL back-end
 
-Finalize a submission to JuliaCon Proceedings (need to finish writing tests across this portfolio and adding documentation in the coming weeks)
-Partnering on potentially running studies within the Julia community using these tools
-Scoping out better support for survey and panel data
-Building additional tooling around observational health research (such as determining fairness metrics, easily deployable interactive results explorers, interop with OHDSI tools, etc.)
-Lowering barrier to entry of prototyping large scale studies for researchers
+\subsection{Strong Composition of Machine Learning, Differential Equations, and JuliaHealth}
+
+\subsection{Potential Future Directions}
 
@@
 
-Paragraph 1: Final thoughts
+\section{Acknowledgements}
 
-\section{Acknowledgements}
+For this work, we would like to thank and recognize the following individuals and groups: Dr. Jon Duke and Charity Hilton for their early support,
 
@@
 \end{document}
+
+% Paragraph 1: Background on observational health research
+
+% Observational health data captures a comprehensive view of patient cohort populations, and they are often used to assess incidence rates, prevalence rates, risk factors, and prognosis associated with physical and mental health conditions and treatment interventions.
+By analyzing healthcare records and clinical events, observational health research can perform cohort characterization, population-estimation, and patient-level predictions to evaluate population health, target interventions and prevention measures, and ultimately transform healthcare policy.
+
+% However, current observational health and epidemiological surveys are limited by their data collection and administration methods, sample size representativeness, cohort phenotype definitions, and statistical/algorithmic methods to combine outside, disparate data sources.
+There is a need for timely, unbiased, accurate, and quality data regarding the mental illness and the ability detect and characterize trends in prevalence, severity, and the association with other chronic illnesses/morbidities.
+
+% This effort could help leading health officials implement better informed interventions that will increase equitable access to mental healthcare across vulnerable populations, support efforts to decrease acute mental health-related morbidities, and improve efficacy of health spending.
+
+% Paragraph 2: What is OHDSI + HADES R ecosystem
+
+% Observational Health Data Sciences and Informatics (OHDSI) is a global, collaborative, interdisciplinary, and open-science network of researchers, healthcare professionals, students, patients, other individuals and organizations committed to large-scale health data analytics and generating reliable evidence, with the goal of promoting better health decisions and better care.
+
+% Since current observational health data sources are aggregated at different levels and cannot be generalized to the overall population, the data is often incomplete, inaccurate, and prone to issues of bias, error handling, and multidimensional complexity.
+The goal of OHDSI is to produce reproducible, robust, and reliable evidence by developing standardized data formats and research protocol designs to enable large-scale collaborative research and improve the transparency and scalability of observational health research.
+
+% The development of the Observational Medical Outcomes Partnership (OMOP) was created in 2009 to address concerns related to data type, study design, and data privacy concerns.
+The OMOP Common Data Model (CDM) is a "person-centered" information model with standardized relationships and structures with other event tables to create a longitudinal view of all clinical, healthcare-related events for each person.
+
+% OHDSI also develops open-source analytical and software tools, such as HADES and ATLAS, which features a comprehensive set of functions needed to conduct network studies from assessing demographic information to generating summary statistics and interactive dashboards.
+The best practices set forth by OHDSI has improved the quality and reliability of evidence in observational studies, and it improved network research study collaboration, supported large-scale analytics, and standardized data structures and medical vocabulary.
+
+% Health Analytics Data-to-Evidence Suite (HADES) is a collection of 20 R packages, developed by the OHDSI team, that interact directly with the OMOP CDM to support large-scale analytics, including cohort construction (CirceR, CohortGenerator), population-level estimation (CohortMethod), patient-level prediction (PatientLevelPrediction), and return statistics, figures, and tables.
+
+% Paragraph 3: Why tooling was needed
+
+% Old Content
+
+% \subsection{Tooling for Observational Health Research}
+
+% Paragraph 1: Introduction of novel packages
+
+% Paragraph 2: Development philosophy and algorithmic design
+
+% \subsection{Example Use Cases}
+
+% \subsubsection{Minimal Characterization Study}
+
+% Paragraph 1: Condensed version of tutorial: https://juliahealth.org/OMOPCDMCohortCreator.jl/dev/tutorials/
+
+% Paragraph 2: Example of using results (base off this paragraph): From here, the potential to calculate valuable metrics disease prevalence across niche populations is rich and opens up the possibility to do even more with these subpopulations one could generate (e.g.
+cross walking with survey panel data, area estimation, etc.).
+Furthermore, I have tested this but have not had an opportunity to write a tutorial on the matter, but these functionalities work with Distributed.jl to enable non-blocking asynchronous code execution (still working out exact approach/functionalities) to fully utilize Julia’s performance when working and transforming large amounts of data that is often found in observational studies.
+
+% \subsubsection{Composability within Ecosystem}
+
+% Paragraph 1 - 2: Distributed computing + Distributed.jl (maybe with Memoization) + Light benchmarking (?)
+
+% Paragraph 3: Support for other databases + FunSQL.jl + SQL output
+
+% \section{Discussion}
+
+% Paragraph 1: How it is being used
+
+% Paragraph 2: Next steps and Future - base off of this: As I was recently funded another year to continue my health disparities research (yay!), I have great ambitions to build out observational health research capacity in JuliaHealth through avenues such as:
+
+% Finalize a submission to JuliaCon Proceedings (need to finish writing tests across this portfolio and adding documentation in the coming weeks)
+% Partnering on potentially running studies within the Julia community using these tools
+% Scoping out better support for survey and panel data
+% Building additional tooling around observational health research (such as determining fairness metrics, easily deployable interactive results explorers, interop with OHDSI tools, etc.)
+% Lowering barrier to entry of prototyping large scale studies for researchers

### 2023-09-30 paper.tex
Added more content, appendix, and links

@@
 These phenotype definitions have varying levels of complexity where they can be very simple (finding all patients with one health condition or diagnosis) or grow to be extremely complex (defining controlled, uncontrolled, and indeterminate patient cohorts based on medications, lab results, etc.)
-The computable phenotype definitions that ATLAS generates based on phenotype definitions are in a variety of serialization formats such as JSON or prepared SQL expressions available in multiple SQL flavors.
-(Examples of a computable phenotype definition serialized to JSON can be found in Appendix 1 and an exmaple of a serialization to PostgreSQL can be found in Appendix 2)
+The computable phenotype definitions that ATLAS generates based on phenotype definitions are in a variety of serialization formats such as JSON (see \ref{appendix:json}) or prepared SQL expressions available in multiple SQL flavors (see \ref{appendix:postgresql}).
 
@@
 
+As shown in this work, there is a rich opportunity for the JuliaHealth organization to become a strong presence within observational health research. 
+JuliaHealth is now sufficiently at a stage of maturity where potential future researchers can leverage JuliaHealth's tools to conduct further future research or build upon existing architecture to target specific needs.
+Some potential future directions that can be taken are as follows.
+
+, efforts can focus on developing a package for assessing fairness and bias in patient datasets, considering factors like data completeness, demographic variables, and fairness algorithms.
+Additionally, researchers can work on creating modular tools for disease pathway analysis, allowing quick insights into disease prevalence, incidence, and comorbidities within patient data.
+
+Moreover, there's potential to advance clinical informatics by developing a flexible clinical type system, supporting computable phenotype definitions and medical ontologies.
+The ecosystem can also benefit from tools to audit the fairness of phenotype definitions.
+Finally, as the landscape of data expands, integrating non-traditional sources like social media and environmental data into standardized patient records will be a critical frontier.
+In these ways, JuliaHealth is primed to drive innovation in observational health informatics, pushing the boundaries of data-driven healthcare research.
+
 \section{Conclusion}
@@
 \section{References}
+
+% APPENDIX
+\appendix
+
+\section{JSON Serialization of Computable Phenotype Definition}\label{appendix:json}
+
+An example of the JSON serialization of computable phenotype definitions can be found within the paper repository here:
+
+In particular, this example seeks to identify individuals with Type 2 Diabetes Mellitus within a given OMOP CDM.
+
+This was generated using ATLAS Version 2.12.1 as well as 
+WebAPI Version 2.12.1.
+A generic definition of the schema that was created for ATLAS 2.7.1 can be explored here: https://github.com/OHDSIBr/ATLAS-JSON-Schema/tree/2.7.4
+
+\section{PostgreSQL Serialization of Computable Phenotype Definition}\label{appendix:postgresql}
+
+An example of a PostgreSQL SQL serialization of computable phenotype definitions can be found within the paper repository here:
+
+In particular, this example seeks to identify individuals with Type 2 Diabetes Mellitus within a given OMOP CDM.
+
+This was generated using ATLAS Version 2.12.1 as well as 
+WebAPI Version 2.12.1.
 

### 2023-11-09 paper.tex
Updates from Overleaf

--- 2023-09-30 paper.tex
@@
 
-Observational health continues to be a growing field in health informatics research as electronic health records, patient medical claims, and other ancilliary patient data source become more readily computable and accessible to researchers.
+Observational health continues to be a growing field in health informatics research as electronic health records (EHR), patient medical claims, and other ancilliary patient data source become more readily computable and accessible to researchers.
 JuliaHealth is poised as an ecosystem to innovate within this area of research by bringing highly performant analytics approaches, composable solutions, and interoperable software that leverages prior state of the art.
@@
 \subsection{What Is JuliaHealth and Its Goals?}
+JuliaHealth is an open-source organization under the umbrella of the Julia language.
+It leverages the high-performance of the Julia language in the health field while for example handling large data sets and manipulating them.
+JuliaHealth aims to be researchers' companion in making use of the observational-health data as Julia is easy to be learnt syntax-wise.
 
@@
 \subsection{Fine-Grain Creation of Cohorts}
+Creating cohorts from a data-set as OMOP CDM has been an interest for researchers to test the outcome.
+OMOPCDMCohortCreator.jl is a package that manipulates a given database of OMOP CDM.
+The Package allows researchers to create cohorts using its tools.
+OMOPCDMCohortCreator.jl connects to the database using OMOPCDMConnector, then provides various tools to extract patients data of specific disease, race, gender, ...etc or filters them to achieve the desired cohort of patients using `executeAudit` function.
 
@@
 TODO: Discuss DBConnector
+
+In this paper we discuss data-sets of type OMOPCDM, DBConnector is a Julia package that connects databases on MySQL, SQLite or postgreSQL using one line command. DBConnector uses the DBInterface
 

### 2023-11-11 paper.tex
Revamped multiple sections and added coding example

@@
 \subsection{Fine-Grain Creation of Cohorts}
-Creating cohorts from a data-set as OMOP CDM has been an interest for researchers to test the outcome.
-OMOPCDMCohortCreator.jl is a package that manipulates a given database of OMOP CDM.
-The Package allows researchers to create cohorts using its tools.
-OMOPCDMCohortCreator.jl connects to the database using OMOPCDMConnector, then provides various tools to extract patients data of specific disease, race, gender, ...etc or filters them to achieve the desired cohort of patients using `executeAudit` function.
 
-TODO: Discuss OMOPCDMCohortCreator.jl
+Although OHDSICohortExpressions.jl is sufficient to translate existing phenotype definitions generated in ATLAS to executable queries, oftentimes, there is a tricky problem of how to best iterate on phenotype definitions.
+Often, one has to take an existing definition and either revise to a more narrow cohort or write manual code to build more niche cohorts after an initial definition is created.
+This refactoring of a phenotype definition can lead to conflating ideas within conversation between expert clinical reviewers, analysts, and health researchers.
+This broadly can be the notions of what we are exploring (the phenotype definition; often the responsibility of a clinical reviewer), what is possible to compute upon (the computable phenotype definition; often the responsibility of an analyst), and the goals of the overall study.
+For situations such as these, the tool OMOPCDMCohortCreator.jl was created.
+
+OMOPCDMCohortCreator.jl allows one to iteratively build a patient population quickly and easily.
+One can use patient pools developed from pre-existing cohorts defined off of phenotype definitions as a starting point in exploring patients or create new patient populations quickly without the need for a well-defined phenotype definition.
+This can enable analysts to more quickly understand what is inside their data assets, do "quick" analysis of a particular smaller research question, and enable a more distinct separation of work between a clinical expert reviewer and analyst teams.
+It is highly recommended to still create phenotype definitions about particular populations one explores.
+
+TODO: Add reference to phenotype paper
+TODO: Add small example showing iterative process 
+TODO: Reference capr somehow
 
@@
 
-TODO: Discuss DBConnector
+One of the biggest styming factors within not only the Julia ecosystem, but observational health research in general has been the difficulty of working with different databases.
+Although the OMOP CDM has been great for standardizing \textit{how} real world data is handled, there is a lack of consensus on \textit{what} is needed to handle these data assets.
+For a variety of reasons, different groups will choose different database architectures (perhaps PostgreSQL for improved permission support, DuckDB for performance, or even SQLite for ease of use) but all are subtly different enough to create headaches in seamlessly working across different research sites.
 
-In this paper we discuss data-sets of type OMOPCDM, DBConnector is a Julia package that connects databases on MySQL, SQLite or postgreSQL using one line command. DBConnector uses the DBInterface
+The way that the Julia ecosystem handles this is via the creation of a common database interface called the DBInterface.
+Specific Julia database packages can implement dispatches for this interace allowing users to have a much better experience working across databases.
+In spite of the such benefits, for some users (and especially within the context of observational health studies), it can beneficial to have a more prescriptive interface where one just needs to pass perhaps a password, username, database name, and schema.
+For that reason, we DBConnector.jl was created.
+
+Although this is not an explicit JuliaHealth package, it was a pain point that was experienced throughout the process of observational health research across multiple partner sites.
+The promise of DBConnector.jl is that with one line command, users can connect simply, and in a common (albeit opinionated) manner, to databases.
+DBConnector unifies Julia database packages that implement the DBInterface to achieve this level of simplified connection making.
+
+Finally, if someone needs more fine-grained control over a particular database connection, they can choose to eschew this package.
+Instead, they can switch to the specific Julia database package that would give them the additional control.
+Although they'll lose the simplicity of this package, core required functionality is not lost when working with databases.
+
+TODO: Add database connector example
+TODO: Reference database connector example
+TODO: Reference that it was inspired by database connector OHDSI package
+TODO: Mention package extensions utility
 
@@
 
-  1. Showing DBConnection
-  2. Using ATLAS Cohort Definition
-  3. Pulling from OMOP CDM Cohort
-  4. Writing characterization code
-  5. Creating artifacts like plots and summary statistics
+With the JuliaHealth tools that have now been created, we can now demonstrate an observational health study workflow that would be used in practice.
+For this paper, we used two datasets: a synthetic patient dataset called \textit{Eunomia} and the MIMIC III patient dataset which had been converted into the OMOP CDM to illustrate the workflow provided by the Observational Health sub-ecosystem.
+Reproducing the workflow using the \textit{Eunomia} dataset is made readily available at the repository here:
+
+TODO: Add tutorial to separate repository
+TODO: Create tutorial about Eunomia
+TODO: Reference Eunomia
+
+\textit{Eunomia} will be used to drive the workflow example looking at patients who have ever had a history of strep throat.
+Results generated by the \textit{Eunomia} dataset will be shown for pedagogical purposes although the actual results of that analysis will be nonsensical given the synthetic nature of the database.
+However, to complement this, a similar study will be done on the MIMIC III dataset characterizing individuals with a history of Type 2 Diabetes Mellitus to encourage discussion on what is practically possible with these tools.
+
+\subsection{Connecting to a Database}
+
+\begin{verbatim}
+  import HealthSampleData: Eunomia
+  import OMOPCDMCohortCreator as OCC
+
+  eunomia = Eunomia()
+  conn = DB(eunomia)
+
+  occ.GenerateDatabaseDetails(
+      :sqlite,
+      "main"
+  )
+
+  occ.GenerateTables(conn)
+\end{verbatim}
+
+TODO: Replace with DBConnector syntax
+
+\subsection{Bridging between OHDSI and JuliaHealth Communities}
+
+Briding between tools used in the OHDSI and JuliaHealth communities is now exceptionally straightforward.
+
+\includegraphics{./atlas_gui.png}
+
+\subsection{Using a Cohort Definition To Build an Initial Cohort}
+
+After having a strep throat phenotype definition defined using ATLAS, we can then execute the resulting computable phenotype definition against the \textit{Eunomia} database.
+
+\begin{verbatim}
+  using JSON3 
+  using OHDSICohortExpressions
+
+  import DBInterface as DBI
+
+  cohort = JSON3.read("strep_throat.json")
+  cohort_expression = cohort[:items][1][:expression]
+
+  model = Model(cdm_version=v"5.3.1", cdm_schema="main",
+                       vocabulary_schema="main", results_schema="main",
+                       target_schema="main", target_table="cohort");
+
+  sql = translate(cohort_expression, dialect=:sqlite, model=model, cohort_definition_id=1);
+
+  [DBI.execute(conn, sub_query) for sub_query in split(sql, ";")[1:end-1]]
+\end{verbatim}
+
+A quick query of the cohort table (where the resulting subjects were stored) shows $\approx 1600$ from the dataset match this criteria.
+
+TODO: Clean up the code a little nicer
+TODO: Make sure the JSON3 reading of a cohort works properly
+
+\subsection{Characterizing Patient Populations}
+
+\begin{verbatim}
+  patients = DBI.execute(conn, "SELECT subject_id AS person_id FROM cohort;") |> DF.DataFrame
+\end{verbatim}
+
+TODO: Replace this line from Jay's PR for getting cohorts
+
+\begin{verbatim}
+  patients_race = occ.GetPatientRace(patients.person_id, conn)
+  patients_gender = occ.GetPatientGender(patients.person_id, conn)
+
+  patients_age_group = occ.GetPatientAgeGroup(
+    patients.person_id, 
+    conn; 
+    age_groupings = [[age, age + 4] for age in 0:5:119]
+  )
+\end{verbatim}
+
+\subsection{Summary of Results}
+
+\begin{verbatim}
+  import DataFrames as DF
+
+  patients_characterized = DF.outerjoin(patients_race,
+                                              patients_gender,
+                                              patients_age_group;
+                                              on = :person_id, 
+                                              matchmissing = :equal)
+  patients_characterized = patients_characterized[:, DF.Not(:person_id)]
+  patient_groups = DF.groupby(patients_characterized, [:race_concept_id, :gender_concept_id, :age_group])
+  patient_groups = DF.combine(patient_groups, DF.nrow => :count)
+\end{verbatim}
 

### 2023-11-13 paper.tex
Added rough crude prevalence section

@@
 
-\subsection{Summary of Results}
+\subsection{Calculating Crudge Prevalence Rate}
 
@@
 \end{verbatim}
+
+Using this analytical approach, tabulation can be quickly conducted across patients.
+Additionally, throughout the process, one can perform relevant analysis methods.
+Here, we'll conduct a crude prevalence calculation across the patient cohort.
+
+\subsection{Summary of Results}
+
+As seen in Table 1 and Table 2, informative results can be rapidly generated by this workflow relevant to observational health research.
+Although the values shown in Table 2 are synthetic and the crude prevalence values themselves are meaningless, we can take a this approach and apply it directly to the MIMIC III dataset.
+In Table 3, we can calculate across this dataset meaningful statistics that could later be further analyzed and drive further questioning (such as ...).
 

### 2023-11-14 paper.tex
Section on future work for JuliaHealth

@@
 
-\subsection{Calculating Crudge Prevalence Rate}
+\subsection{Calculating Crude Prevalence Rate}
 
@@
 
-, efforts can focus on developing a package for assessing fairness and bias in patient datasets, considering factors like data completeness, demographic variables, and fairness algorithms.
-Additionally, researchers can work on creating modular tools for disease pathway analysis, allowing quick insights into disease prevalence, incidence, and comorbidities within patient data.
+Due to the strong composition that exists within the JuliaHealth and broader Julia ecosystem, solutions may already exist for solving various health informatics related research.
+However, what has consistently been a problem within the Julia ecosystem is the lack of sufficient documentation dedicated to describing these novel compositions and their application to various endeavors.
+Future efforts can be dedicated to taking existing tools and reframing them in a JuliaHealth context to assist newcomers to Julia and JuliaHealth in determining how to explore and address problems that they are interested in.
+
+Additionally, due to the composable and modular nature of the observational health sub-ecosystem, research analyses and visualizations can be decoupled and interchanged with one another.
+Whereas in other ecosystems that make technologies for tools such as dashboards tightly coupled to analytics regimens, the subecosystem is mature enough to enable new users and developers to create analytics tools that can be separately woven into more graphical user interfaces.
+For example, treatment pathways is a novel observational health approach to understand and visualize the care a patient receives within a care setting.
+Only a few tools exist to visualize such hidden patient narratives but with the JuliaHealth ecosystem, the possibility of making tools that can interrogate patient care pathways more deeply could be fully realized.
+
+Furthermore, again exploiting the composition within the JuliaHealth ecosystem, as mentioned in the context of JuliaHealth and machine learning applications, future efforts could focus on developing fairness and bias auditing tools in patient datasets.
+These tools could consider factors commonly explored in health equity literature like data completeness, demographic variables, and fairness algorithms.
 

### 2023-11-24 paper.tex
Added example summary output table and generating a workflow companion script
Finished first full draft of paper

@@
 Additionally, due to the strong interoperability capabilities of Julia, rather than forgoing or completely reinventing aspects of pre-existing observational health research tools, the Julia observational health subecosystem builds on some of these innovations.
+
+Observational Health Data Sciences and Informatics (OHDSI) is a global, collaborative, interdisciplinary, and open-science network of researchers, healthcare professionals, students, patients, other individuals and organizations committed to large-scale health data analytics and generating reliable evidence, with the goal of promoting better health decisions and better care.
+
+Since current observational health data sources are aggregated at different levels and cannot be generalized to the overall population, the data is often incomplete, inaccurate, and prone to issues of bias, error handling, and multidimensional complexity.
+The goal of OHDSI is to produce reproducible, robust, and reliable evidence by developing standardized data formats and research protocol designs to enable large-scale collaborative research and improve the transparency and scalability of observational health research.
+
+The development of the Observational Medical Outcomes Partnership (OMOP) was created in 2009 to address concerns related to data type, study design, and data privacy concerns.
+The OMOP Common Data Model (CDM) is a "person-centered" information model with standardized relationships and structures with other event tables to create a longitudinal view of all clinical, healthcare-related events for each person.
+
+OHDSI also develops open-source analytical and software tools, such as HADES and ATLAS, which features a comprehensive set of functions needed to conduct network studies from assessing demographic information to generating summary statistics and interactive dashboards.
+The best practices set forth by OHDSI has improved the quality and reliability of evidence in observational studies, and it improved network research study collaboration, supported large-scale analytics, and standardized data structures and medical vocabulary.
+
+Health Analytics Data-to-Evidence Suite (HADES) is a collection of 20 R packages, developed by the OHDSI team, that interact directly with the OMOP CDM to support large-scale analytics, including cohort construction (CirceR, CohortGenerator), population-level estimation (CohortMethod), patient-level prediction (PatientLevelPrediction), and return statistics, figures, and tables.
 
@@
 
-\includegraphics{./atlas_gui.png}
+% \includegraphics{./atlas_gui.png}
 
@@
 
+\begin{table}[!ht]
+    \centering
+    \begin{tabular}{|l|l|l|l|l|l|}
+    \hline
+        Race & Gender & Age Group & Cohort Count & Total Count & Prev. (\%) \\ \hline
+        Black or African American & Female & 60 - 64 & 13 & 21 & 61.90 \\ \hline
+        White & Male & 70 - 74 & 68 & 96 & 70.83 \\ \hline
+        White & Male & 45 - 49 & 47 & 74 & 63.51 \\ \hline
+        White & Female & 50 - 54 & 59 & 111 & 53.15 \\ \hline
+        Black or African American & Female & 55 - 59 & 7 & 19 & 36.84 \\ \hline
+        White & Male & 75 - 79 & 45 & 56 & 80.35 \\ \hline
+        White & Female & 55 - 59 & 62 & 97 & 63.91 \\ \hline
+        White & Female & 45 - 49 & 41 & 92 & 44.56 \\ \hline
+        Asian & Female & 80 - 84 & 4 & 5 & 80.0  \hline
+        Asian & Male & 80 - 84 & 6 & 10 & 60.00 \\ \hline
+        Black or African American & Female & 45 - 49 & 10 & 23 & 43.47 \\ \hline
+        No matching concept & Female & 60 - 64 & 29 & 39 & 74.35 \\ \hline
+        Black or African American & Male & 65 - 69 & 12 & 18 & 66.66 \\ \hline
+    \end{tabular}
+\end{table}
+
 \subsection{Summary of Results}
@@
     2. Distributed computing
-    3. FunSQL back-end
 
-\subsection{Strong Composition of Machine Learning, Differential Equations, and JuliaHealth}
+\subsubsection{Modularization of Code and Chaining}
+
+An exciting emergent aspect of the JuliaHealth ecosystem that came about as a result of developing OMOPCDMCohortCreator.jl was the prospect of radical modularization and chaining of package functionalities together.
+For example, taking our workflow example from the earlier section where we characterized patients by race, gender, and age group, we can do a few "tricks" to modularize these characterization functions as follows:
+
+We can leverage a unique property of Julia in that Julia supports partial evaluation of functions (also known as currying) to simplify repeated function calls to the same function over and over again.
+First, we'll take our original functions and fix a connection object to each of our cofactor functions.
+This simplifies having to continuously pass forward a connection object and opens up the ability to chain together these functions simply:
+
+TODO: Check that I equated currying correctly here.
+
+\begin{verbatim}
+import Base:
+  Fix2
+
+CGetPatientRace = Base.Fix2(occ.GetPatientRace, conn)
+CGetPatientGender = Base.Fix2(occ.GetPatientGender, conn)
+CGetPatientAgeGroup = Base.Fix2(occ.GetPatientAgeGroup, conn)
+\end{verbatim}
+
+NOTE: The "C" prefix for each of our curried functions denotes that it is curried.
+
+Second, as OMOPCDMCohortCreator.jl is built on top of DataFrames.jl and one has the option of creating a DataFrame with the result from every OCC function, we can use the package, Chain.jl.
+This package abstracts away some of the explicit functionalities of DataFrames.jl but instead gives much easier ways to define a flow of functions:
+
+\begin{verbatim}
+import Chain:
+  @chain
+
+CofactorPatients(patients) = @chain patients begin
+    CGetPatientRace_fixed
+    CGetPatientGender_fixed
+    CGetPatientAgeGroup_fixed
+end
+\end{verbatim}
+
+\begin{verbatim}
+CharacterizePatients(patients) = @chain patients begin
+    _[:, DF.Not(:person_id)]
+    DF.groupby(_, names(_))
+    DF.combine(_, DF.nrow => :count)
+    occ.ExecuteAudit
+end
+\end{verbatim}
+
+This gives one the ability to simply reason about an analysis in a very straightforward and readable manner:
+
+\begin{verbatim}
+RunStudy(patients) = @chain patients begin
+    CofactorPatients
+    CharacterizePatients
+end
+\end{verbatim}
+
+In this way, future researchers using tools such as OMOPCDMCohortCreator.jl can create these more convenient abstractions.
+In fact, it could be a point of exploration in the future to build on top of OMOPCDMCohortCreator.jl even more abstracted and reusable functionalities to simplify ease of use while also enabling more readily auditable functionality.
+Finally, as these functions have now been fully modularized, these can now be readily distributed using Julia's distributed computing methods:
+
+\begin{verbatim}
+import Distributed:
+  @distributed
+
+result = @distributed (append!) for i in 1:10
+   vals = 
+      RunStudy(strep_patients[i*100 + 1:(i+1)*100, :])
+   [vals]
+end
+
+vcat(result...) |> audit_patient_groups
+\end{verbatim}
+
+Distributed computing here then allows a researcher to run multiple investigations at once as well as optimize performance to a given database.
+
+\subsubsection{Auto-Generating SQL from JuliaHealth Commands}
+
+One crucial aspect of the observational health sub-ecosystem is that it is unavoidable to work with databases that contain patient health information.
+For better or worse, the tool that is used to best interact with these databases are SQL dialects that loosely follow consistently the ANSI SQL guidance.
+To accomodate for this, several of the core tools developed in this sub-ecosystem generate SQL that can be used in the case where Julia may not be available in a given environment housing patient health data.
+As an example, \textit{OMOPCDMCohortCreator.jl} functions build upon \textit{FunSQL.jl} allowing one to use a function such `GetPatientAgeGroup` without a connection object which would result in SQL.
+
+\begin{verbatim}
+  GetPatientAgeGroup([1]; age_groupings = [[0, 40], [41, 80]])
+\end{verbatim}
+
+\begin{verbatim}
+  SELECT
+    "PERSON_2"."person_id",
+    (CASE 
+      WHEN ("PERSON_2"."age" < 41) 
+        THEN '0 - 40' 
+      WHEN ("PERSON_2"."age" < 81) 
+        THEN '41 - 80' END) AS "age_group"
+  FROM (
+    SELECT
+      "PERSON_1"."person_id",
+      (2023 - "PERSON_1"."year_of_birth") AS "age"
+    FROM "PERSON" AS "PERSON_1"
+    WHERE ("PERSON_1"."person_id" IN (1))
+  ) AS "PERSON_2"
+\end{verbatim}
+
+\subsection{Strong Composition within JuliaHealth}
+
+Within Julia, one aspect of the language that is often viewed as a technical problem within other languages emerges more as a social problem within Julia.
+That problem is the problem of strong composition within Julia.
+What composition within Julia is thought to be generally is that a user of Julia packages A and B can combine these packages oftentimes seamlessly together in such a way as to solve a problem that the original designers of A and B did not think about.
+This has the virtue of Julia packages being strongly flexible to suit problems at hand but a discoverability problem where users developing these novel compositions may not share about these combinations.
+As a result, there arise a paradox wherein strongly compositional systems, like Julia, simultaneously give rise to obfuscational logic.
+
+Within JuliaHealth, this problem is an active area of exploration.
+During the summer of 2023, this problem was one of the focuses of Google Summer of Code mentee, Fareeda Abdelazeez.
+Originally, the project dealt with determining if another package in the JuliaHealth ecosystem was necessary to support prediction for patient outcomes.
+Instead, what was realized instead is that what should be further investigated is novel compositions that can be viewed through the lens of JuliaHealth.
+As shown in this workflow, although there were some packages that were made specifically for JuliaHealth, other packages that were not JuliaHealth specific were used alongside of it that composed flawlessly with the rest of the JuliaHealth observational health subecosystem.
 
@@
 
+In conclusion, this paper has demonstrated the capabilities of the JuliaHealth observational health sub-ecosystem and its potential to further investigations in observational health research.
+Although it is an ecosystem that is relatively new, its robustness allows for novel investigations to be pursued within this domain of research. 
+Additionally, due to the strongly compositional nature of the Julia ecosystem, these JuliaHealth tools can readily compose with tooling to support in-depth statistical analyses, mathematical modeling, and visualization capacities.
+
+The deep dive into the JuliaHealth ecosystem unravels its arsenal of tools tailored for observational health research, meticulously designed to navigate the complexities of handling OMOP CDM data.
+Notably, the sub-ecosystem within JuliaHealth presents mature, internationally federated research capabilities, building upon existing innovations like OHDSI's HADES, while offering novel tools such as OMOPCDMCohortCreator.jl.
+In considering future directions, the paper posits the promising avenues for JuliaHealth, envisioning expanded documentation, the development of visualization tools, and integration of machine learning applications for fairness auditing.
+The ecosystem's potential to revolutionize clinical informatics by integrating non-traditional data sources, auditing phenotype definitions, and fostering innovation in data-driven healthcare research is paramount.
+In essence, the JuliaHealth ecosystem stands as a catalyst for transformative observational health research, poised to push the boundaries of innovation, collaboration, and standardization in the pursuit of better healthcare outcomes.
+As it continues to evolve, JuliaHealth emerges as a beacon of progress in the dynamic landscape of health informatics.
+
 \section{Acknowledgements}
 
-For this work, we would like to thank and recognize the following individuals and groups: Dr. Jon Duke and Charity Hilton for their early support,
+For this work, we would like to thank and recognize the following individuals and groups:
+
+\begin{enumerate}
+
+  \item Dr. Jon Duke for his early support
+  \item Kristin Kostka for being an early champion in trying novel ideas
+  \item Dilum Aluthge for helping this work find a home in JuliaHealth
+  \item Adam Black for thinking through potential approaches
+  \item The Roux Institute's OHDSI Center for their encouragement
+  \item The JuliaHealth community and contributors for fearlessly trying to push boundaries on what observational health informatics can truly be
+
+\end{enumerate}
 
@@
 \end{document}
-
-% Paragraph 1: Background on observational health research
-
-% Observational health data captures a comprehensive view of patient cohort populations, and they are often used to assess incidence rates, prevalence rates, risk factors, and prognosis associated with physical and mental health conditions and treatment interventions.
-By analyzing healthcare records and clinical events, observational health research can perform cohort characterization, population-estimation, and patient-level predictions to evaluate population health, target interventions and prevention measures, and ultimately transform healthcare policy.
-
-% However, current observational health and epidemiological surveys are limited by their data collection and administration methods, sample size representativeness, cohort phenotype definitions, and statistical/algorithmic methods to combine outside, disparate data sources.
-There is a need for timely, unbiased, accurate, and quality data regarding the mental illness and the ability detect and characterize trends in prevalence, severity, and the association with other chronic illnesses/morbidities.
-
-% This effort could help leading health officials implement better informed interventions that will increase equitable access to mental healthcare across vulnerable populations, support efforts to decrease acute mental health-related morbidities, and improve efficacy of health spending.
-
-% Paragraph 2: What is OHDSI + HADES R ecosystem
-
-% Observational Health Data Sciences and Informatics (OHDSI) is a global, collaborative, interdisciplinary, and open-science network of researchers, healthcare professionals, students, patients, other individuals and organizations committed to large-scale health data analytics and generating reliable evidence, with the goal of promoting better health decisions and better care.
-
-% Since current observational health data sources are aggregated at different levels and cannot be generalized to the overall population, the data is often incomplete, inaccurate, and prone to issues of bias, error handling, and multidimensional complexity.
-The goal of OHDSI is to produce reproducible, robust, and reliable evidence by developing standardized data formats and research protocol designs to enable large-scale collaborative research and improve the transparency and scalability of observational health research.
-
-% The development of the Observational Medical Outcomes Partnership (OMOP) was created in 2009 to address concerns related to data type, study design, and data privacy concerns.
-The OMOP Common Data Model (CDM) is a "person-centered" information model with standardized relationships and structures with other event tables to create a longitudinal view of all clinical, healthcare-related events for each person.
-
-% OHDSI also develops open-source analytical and software tools, such as HADES and ATLAS, which features a comprehensive set of functions needed to conduct network studies from assessing demographic information to generating summary statistics and interactive dashboards.
-The best practices set forth by OHDSI has improved the quality and reliability of evidence in observational studies, and it improved network research study collaboration, supported large-scale analytics, and standardized data structures and medical vocabulary.
-
-% Health Analytics Data-to-Evidence Suite (HADES) is a collection of 20 R packages, developed by the OHDSI team, that interact directly with the OMOP CDM to support large-scale analytics, including cohort construction (CirceR, CohortGenerator), population-level estimation (CohortMethod), patient-level prediction (PatientLevelPrediction), and return statistics, figures, and tables.
-
-% Paragraph 3: Why tooling was needed
-
-% Old Content
-
-% \subsection{Tooling for Observational Health Research}
-
-% Paragraph 1: Introduction of novel packages
-
-% Paragraph 2: Development philosophy and algorithmic design
-
-% \subsection{Example Use Cases}
-
-% \subsubsection{Minimal Characterization Study}
-
-% Paragraph 1: Condensed version of tutorial: https://juliahealth.org/OMOPCDMCohortCreator.jl/dev/tutorials/
-
-% Paragraph 2: Example of using results (base off this paragraph): From here, the potential to calculate valuable metrics disease prevalence across niche populations is rich and opens up the possibility to do even more with these subpopulations one could generate (e.g.
-cross walking with survey panel data, area estimation, etc.).
-Furthermore, I have tested this but have not had an opportunity to write a tutorial on the matter, but these functionalities work with Distributed.jl to enable non-blocking asynchronous code execution (still working out exact approach/functionalities) to fully utilize Julia’s performance when working and transforming large amounts of data that is often found in observational studies.
-
-% \subsubsection{Composability within Ecosystem}
-
-% Paragraph 1 - 2: Distributed computing + Distributed.jl (maybe with Memoization) + Light benchmarking (?)
-
-% Paragraph 3: Support for other databases + FunSQL.jl + SQL output
-
-% \section{Discussion}
-
-% Paragraph 1: How it is being used
-
-% Paragraph 2: Next steps and Future - base off of this: As I was recently funded another year to continue my health disparities research (yay!), I have great ambitions to build out observational health research capacity in JuliaHealth through avenues such as:
-
-% Finalize a submission to JuliaCon Proceedings (need to finish writing tests across this portfolio and adding documentation in the coming weeks)
-% Partnering on potentially running studies within the Julia community using these tools
-% Scoping out better support for survey and panel data
-% Building additional tooling around observational health research (such as determining fairness metrics, easily deployable interactive results explorers, interop with OHDSI tools, etc.)
-% Lowering barrier to entry of prototyping large scale studies for researchers

### 2023-11-24 paper.tex
Took care of some outstanding todos

@@
 
-\section{Describing an Observational Health Research Workflow Using JuliaHealth}
+\section{Describing Common Observational Health Research Tasks within the Julia Ecosystem}
 
-TODO: Discuss the "big idea" of what this workflow could look like
+This section guides readers through common steps part of observational health research and to what capacity the Julia ecosystem supports these tasks.
+Here are the steps covered
 
-\subsection{Cohort Expressions Usability}
+\begin{enumerate}
+  \item \textbf{Using Phenotype Definitions} - brief overview of phenotype definitions and how to work with them
+  \item \textbf{Rapid Cohort Iteration and Creation} - how to use JuliaHealth tools to iteratively develop patient cohorts to explore
+  \item \textbf{Working with Patient Databases} - current approaches to working with different patient databases
+  \item \textbf{Ecosystem Interoperability} - how one can bridge between JuliaHealth's ecosystem to other research ecosystems
+\end{enumerate}
+
+\subsection{Using Phenotype Definitions}
 
@@
 
-\subsection{Fine-Grain Creation of Cohorts}
+\subsection{Rapid Cohort Iteration and Creation}
 
 Although OHDSICohortExpressions.jl is sufficient to translate existing phenotype definitions generated in ATLAS to executable queries, oftentimes, there is a tricky problem of how to best iterate on phenotype definitions.
+\cite{zelkoDevelopingRobustComputable2023}
 Often, one has to take an existing definition and either revise to a more narrow cohort or write manual code to build more niche cohorts after an initial definition is created.
@@
 
-TODO: Add reference to phenotype paper
 TODO: Add small example showing iterative process 
@@
 
-\subsection{Building on Current Tools and State of the Art}
+\subsection{Ecosystem Interoperability}
 
@@
 
-With the JuliaHealth tools that have now been created, we can now demonstrate an observational health study workflow that would be used in practice.
+With the JuliaHealth tools that have been created, we can now demonstrate an observational health study workflow that would be used in practice.
 For this paper, we used two datasets: a synthetic patient dataset called \textit{Eunomia} and the MIMIC III patient dataset which had been converted into the OMOP CDM to illustrate the workflow provided by the Observational Health sub-ecosystem.

### 2023-11-26 paper.tex
Worked on interoperability todos

@@
 The best practices set forth by OHDSI has improved the quality and reliability of evidence in observational studies, and it improved network research study collaboration, supported large-scale analytics, and standardized data structures and medical vocabulary.
-
-Health Analytics Data-to-Evidence Suite (HADES) is a collection of 20 R packages, developed by the OHDSI team, that interact directly with the OMOP CDM to support large-scale analytics, including cohort construction (CirceR, CohortGenerator), population-level estimation (CohortMethod), patient-level prediction (PatientLevelPrediction), and return statistics, figures, and tables.
 
@@
 
+DBConnector.jl was inspired by the package \textit{DatabaseConnector} (Link: https://github.com/OHDSI/DatabaseConnector)
+
 TODO: Add database connector example
 TODO: Reference database connector example
-TODO: Reference that it was inspired by database connector OHDSI package
 TODO: Mention package extensions utility
@@
 
-TODO: Discuss OHDSIAPI.jl
+In order to more effectively bridge the JuliaHealth ecosystem with ongoing efforts in other ecosystems, Julia provides methods to allow one to seamlessly interoperate with not only other ecosystems but entirely different languages as well.
+For example, through the work of packages like RCall.jl or PythonCall.jl, one can utilize R or Python packages and interfaces directly within Julia. \cite{PythonCall.jl} \cite{RCallJl}
+This embedding allows one to patch one's workflow where proper tools are missing with the equivalent package in Python or R.
+
+One particular ecosystem is the OHDSI Health Analytics Data-to-Evidence Suite (HADES) ecosystem where the majority of its tools are written in a combination of R and Java.
+HADES is a collection of approximately 20 packages, developed by the OHDSI community to interact directly with the OMOP CDM to support large-scale analytics.
+The goals of HADES is much the same as the JuliaHealth observational health subecosystem such as supporting cohort construction, population-level estimation, patient-level prediction, and calculating other sorts of relevant research artifacts.
+
+For HADES tools that expose APIs, rather than having to somehow wrap around this with a tool like RCall.jl, instead, HADES software can continue to run as a service and OHDSIAPI.jl can be used.
+OHDSIAPI.jl is a Julia wrapper around a variety of OHDSI API-based services that can directly interface with such services.
+This can be a great bridging mechanism so that observational health researchers can continue doing work and still interact with necessary research endpoints as needed without interrupting their workflow.
 

### 2023-12-03 header.tex
Added Jay Sanjay as coauthor

% **************GENERATED FILE, DO NOT EDIT**************

\title{Developing Observational Health Research Capacity in the Julia Ecosystem}

\author[1]{Jacob S. Zelko}
\author[3]{Fareeda Abdelazeez}
\author[1, 2]{Malina Hy}
\author[1, 2]{Varshini Chinta}
\author[3]{Jay Sanjay}
\affil[1]{Georgia Tech Research Institute}
\affil[2]{Georgia Institute of Technology}
\affil[3]{Foobar}

\keywords{Observational Health, OMOP Common Data Model, Database Management, Characterization}

\hypersetup{
pdftitle = {Developing Observational Health Research Capacity in the Julia Ecosystem},
pdfsubject = {JuliaCon 2019 Proceedings},
pdfauthor = {Jacob S. Zelko, Fareeda Abdelazeez, Malina Hy, Varshini Chinta, Jay Sanjay},
pdfkeywords = {Observational Health, OMOP Common Data Model, Database Management, Characterization},
}

### 2023-12-03 paper.tex
Squashing TODOs across the paper

@@
 
-\subsection{What Is JuliaHealth and Its Goals?}
-JuliaHealth is an open-source organization under the umbrella of the Julia language.
-It leverages the high-performance of the Julia language in the health field while for example handling large data sets and manipulating them.
-JuliaHealth aims to be researchers' companion in making use of the observational-health data as Julia is easy to be learnt syntax-wise.
+\subsection{What Is JuliaHealth?}
+
+DONE:
+JuliaHealth is an open-source collective dedicated to bringing the Julia programming language to various avenues of health research domains such as observational health, medical imaging, standards interoperability, etc.
+It leverages the strengths of the Julia language to introduce new approaches to these spaces, augment and improve existing workflows, and provide a central community structure for health researchers both part of and entering into the Julia ecosystem.
+Eventually, JuliaHealth will host a variety of subecosystems and associated resources that will enable novel research to be conducted within Julia.
 
@@
 
-Over the history of JuliaHealth, a sub-ecosystem within the JuliaHealth ecosystem has developed which is dedicated to performing observational health research around the OMOP CDM.
-This sub-ecosystem is now mature enough to conduct internationally federated observational health research using tools housed within JuliaHealth and Julia exclusively.
-Additionally, due to the strong interoperability capabilities of Julia, rather than forgoing or completely reinventing aspects of pre-existing observational health research tools, the Julia observational health subecosystem builds on some of these innovations.
+JuliaHealth itself is a loose collective of various research software engineers, health researchers, and hobbyist programmers that had its origins in $\approx 2020$.
+Being that JuliaHealth is a federated network of contributors, varying aspects of the ecosystem have developed at differing rates as a function of multiple different factors such as developer time, workplace incentives, and more.
+One area of the subecosystems within JuliaHealth that has matured significantly over the years has been the observational health subecosystem (referred to as OHS for the remainder of this paper).
 
-Observational Health Data Sciences and Informatics (OHDSI) is a global, collaborative, interdisciplinary, and open-science network of researchers, healthcare professionals, students, patients, other individuals and organizations committed to large-scale health data analytics and generating reliable evidence, with the goal of promoting better health decisions and better care.
+Although difficult to date, it can said that OHS had its inception in work done by Evans and Simonov with efforts such as \textit{FunOHDSI.jl}, \textit{OHDSICohortExpressions}, and \textit{FunSQL.jl} around $2021$  \cite{kirill_simonov_2023_7705325}
+Since then, Zelko et.
+al.
+have spearheaded most recent developments within OHS via the creation of a variety of tools and demonstrated their effectiveness in using these tools for novel observational health research endeavors \cite{zelko2022pilot}
+Through such efforts, OHS is becoming further capable of executing novel observational health research although, understandably, not as mature as similar ecosystems.
 
+TODO: Cite FunOHDSI
+TODO: Cite OHDSICohortExpressions
+TODO: Cite OHDSI Symposium poster on JuliaHealth
+TODO: Register OHDSI Symposium poster to Zenodo
+TODO: Cite JuliaCon 2023 presentation
+TODO: Register JuliaCon 2023 presentation to Zenodo
+
+In particular, the strongest influence and inspiration to OHS is the OHDSI open-science network.
+OHDSI (Observational Health Data Sciences and Informatics) is a global, collaborative, interdisciplinary, and open network of researchers, healthcare professionals, students, patients, and other entities committed to large-scale health data analytics and generating reliable evidence, with the goal of promoting better health decisions and better care.
 Since current observational health data sources are aggregated at different levels and cannot be generalized to the overall population, the data is often incomplete, inaccurate, and prone to issues of bias, error handling, and multidimensional complexity.
@@
 
-The development of the Observational Medical Outcomes Partnership (OMOP) was created in 2009 to address concerns related to data type, study design, and data privacy concerns.
-The OMOP Common Data Model (CDM) is a "person-centered" information model with standardized relationships and structures with other event tables to create a longitudinal view of all clinical, healthcare-related events for each person.
+OHDSI had its inception in the development of the Observational Medical Outcomes Partnership (OMOP) that was created in 2009 by the US Food and Drug Administration to address concerns related to data type, study design, and data privacy concerns of "Real World Data" \cite{ohdsi2019book} \cite{FDARealWorldEvidence}
+What could be considered the magnum opus of the OMOP group was the development of what is called the OMOP Common Data Model (OMOP CDM) to house and ingest this sort of data in a consistent format that has been optimized to assist in extracting meaningful population health information.
+In brief, the OMOP Common Data Model (CDM) is a "person-centered" model that has standardized common events in a given patient encounter with database tables that represent the unique structures found in patient events.
+This creates a longitudinal view of all healthcare-related events per person recorded in a singular health system.
+OHDSI was developed to focus on leveraging observational health data stored within the OMOP CDM.
 
-OHDSI also develops open-source analytical and software tools, such as HADES and ATLAS, which features a comprehensive set of functions needed to conduct network studies from assessing demographic information to generating summary statistics and interactive dashboards.
-The best practices set forth by OHDSI has improved the quality and reliability of evidence in observational studies, and it improved network research study collaboration, supported large-scale analytics, and standardized data structures and medical vocabulary.
+At this time, OHS prioritizes creation and maintenance of tools that support analysis of "OMOP-ified"\footnotemark data (that is, data which is formatted in the OMOP CDM), in much the same manner that OHDSI does.
 
-As an example, OHDSI's HADES (Health Analytics Data-To-Evidence Suite) has been in existence for over $10$ years and has several rich packages.
+\footnotetext{I am unsure of where this notion came from. I believe it may have originally been coined by my former mentor and OHDSI founder, Dr. Jon Duke. 
+The provenance of related terms like "OMOP-ification" (the process of converting a database to the OMOP CDM) is even murkier.}
+
+OHDSI supports the development of open-source software tools in a centralized ecosystem called HADES (Health Analytics Data-to-Evidence Suite) for particularly managing and working with OMOP CDM formatted data.
+The HADES ecosystem has been in existence for $\approx 10$ years and has given rise to improvements in the quality and reliability of observational health research as well as motivating novel collaboration mechanisms.
+
+In particular, one collaboration mechanism that OHDSI has been able to operationalize meaningfully while working with observational health data is the notion of "network studies".
+These studies, are designed to be transparent and encompasses research endeavors being explored by members of the OHDSI comprehensive of relevant documentation, analysis code, results, and study protocols describing the study's scope and objectives.
+One core innovation brought about by OHDSI in the space of network science is how, given the HADES tools and procedures established by OHDSI, a research group can develop their analysis on their own OMOP CDM instance and then bundle together their study.
+Once this packaging is done, they can pass along the study package to another site to collaborate on the same analysis but on different data assets.
+Data confidentiality is preserved in this approach as data from individual participating sites remains exclusive to each site, with only aggregate results shared across a given collaborator network network.
+As such, no risk for re-identification comes about through this network collaboration and has ushered in the possibility for similar research approaches to be taken across the globe.
+\cite{ohdsi2019book}
+
 One of the premier tools within HADES is the ATLAS Shiny application which was developed to "support the design and execution of observational analyses to generate real world evidence from patient level observational data."
@@
 
-This section guides readers through common steps part of observational health research and to what capacity the Julia ecosystem supports these tasks.
-Here are the steps covered
+DONE:
+
+Achieving the notion of a network study in an OHDSI-sense is a highly valuable goal to multiple communities.
+JuliaHealth now has capacity to do just the same and can harness the best of OHDSI and the best of Julia to achieve novel observational health workflows.
+The JuliaHealth observational health subecosystem (referred to as OHS for the remainder of this paper) supports a variety of use cases designed to support and enable investigations into observational health data.
+This paper summarizes some of the current capacity OHS possesses to conduct such endeavors.
+In particular, the following supported areas will be discussed:
 
@@
 HADES is a collection of approximately 20 packages, developed by the OHDSI community to interact directly with the OMOP CDM to support large-scale analytics.
-The goals of HADES is much the same as the JuliaHealth observational health subecosystem such as supporting cohort construction, population-level estimation, patient-level prediction, and calculating other sorts of relevant research artifacts.
+The goals of HADES is much the same as the JuliaHealth OHS such as supporting cohort construction, population-level estimation, patient-level prediction, and calculating other sorts of relevant research artifacts.
 
@@
 For this paper, we used two datasets: a synthetic patient dataset called \textit{Eunomia} and the MIMIC III patient dataset which had been converted into the OMOP CDM to illustrate the workflow provided by the Observational Health sub-ecosystem.
-Reproducing the workflow using the \textit{Eunomia} dataset is made readily available at the repository here:
+Reproducing the workflow using the \textit{Eunomia} dataset is made readily available at the repository here: \cite{schuemieEunomia2023}
 
@@
 
-One crucial aspect of the observational health sub-ecosystem is that it is unavoidable to work with databases that contain patient health information.
+One crucial aspect of the OHS is that it is unavoidable to work with databases that contain patient health information.
 For better or worse, the tool that is used to best interact with these databases are SQL dialects that loosely follow consistently the ANSI SQL guidance.
@@
 Instead, what was realized instead is that what should be further investigated is novel compositions that can be viewed through the lens of JuliaHealth.
-As shown in this workflow, although there were some packages that were made specifically for JuliaHealth, other packages that were not JuliaHealth specific were used alongside of it that composed flawlessly with the rest of the JuliaHealth observational health subecosystem.
+As shown in this workflow, although there were some packages that were made specifically for JuliaHealth, other packages that were not JuliaHealth specific were used alongside of it that composed flawlessly with the rest of the JuliaHealth OHS.
 
@@
 
-Additionally, due to the composable and modular nature of the observational health sub-ecosystem, research analyses and visualizations can be decoupled and interchanged with one another.
+Additionally, due to the composable and modular nature of the OHS, research analyses and visualizations can be decoupled and interchanged with one another.
 Whereas in other ecosystems that make technologies for tools such as dashboards tightly coupled to analytics regimens, the subecosystem is mature enough to enable new users and developers to create analytics tools that can be separately woven into more graphical user interfaces.
@@
 
-In conclusion, this paper has demonstrated the capabilities of the JuliaHealth observational health sub-ecosystem and its potential to further investigations in observational health research.
+In conclusion, this paper has demonstrated the capabilities of the JuliaHealth OHS and its potential to further investigations in observational health research.
 Although it is an ecosystem that is relatively new, its robustness allows for novel investigations to be pursued within this domain of research. 

### 2023-12-05 paper.tex
Fixed TODOs and code listings

@@
 \documentclass{juliacon}
+
+\usepackage{listings}
+
 \setcounter{page}{1}
@@
 
-DONE:
+% DONE:
 JuliaHealth is an open-source collective dedicated to bringing the Julia programming language to various avenues of health research domains such as observational health, medical imaging, standards interoperability, etc.
@@
 
-TODO: Cite FunOHDSI
-TODO: Cite OHDSICohortExpressions
-TODO: Cite OHDSI Symposium poster on JuliaHealth
-TODO: Register OHDSI Symposium poster to Zenodo
-TODO: Cite JuliaCon 2023 presentation
-TODO: Register JuliaCon 2023 presentation to Zenodo
+% TODO: Cite FunOHDSI
+% TODO: Cite OHDSICohortExpressions
+% 
+% TODO: Cite OHDSI Symposium poster on JuliaHealth
+% TODO: Register OHDSI Symposium poster to Zenodo
+% TODO: Cite JuliaCon 2023 presentation
+% TODO: Register JuliaCon 2023 presentation to Zenodo
 
@@
 
-DONE:
+% DONE:
 
@@
 
-TODO: Add small example showing iterative process 
-TODO: Reference capr somehow
+% TODO: Mention capr here somehow
 
@@
 
-TODO: Add database connector example
-TODO: Reference database connector example
-TODO: Mention package extensions utility
+% TODO: Add database connector example
+% TODO: Reference database connector example
 
@@
 
-TODO: Add tutorial to separate repository
-TODO: Create tutorial about Eunomia
-TODO: Reference Eunomia
+% TODO: Add tutorial to separate repository
 
@@
 
-\begin{verbatim}
-  import HealthSampleData: Eunomia
+\begin{lstlisting}[language=Julia,numbers=none]
+import DBConnector: 
+  DBConnection
+import HealthSampleData: 
+  Eunomia
 import OMOPCDMCohortCreator as OCC
 
-  eunomia = Eunomia()
-  conn = DB(eunomia)
+conn = DBConnection("sqlite", Eunomia())
 
@@
 occ.GenerateTables(conn)
-\end{verbatim}
-
-TODO: Replace with DBConnector syntax
+\end{lstlisting}
 
@@
 
-\begin{verbatim}
+\begin{lstlisting}[language=Julia,numbers=none]
 using JSON3 
@@
 
-  model = Model(cdm_version=v"5.3.1", cdm_schema="main",
-                       vocabulary_schema="main", results_schema="main",
-                       target_schema="main", target_table="cohort");
+model = Model(cdm_version = v"5.3.1", 
+              cdm_schema = "main",
+              vocabulary_schema = "main", 
+              results_schema = "main",
+              target_schema = "main", 
+              target_table = "cohort");
 
-  sql = translate(cohort_expression, dialect=:sqlite, model=model, cohort_definition_id=1);
+sql = translate(cohort_expression, 
+                dialect = :sqlite, 
+                model = model, 
+                cohort_definition_id = 1);
 
-  [DBI.execute(conn, sub_query) for sub_query in split(sql, ";")[1:end-1]]
-\end{verbatim}
+for query in split(sql, ";")[1:end-1]
+  DBI.execute(conn, sub_query)
+end
+\end{lstlisting}
 
@@
 
-TODO: Clean up the code a little nicer
-TODO: Make sure the JSON3 reading of a cohort works properly
-
 \subsection{Characterizing Patient Populations}
 
-\begin{verbatim}
-  patients = DBI.execute(conn, "SELECT subject_id AS person_id FROM cohort;") |> DF.DataFrame
-\end{verbatim}
+\begin{lstlisting}[language=Julia,numbers=none]
+C = GetCohortSubjects(1, conn)
+rename!(C, :subject_id => :person_id)
 
-TODO: Replace this line from Jay's PR for getting cohorts
+C_race = occ.GetPatientRace(C.person_id, conn)
+C_gender = occ.GetPatientGender(C.person_id, conn)
 
-\begin{verbatim}
-  patients_race = occ.GetPatientRace(patients.person_id, conn)
-  patients_gender = occ.GetPatientGender(patients.person_id, conn)
-
-  patients_age_group = occ.GetPatientAgeGroup(
-    patients.person_id, 
+C_age_group = occ.GetPatientAgeGroup(
+  C.person_id, 
   conn; 
-    age_groupings = [[age, age + 4] for age in 0:5:119]
 )
-\end{verbatim}
+\end{lstlisting}
 
@@
 
-\begin{verbatim}
+\begin{lstlisting}[language=Julia,numbers=none]
 import DataFrames as DF
 
-  patients_characterized = DF.outerjoin(patients_race,
-                                              patients_gender,
-                                              patients_age_group;
+C_features = DF.outerjoin(C_race,
+                          C_gender,
+                          C_age_group;
                           on = :person_id, 
                           matchmissing = :equal)
-  patients_characterized = patients_characterized[:, DF.Not(:person_id)]
-  patient_groups = DF.groupby(patients_characterized, [:race_concept_id, :gender_concept_id, :age_group])
-  patient_groups = DF.combine(patient_groups, DF.nrow => :count)
-\end{verbatim}
+
+C_features = C_features[:, DF.Not(:person_id)]
+
+C_groups = DF.groupby(C_features, 
+                      [
+                        :race_concept_id, 
+                        :gender_concept_id, 
+                        :age_group
+                      ])
+
+C_groups = DF.combine(C_groups, DF.nrow => :count)
+\end{lstlisting}
 
@@
     \hline
-        Race & Gender & Age Group & Cohort Count & Total Count & Prev. (\%) \\ \hline
-        Black or African American & Female & 60 - 64 & 13 & 21 & 61.90 \\ \hline
-        White & Male & 70 - 74 & 68 & 96 & 70.83 \\ \hline
-        White & Male & 45 - 49 & 47 & 74 & 63.51 \\ \hline
-        White & Female & 50 - 54 & 59 & 111 & 53.15 \\ \hline
-        Black or African American & Female & 55 - 59 & 7 & 19 & 36.84 \\ \hline
-        White & Male & 75 - 79 & 45 & 56 & 80.35 \\ \hline
-        White & Female & 55 - 59 & 62 & 97 & 63.91 \\ \hline
-        White & Female & 45 - 49 & 41 & 92 & 44.56 \\ \hline
-        Asian & Female & 80 - 84 & 4 & 5 & 80.0  \hline
-        Asian & Male & 80 - 84 & 6 & 10 & 60.00 \\ \hline
-        Black or African American & Female & 45 - 49 & 10 & 23 & 43.47 \\ \hline
-        No matching concept & Female & 60 - 64 & 29 & 39 & 74.35 \\ \hline
-        Black or African American & Male & 65 - 69 & 12 & 18 & 66.66 \\ \hline
+        Race & Gender & Age Group & Cohort & Total & Prev. (\%) \\ \hline
+        B/AA & F & 60 - 64 & 13 & 21 & 61.90 \\ \hline
+        W & F & 50 - 54 & 59 & 111 & 53.15 \\ \hline
+        W & F & 55 - 59 & 62 & 97 & 63.91 \\ \hline
+        A & M & 80 - 84 & 6 & 10 & 60.00 \\ \hline
+        UNK & F & 60 - 64 & 29 & 39 & 74.35 \\ \hline
+        B/AA & M & 65 - 69 & 12 & 18 & 66.66 \\ \hline
     \end{tabular}
@@
 
-TODO: Check that I equated currying correctly here.
-
-\begin{verbatim}
+\begin{lstlisting}[language=Julia,numbers=none]
 import Base:
@@
 
-CGetPatientRace = Base.Fix2(occ.GetPatientRace, conn)
-CGetPatientGender = Base.Fix2(occ.GetPatientGender, conn)
-CGetPatientAgeGroup = Base.Fix2(occ.GetPatientAgeGroup, conn)
-\end{verbatim}
+import OMOPCDMCohortCreator:
+  GetPatientRace,
+  GetPatientGender,
+  GetPatientAgeGroup
+
+CGetPatientRace = Fix2(GetPatientRace, c)
+CGetPatientGender = Fix2(GetPatientGender, c)
+CGetPatientAgeGroup = Fix2(GetPatientAgeGroup, c)
+\end{lstlisting}
 
@@
 
-\begin{verbatim}
+\begin{lstlisting}[language=Julia,numbers=none]
 import Chain:
@@
 end
-\end{verbatim}
+\end{lstlisting}
 
-\begin{verbatim}
+\begin{lstlisting}[language=Julia,numbers=none]
 CharacterizePatients(patients) = @chain patients begin
@@
 end
-\end{verbatim}
+\end{lstlisting}
 
@@
 
-\begin{verbatim}
+\begin{lstlisting}[language=Julia,numbers=none]
 RunStudy(patients) = @chain patients begin
@@
 
-\begin{verbatim}
+\begin{lstlisting}[language=Julia,numbers=none]
 import Distributed:
@@
 vcat(result...) |> audit_patient_groups
-\end{verbatim}
+\end{lstlisting}
 
@@
 
-\begin{verbatim}
+\begin{lstlisting}[language=Julia,numbers=none]
   GetPatientAgeGroup([1]; age_groupings = [[0, 40], [41, 80]])
-\end{verbatim}
+\end{lstlisting}
 
-\begin{verbatim}
+\begin{lstlisting}[language=Julia,numbers=none]
   SELECT
@@
   ) AS "PERSON_2"
-\end{verbatim}
+\end{lstlisting}
 

### 2023-12-08 paper.tex
Draft version 0.2.0
Started review and adjusted code listings

@@
+%! TeX program = lualatex
+
 \documentclass{juliacon}
@@
 \usepackage{listings}
+
+\usepackage{fontspec}
+\usepackage{minted}
+
+\newfontfamily \JuliaMono {JuliaMono.ttf}[
+    Path      = ./,
+    Extension = .ttf
+    ]
+
+\newfontface \JuliaMono{JuliaMono}
+
+\setmonofont{JuliaMono}[
+    Contextuals=Alternate
+]
 
@@
 Observational health continues to be a growing field in health informatics research as electronic health records (EHR), patient medical claims, and other ancilliary patient data source become more readily computable and accessible to researchers.
+% TODO: 
 JuliaHealth is poised as an ecosystem to innovate within this area of research by bringing highly performant analytics approaches, composable solutions, and interoperable software that leverages prior state of the art.
+% TODO: 
 This paper will discuss the state of the art observational health research tools within the JuliaHealth ecosystem and how JuliaHealth is prepared to further research goals within this domain.
@@
 
-Over the past $10$ years, there have been several innovations within the observational health research.
+% EXPLAIN: What is observational health research?
+% EXPLAIN: What sort of innovations?
+Over the past $10$ years, there have been several innovations within the observational health research community.
+
+% EXPLAIN: Why is this a crucial innovation?
+% EXPLAIN: What is meant by a transferable analysis
 Arguably, the most crucial innovation that has emerged is the Observational Medical Outcomes Partnership Common Data Model (OMOP CDM) which was designed specifically to standardize observational health data to enable transferable analyses.
+% TODO: Add citation
+% NOTE: Why is mapping relevant here?
 The importance of this model has become apparent within open science organizations like the OHDSI (Observational Health Data Science and Informatics) network where members across $100+$ countries with access to various health systems have been able to successfully map their patient data to the OMOP CDM.
+% TODO: Add citation about successful collaborations
+% EXPLAIN: What is a network study
+% TODO: Cite the quote "write once, run everywhere"
 This has led to successful collaboration relationships across countries where one research group can develop a network study and then disseminate the study across collaborators leading to the idea that you can "write once, run everywhere."
 
+As the OMOP CDM has become a cornerstone in running these widely distributed network studies, special attention to developing tools around handling OMOP CDM data is necessary to meaningfully contribute within the observational health research space.
 The OMOP CDM itself does not require a specific technology to work with the data stored in this standard.
 It features a person-centric design where each domain records personal identity while prioritizing data protection through the limiting of information that could endanger patient anonymity.
-As the OMOP CDM has become a cornerstone in running these widely distributed network studies, special attention to developing tools around handling OMOP CDM data is necessary to meaningfully contribute within the observational health research space.
 
@@
 
-% DONE:
-JuliaHealth is an open-source collective dedicated to bringing the Julia programming language to various avenues of health research domains such as observational health, medical imaging, standards interoperability, etc.
-It leverages the strengths of the Julia language to introduce new approaches to these spaces, augment and improve existing workflows, and provide a central community structure for health researchers both part of and entering into the Julia ecosystem.
+% DONE
+JuliaHealth is an open-science collective dedicated to bringing the Julia programming language to various avenues of health research domains including, but not limited to, observational health, medical imaging, standards interoperability and more.
+It leverages the strengths of the Julia language to introduce new methods into these spaces, augment existing workflows, and provide a community structure for health researchers within the global Julia ecosystem.
 Eventually, JuliaHealth will host a variety of subecosystems and associated resources that will enable novel research to be conducted within Julia.
@@
 
-JuliaHealth itself is a loose collective of various research software engineers, health researchers, and hobbyist programmers that had its origins in $\approx 2020$.
-Being that JuliaHealth is a federated network of contributors, varying aspects of the ecosystem have developed at differing rates as a function of multiple different factors such as developer time, workplace incentives, and more.
-One area of the subecosystems within JuliaHealth that has matured significantly over the years has been the observational health subecosystem (referred to as OHS for the remainder of this paper).
+As JuliaHealth is an ad hoc network of contributors, varying aspects of the ecosystem have developed at differing rates as a function of factors such as developer time or workplace incentives.
+One area of the subecosystems within JuliaHealth that has matured significantly over the years has been the JuliaHealth observational health subecosystem (referred to as OHS or the "subecosystem" for the remainder of this paper).
 
-Although difficult to date, it can said that OHS had its inception in work done by Evans and Simonov with efforts such as \textit{FunOHDSI.jl}, \textit{OHDSICohortExpressions}, and \textit{FunSQL.jl} around $2021$  \cite{kirill_simonov_2023_7705325}
+Although difficult to date, it can said that OHS had its inception in work done by Evans and Simonov with efforts such as \textit{FunOHDSI.jl}, \textit{OHDSICohortExpressions.jl}, and \textit{FunSQL.jl} around $2021$ \cite{kirill_simonov_2023_7705325}
 Since then, Zelko et.
@@
 have spearheaded most recent developments within OHS via the creation of a variety of tools and demonstrated their effectiveness in using these tools for novel observational health research endeavors \cite{zelko2022pilot}
-Through such efforts, OHS is becoming further capable of executing novel observational health research although, understandably, not as mature as similar ecosystems.
+Through such efforts, OHS is becoming further capable of executing novel observational health research.
 
@@
 
-In particular, the strongest influence and inspiration to OHS is the OHDSI open-science network.
-OHDSI (Observational Health Data Sciences and Informatics) is a global, collaborative, interdisciplinary, and open network of researchers, healthcare professionals, students, patients, and other entities committed to large-scale health data analytics and generating reliable evidence, with the goal of promoting better health decisions and better care.
+Driving development within the subecosystem is the influence the OHDSI open-science network.
+As mentioned previously, OHDSI is a global, collaborative, interdisciplinary, and open network of researchers, healthcare professionals, students, patients, and other entities committed to large-scale health data analytics and generating reliable evidence, with the goal of promoting better health decisions and better care.
 Since current observational health data sources are aggregated at different levels and cannot be generalized to the overall population, the data is often incomplete, inaccurate, and prone to issues of bias, error handling, and multidimensional complexity.
@@
 
-\begin{lstlisting}[language=Julia,numbers=none]
+\begin{minted}[breaklines,escapeinside=||,mathescape=true, numbersep=3pt, gobble=2, frame=lines, fontsize=\scriptsize, framesep=2mm]{julia}
 using JSON3 
@@
 end
-\end{lstlisting}
+\end{minted}
 

### 2023-12-10 paper.tex
Working on making major revisions

@@
 \documentclass{juliacon}
-
-\usepackage{listings}
 
@@
 As the OMOP CDM has become a cornerstone in running these widely distributed network studies, special attention to developing tools around handling OMOP CDM data is necessary to meaningfully contribute within the observational health research space.
+% TODO: Move this to the section: Status of Conducting Observational Health Research in Julia
 The OMOP CDM itself does not require a specific technology to work with the data stored in this standard.
@@
 
-% DONE
 JuliaHealth is an open-science collective dedicated to bringing the Julia programming language to various avenues of health research domains including, but not limited to, observational health, medical imaging, standards interoperability and more.
@@
 \section{Describing Common Observational Health Research Tasks within the Julia Ecosystem}
-
-% DONE:
 
@@
 
-\begin{lstlisting}[language=Julia,numbers=none]
+\begin{minted}[breaklines,escapeinside=||,mathescape=true, numbersep=3pt, gobble=2, frame=lines, fontsize=\scriptsize, framesep=2mm]{julia}
 import DBConnector: 
@@
 occ.GenerateTables(conn)
-\end{lstlisting}
+\end{minted}
 
@@
 for query in split(sql, ";")[1:end-1]
-  DBI.execute(conn, sub_query)
+  DBI.execute(conn, query)
 end
@@
 
-\begin{lstlisting}[language=Julia,numbers=none]
+\begin{minted}[breaklines,escapeinside=||,mathescape=true, numbersep=3pt, gobble=2, frame=lines, fontsize=\scriptsize, framesep=2mm]{julia}
 C = GetCohortSubjects(1, conn)
@@
 )
-\end{lstlisting}
+\end{minted}
 
@@
 
-\begin{lstlisting}[language=Julia,numbers=none]
+\begin{minted}[breaklines,escapeinside=||,mathescape=true, numbersep=3pt, gobble=2, frame=lines, fontsize=\scriptsize, framesep=2mm]{julia}
 import DataFrames as DF
@@
 C_groups = DF.combine(C_groups, DF.nrow => :count)
-\end{lstlisting}
+\end{minted}
 
@@
 
-\begin{lstlisting}[language=Julia,numbers=none]
+\begin{minted}[breaklines,escapeinside=||,mathescape=true, numbersep=3pt, gobble=2, frame=lines, fontsize=\scriptsize, framesep=2mm]{julia}
 import Base:
@@
 CGetPatientAgeGroup = Fix2(GetPatientAgeGroup, c)
-\end{lstlisting}
+\end{minted}
 
@@
 
-\begin{lstlisting}[language=Julia,numbers=none]
+\begin{minted}[breaklines,escapeinside=||,mathescape=true, numbersep=3pt, gobble=2, frame=lines, fontsize=\scriptsize, framesep=2mm]{julia}
 import Chain:
@@
 end
-\end{lstlisting}
+\end{minted}
 
-\begin{lstlisting}[language=Julia,numbers=none]
+\begin{minted}[breaklines,escapeinside=||,mathescape=true, numbersep=3pt, gobble=2, frame=lines, fontsize=\scriptsize, framesep=2mm]{julia}
 CharacterizePatients(patients) = @chain patients begin
@@
 
-\begin{lstlisting}[language=Julia,numbers=none]
+\begin{minted}[breaklines,escapeinside=||,mathescape=true, numbersep=3pt, gobble=2, frame=lines, fontsize=\scriptsize, framesep=2mm]{julia}
 RunStudy(patients) = @chain patients begin
@@
 
-\begin{lstlisting}[language=Julia,numbers=none]
+\begin{minted}[breaklines,escapeinside=||,mathescape=true, numbersep=3pt, gobble=2, frame=lines, fontsize=\scriptsize, framesep=2mm]{julia}
 import Distributed:
@@
 vcat(result...) |> audit_patient_groups
-\end{lstlisting}
+\end{minted}
 
@@
 
-\begin{lstlisting}[language=Julia,numbers=none]
+\begin{minted}[breaklines,escapeinside=||,mathescape=true, numbersep=3pt, gobble=2, frame=lines, fontsize=\scriptsize, framesep=2mm]{julia}
   GetPatientAgeGroup([1]; age_groupings = [[0, 40], [41, 80]])
-\end{lstlisting}
+\end{minted}
 
-\begin{lstlisting}[language=Julia,numbers=none]
+\begin{minted}[breaklines,escapeinside=||,mathescape=true, numbersep=3pt, gobble=2, frame=lines, fontsize=\scriptsize, framesep=2mm]{julia}
   SELECT
@@
   ) AS "PERSON_2"
-\end{lstlisting}
+\end{minted}
 

### 2023-12-12 paper.tex
Continuing massive reviews of the JuliaHealth paper

@@
 % TODO: Add citation about successful collaborations
-% EXPLAIN: What is a network study
+% EXPLAIN: What is a network study?
 % TODO: Cite the quote "write once, run everywhere"
 This has led to successful collaboration relationships across countries where one research group can develop a network study and then disseminate the study across collaborators leading to the idea that you can "write once, run everywhere."
+
+\subsection{Background}
 
@@
 It features a person-centric design where each domain records personal identity while prioritizing data protection through the limiting of information that could endanger patient anonymity.
-
-\subsection{What Is JuliaHealth?}
-
-JuliaHealth is an open-science collective dedicated to bringing the Julia programming language to various avenues of health research domains including, but not limited to, observational health, medical imaging, standards interoperability and more.
-It leverages the strengths of the Julia language to introduce new methods into these spaces, augment existing workflows, and provide a community structure for health researchers within the global Julia ecosystem.
-Eventually, JuliaHealth will host a variety of subecosystems and associated resources that will enable novel research to be conducted within Julia.
-
-\subsection{Status of Conducting Observational Health Research in Julia}
-
-As JuliaHealth is an ad hoc network of contributors, varying aspects of the ecosystem have developed at differing rates as a function of factors such as developer time or workplace incentives.
-One area of the subecosystems within JuliaHealth that has matured significantly over the years has been the JuliaHealth observational health subecosystem (referred to as OHS or the "subecosystem" for the remainder of this paper).
-
-Although difficult to date, it can said that OHS had its inception in work done by Evans and Simonov with efforts such as \textit{FunOHDSI.jl}, \textit{OHDSICohortExpressions.jl}, and \textit{FunSQL.jl} around $2021$ \cite{kirill_simonov_2023_7705325}
-Since then, Zelko et.
-al.
-have spearheaded most recent developments within OHS via the creation of a variety of tools and demonstrated their effectiveness in using these tools for novel observational health research endeavors \cite{zelko2022pilot}
-Through such efforts, OHS is becoming further capable of executing novel observational health research.
-
-% TODO: Cite FunOHDSI
-% TODO: Cite OHDSICohortExpressions
-% 
-% TODO: Cite OHDSI Symposium poster on JuliaHealth
-% TODO: Register OHDSI Symposium poster to Zenodo
-% TODO: Cite JuliaCon 2023 presentation
-% TODO: Register JuliaCon 2023 presentation to Zenodo
 
@@
 \cite{ohdsi2019book}
+
+\subsection{What Is JuliaHealth?}
+
+% TODO: Cite Discourse post on start of JuliaHealth
+JuliaHealth is an open-science collective dedicated to bringing the Julia programming language to various avenues of health research domains including, but not limited to, observational health, medical imaging, standards interoperability and more.
+It leverages the strengths of the Julia language to introduce new methods into these spaces, augment existing workflows, and provide a community structure for health researchers within the global Julia ecosystem.
+% EXPLAIN: What is a subecosystem?
+% EXPLAIN: How does having more subecosystems enable novel research to be done?
+Eventually, JuliaHealth will host a variety of subecosystems and associated resources that will enable novel research to be conducted within Julia.
+
+\subsection{Status of Conducting Observational Health Research in Julia}
+
+% TODO: Move this to What Is JuliaHealth section
+As JuliaHealth is an ad hoc network of contributors, varying aspects of the ecosystem have developed at differing rates as a function of factors such as developer time or workplace incentives.
+
+One area of the subecosystems within JuliaHealth that has matured significantly over the years has been the JuliaHealth observational health subecosystem (OHS).
+It can be said that OHS had its inception in work done by Evans and Simonov with efforts such as \textit{FunOHDSI.jl}, \textit{OHDSICohortExpressions.jl}, and \textit{FunSQL.jl} around $2021$ \cite{kirill_simonov_2023_7705325}
+% TODO: Cite FunOHDSI
+% TODO: Cite OHDSICohortExpressions
+% TODO: Cite OHDSI Symposium poster on JuliaHealth
+% TODO: Register OHDSI Symposium poster to Zenodo
+% TODO: Cite JuliaCon 2023 presentation
+% TODO: Register JuliaCon 2023 presentation to Zenodo
+Since then, Zelko et.
+al.
+have spearheaded most recent developments within OHS via the creation of a variety of tools and demonstrated their effectiveness in using these tools for novel observational health research endeavors \cite{zelko2022pilot}
+% EXPLAIN: What does this mean?
+Through such efforts, OHS is becoming further capable of executing novel observational health research.
 

### 2023-12-14 paper.tex
Working on TODOs
Added additional reviews in background section

@@
 
+TODO: Review the abstract after reviewing the paper fully
 Observational health continues to be a growing field in health informatics research as electronic health records (EHR), patient medical claims, and other ancilliary patient data source become more readily computable and accessible to researchers.
-% TODO: 
 JuliaHealth is poised as an ecosystem to innovate within this area of research by bringing highly performant analytics approaches, composable solutions, and interoperable software that leverages prior state of the art.
-% TODO: 
 This paper will discuss the state of the art observational health research tools within the JuliaHealth ecosystem and how JuliaHealth is prepared to further research goals within this domain.
@@
 % EXPLAIN: What is meant by a transferable analysis
+% TODO: Cite: J.
+M.
+Overhage, P.
+B.
+Ryan, C.
+G.
+Reich, A.
+G.
+Hartzema, and P.
+E.
+Stang, “Validation of a common data model for active safety surveillance research,” Journal of the American Medical Informatics Association, vol.
+19, no.
+1, pp.
+54–60, 2012.
 Arguably, the most crucial innovation that has emerged is the Observational Medical Outcomes Partnership Common Data Model (OMOP CDM) which was designed specifically to standardize observational health data to enable transferable analyses.
@@
 % TODO: Cite the quote "write once, run everywhere" 
+% http://www.sun.com/smi/Press/sunflash/1996-01/sunflash.960123.10561.xml} 
 This has led to successful collaboration relationships across countries where one research group can develop a network study and then disseminate the study across collaborators leading to the idea that you can "write once, run everywhere."
@@
 
+% TODO: Reframe this paragraph to make it like a better background introduction.
 As the OMOP CDM has become a cornerstone in running these widely distributed network studies, special attention to developing tools around handling OMOP CDM data is necessary to meaningfully contribute within the observational health research space.
-% TODO: Move this to the section: Status of Conducting Observational Health Research in Julia
 The OMOP CDM itself does not require a specific technology to work with the data stored in this standard.
@@
 
+% TODO: Reframe this to suggest the more general impact that OHDSI has had within the observational health research space.
 Driving development within the subecosystem is the influence the OHDSI open-science network.
 As mentioned previously, OHDSI is a global, collaborative, interdisciplinary, and open network of researchers, healthcare professionals, students, patients, and other entities committed to large-scale health data analytics and generating reliable evidence, with the goal of promoting better health decisions and better care.
-Since current observational health data sources are aggregated at different levels and cannot be generalized to the overall population, the data is often incomplete, inaccurate, and prone to issues of bias, error handling, and multidimensional complexity.
-The goal of OHDSI is to produce reproducible, robust, and reliable evidence by developing standardized data formats and research protocol designs to enable large-scale collaborative research and improve the transparency and scalability of observational health research.
-
 OHDSI had its inception in the development of the Observational Medical Outcomes Partnership (OMOP) that was created in 2009 by the US Food and Drug Administration to address concerns related to data type, study design, and data privacy concerns of "Real World Data" \cite{ohdsi2019book} \cite{FDARealWorldEvidence}
@@
 OHDSI was developed to focus on leveraging observational health data stored within the OMOP CDM.
-
-At this time, OHS prioritizes creation and maintenance of tools that support analysis of "OMOP-ified"\footnotemark data (that is, data which is formatted in the OMOP CDM), in much the same manner that OHDSI does.
-
-\footnotetext{I am unsure of where this notion came from. I believe it may have originally been coined by my former mentor and OHDSI founder, Dr. Jon Duke. 
-The provenance of related terms like "OMOP-ification" (the process of converting a database to the OMOP CDM) is even murkier.}
+Since current observational health data sources are aggregated at different levels and cannot be generalized to the overall population, the data is often incomplete, inaccurate, and prone to issues of bias, error handling, and multidimensional complexity.
+The goal of OHDSI is to produce reproducible, robust, and reliable evidence by developing standardized data formats and research protocol designs to enable large-scale collaborative research and improve the transparency and scalability of observational health research.
 
@@
 
+% TODO: Use this content to address what the notion of "network studies" are; move this to the "Introduction" section
 In particular, one collaboration mechanism that OHDSI has been able to operationalize meaningfully while working with observational health data is the notion of "network studies".
@@
 \cite{ohdsi2019book}
+
+At this time, OHS prioritizes creation and maintenance of tools that support analysis of "OMOP-ified"\footnotemark data (that is, data which is formatted in the OMOP CDM), in much the same manner that OHDSI does.
+
+\footnotetext{I am unsure of where this notion came from. I believe it may have originally been coined by my former mentor and OHDSI founder, Dr. Jon Duke. 
+The provenance of related terms like "OMOP-ification" (the process of converting a database to the OMOP CDM) is even murkier.}
 

### 2023-12-14 paper.tex
Major refactor of introduction section

@@
 
-\subsection{Background}
-
-% TODO: Reframe this paragraph to make it like a better background introduction.
-As the OMOP CDM has become a cornerstone in running these widely distributed network studies, special attention to developing tools around handling OMOP CDM data is necessary to meaningfully contribute within the observational health research space.
-The OMOP CDM itself does not require a specific technology to work with the data stored in this standard.
-It features a person-centric design where each domain records personal identity while prioritizing data protection through the limiting of information that could endanger patient anonymity.
-
-% TODO: Reframe this to suggest the more general impact that OHDSI has had within the observational health research space.
-Driving development within the subecosystem is the influence the OHDSI open-science network.
-As mentioned previously, OHDSI is a global, collaborative, interdisciplinary, and open network of researchers, healthcare professionals, students, patients, and other entities committed to large-scale health data analytics and generating reliable evidence, with the goal of promoting better health decisions and better care.
-OHDSI had its inception in the development of the Observational Medical Outcomes Partnership (OMOP) that was created in 2009 by the US Food and Drug Administration to address concerns related to data type, study design, and data privacy concerns of "Real World Data" \cite{ohdsi2019book} \cite{FDARealWorldEvidence}
-What could be considered the magnum opus of the OMOP group was the development of what is called the OMOP Common Data Model (OMOP CDM) to house and ingest this sort of data in a consistent format that has been optimized to assist in extracting meaningful population health information.
-In brief, the OMOP Common Data Model (CDM) is a "person-centered" model that has standardized common events in a given patient encounter with database tables that represent the unique structures found in patient events.
-This creates a longitudinal view of all healthcare-related events per person recorded in a singular health system.
-OHDSI was developed to focus on leveraging observational health data stored within the OMOP CDM.
-Since current observational health data sources are aggregated at different levels and cannot be generalized to the overall population, the data is often incomplete, inaccurate, and prone to issues of bias, error handling, and multidimensional complexity.
-The goal of OHDSI is to produce reproducible, robust, and reliable evidence by developing standardized data formats and research protocol designs to enable large-scale collaborative research and improve the transparency and scalability of observational health research.
-
-OHDSI supports the development of open-source software tools in a centralized ecosystem called HADES (Health Analytics Data-to-Evidence Suite) for particularly managing and working with OMOP CDM formatted data.
-The HADES ecosystem has been in existence for $\approx 10$ years and has given rise to improvements in the quality and reliability of observational health research as well as motivating novel collaboration mechanisms.
-
-% TODO: Use this content to address what the notion of "network studies" are; move this to the "Introduction" section
+% TODO: Use this to address the notion of what "network studies" are
 In particular, one collaboration mechanism that OHDSI has been able to operationalize meaningfully while working with observational health data is the notion of "network studies".
@@
 
+\subsection{Background}
+
+\subsubsection{Core Technologies within Observational Health Research}
+
+% TODO: Reframe this paragraph to make it like a better background introduction.
+What could be considered the magnum opus of the OMOP group was the development of what is called the OMOP Common Data Model (OMOP CDM) to house and ingest this sort of data in a consistent format that has been optimized to assist in extracting meaningful population health information.
+In brief, the OMOP Common Data Model (CDM) is a "person-centered" model that has standardized common events in a given patient encounter with database tables that represent the unique structures found in patient events.
+This creates a longitudinal view of all healthcare-related events per person recorded in a singular health system.
+As the OMOP CDM has become a cornerstone in running these widely distributed network studies, special attention to developing tools around handling OMOP CDM data is necessary to meaningfully contribute within the observational health research space.
+The OMOP CDM itself does not require a specific technology to work with the data stored in this standard.
+It features a person-centric design where each domain records personal identity while prioritizing data protection through the limiting of information that could endanger patient anonymity.
+
+\subsubsection{The Impact of the Observational Health Data Science and Informatics (OHDSI) Consortium}
+
+% TODO: Reframe this to suggest the more general impact that OHDSI has had within the observational health research space.
+Driving development within the subecosystem is the influence the OHDSI open-science network.
+As mentioned previously, OHDSI is a global, collaborative, interdisciplinary, and open network of researchers, healthcare professionals, students, patients, and other entities committed to large-scale health data analytics and generating reliable evidence, with the goal of promoting better health decisions and better care.
+OHDSI had its inception in the development of the Observational Medical Outcomes Partnership (OMOP) that was created in 2009 by the US Food and Drug Administration to address concerns related to data type, study design, and data privacy concerns of "Real World Data" \cite{ohdsi2019book} \cite{FDARealWorldEvidence}
+OHDSI was developed to focus on leveraging observational health data stored within the OMOP CDM.
+Since current observational health data sources are aggregated at different levels and cannot be generalized to the overall population, the data is often incomplete, inaccurate, and prone to issues of bias, error handling, and multidimensional complexity.
+The goal of OHDSI is to produce reproducible, robust, and reliable evidence by developing standardized data formats and research protocol designs to enable large-scale collaborative research and improve the transparency and scalability of observational health research.
+
+OHDSI supports the development of open-source software tools in a centralized ecosystem called HADES (Health Analytics Data-to-Evidence Suite) for particularly managing and working with OMOP CDM formatted data.
+The HADES ecosystem has been in existence for $\approx 10$ years and has given rise to improvements in the quality and reliability of observational health research as well as motivating novel collaboration mechanisms.
+
+\subsubsection{JuliaHealth's Role within Observational Health Research}
+
+As JuliaHealth is an ad hoc network of contributors, varying aspects of the ecosystem have developed at differing rates as a function of factors such as developer time or workplace incentives.
+
 At this time, OHS prioritizes creation and maintenance of tools that support analysis of "OMOP-ified"\footnotemark data (that is, data which is formatted in the OMOP CDM), in much the same manner that OHDSI does.
 
-\footnotetext{I am unsure of where this notion came from. I believe it may have originally been coined by my former mentor and OHDSI founder, Dr. Jon Duke. 
-The provenance of related terms like "OMOP-ification" (the process of converting a database to the OMOP CDM) is even murkier.}
-
-\subsection{What Is JuliaHealth?}
+\footnotetext{It is unclear where terms like "OMOP-ify" or "OMOP-ification" (the process of converting a data to the OMOP CDM) originated; it may have been coined by Dr. Jon Duke.}
 
@@
 Eventually, JuliaHealth will host a variety of subecosystems and associated resources that will enable novel research to be conducted within Julia.
-
-\subsection{Status of Conducting Observational Health Research in Julia}
-
-% TODO: Move this to What Is JuliaHealth section
-As JuliaHealth is an ad hoc network of contributors, varying aspects of the ecosystem have developed at differing rates as a function of factors such as developer time or workplace incentives.
 
@@
 Through such efforts, OHS is becoming further capable of executing novel observational health research.
-
-One of the premier tools within HADES is the ATLAS Shiny application which was developed to "support the design and execution of observational analyses to generate real world evidence from patient level observational data."
 
@@
 \subsection{Using Phenotype Definitions}
+
+One of the premier tools within HADES is the ATLAS Shiny application which was developed to "support the design and execution of observational analyses to generate real world evidence from patient level observational data."
 

### 2023-12-14 paper.tex
Added more todos and restructuring

@@
 Arguably, the most crucial innovation that has emerged is the Observational Medical Outcomes Partnership Common Data Model (OMOP CDM) which was designed specifically to standardize observational health data to enable transferable analyses.
-% TODO: Add citation
-% NOTE: Why is mapping relevant here?
+% TODO: Add citation from OHDSI website's collaborator statistics
+% EXPLAIN: Why is mapping relevant here?
 The importance of this model has become apparent within open science organizations like the OHDSI (Observational Health Data Science and Informatics) network where members across $100+$ countries with access to various health systems have been able to successfully map their patient data to the OMOP CDM.
@@
 
+\subsection{Core Technologies}
+
+Modern observational health research centers around a variety of core technologies.
+A few of these technologies highlighted in this paper are:
+
+\begin{enumerate}
+
+\item Data standards within observational health research
+\item Collaborative networks and studies using observational health data
+\item Research software to support the analysis of observational health data
+
+\end{enumerate}
+
+In particular, the data standard that will be most highlighted for the remainder of this paper is the Observational Medical Outcomes Partnership Common Data Model (OMOP CDM).
+Subsequent discussion will be focused on the opportunities that this specific data model has given rise to across observational health as well as research communities such as the Observational Health Data Sciences and Informatics (OHDSI) open science consortium.
+
+% TODO: This paragraph should be centered on the core technology of data standards -- in particular, the OMOP CDM
+% EXPLAIN: What is real world data? Use https://jacobzelko.com/10282021140730-real-world-evidence/
+OMOP was created in 2009 by the US Food and Drug Administration to address concerns related to data type, study design, and data privacy concerns of "Real World Data" \cite{ohdsi2019book} \cite{FDARealWorldEvidence}
+What could arguably be considered the longest standing achievement of OMOP was the development of the OMOP Common Data Model (OMOP CDM) to house and ingest real world data.
+% TODO: Combine the next two sentences together
+% EXPLAIN: What does "person-centered" mean?
+% EXPLAIN: What is a patient event? Could you give an example?
+in a consistent format that has been optimized to assist in extracting meaningful population health information.
+In brief, the OMOP Common Data Model (CDM) is a "person-centered" model that has standardized common events in a given patient encounter with database tables that represent the unique structures found in patient events.
+% TODO: Develop this fragment into a full sentence
+prioritizing data protection through the limiting of information that could endanger patient anonymity.
+% EXPLAIN: What is a health system? Could you give an example?
+This creates a longitudinal view of all healthcare-related events per person recorded in a singular health system.
+% TODO: Make this more coherent to the rest of the paragraph
+% TODO: Explicitly mention that although the model prescribes a database schema, it does not need a specific database type
+The OMOP CDM itself does not require a specific technology to work with the data stored in this standard.
+
+% TODO: This paragraph should be centered on the core technology of the concept of network studies
 % TODO: Use this to address the notion of what "network studies" are
+% TODO: Make this initially less OHDSI specific; pull from https://jacobzelko.com/05122023034151-network-macarthur
 In particular, one collaboration mechanism that OHDSI has been able to operationalize meaningfully while working with observational health data is the notion of "network studies".
-These studies, are designed to be transparent and encompasses research endeavors being explored by members of the OHDSI comprehensive of relevant documentation, analysis code, results, and study protocols describing the study's scope and objectives.
-One core innovation brought about by OHDSI in the space of network science is how, given the HADES tools and procedures established by OHDSI, a research group can develop their analysis on their own OMOP CDM instance and then bundle together their study.
-Once this packaging is done, they can pass along the study package to another site to collaborate on the same analysis but on different data assets.
-Data confidentiality is preserved in this approach as data from individual participating sites remains exclusive to each site, with only aggregate results shared across a given collaborator network network.
-As such, no risk for re-identification comes about through this network collaboration and has ushered in the possibility for similar research approaches to be taken across the globe.
-\cite{ohdsi2019book}
-
-\subsection{Background}
-
-\subsubsection{Core Technologies within Observational Health Research}
-
-% TODO: Reframe this paragraph to make it like a better background introduction.
-What could be considered the magnum opus of the OMOP group was the development of what is called the OMOP Common Data Model (OMOP CDM) to house and ingest this sort of data in a consistent format that has been optimized to assist in extracting meaningful population health information.
-In brief, the OMOP Common Data Model (CDM) is a "person-centered" model that has standardized common events in a given patient encounter with database tables that represent the unique structures found in patient events.
-This creates a longitudinal view of all healthcare-related events per person recorded in a singular health system.
-As the OMOP CDM has become a cornerstone in running these widely distributed network studies, special attention to developing tools around handling OMOP CDM data is necessary to meaningfully contribute within the observational health research space.
-The OMOP CDM itself does not require a specific technology to work with the data stored in this standard.
-It features a person-centric design where each domain records personal identity while prioritizing data protection through the limiting of information that could endanger patient anonymity.
-
-\subsubsection{The Impact of the Observational Health Data Science and Informatics (OHDSI) Consortium}
-
-% TODO: Reframe this to suggest the more general impact that OHDSI has had within the observational health research space.
-Driving development within the subecosystem is the influence the OHDSI open-science network.
-As mentioned previously, OHDSI is a global, collaborative, interdisciplinary, and open network of researchers, healthcare professionals, students, patients, and other entities committed to large-scale health data analytics and generating reliable evidence, with the goal of promoting better health decisions and better care.
+These studies, are designed to be transparent and encompasses research endeavors being explored by members of OHDSI comprehensive of relevant documentation, analysis code, results, and study protocols describing the study's scope and objectives.
 OHDSI had its inception in the development of the Observational Medical Outcomes Partnership (OMOP) that was created in 2009 by the US Food and Drug Administration to address concerns related to data type, study design, and data privacy concerns of "Real World Data" \cite{ohdsi2019book} \cite{FDARealWorldEvidence}
@@
 The goal of OHDSI is to produce reproducible, robust, and reliable evidence by developing standardized data formats and research protocol designs to enable large-scale collaborative research and improve the transparency and scalability of observational health research.
+% TODO: Carry HADES discussion into the paragraph talking about research tooling
+One core innovation brought about by OHDSI in the space of network science is how, given the HADES tools and procedures established by OHDSI, 
+% TODO: Make this more general such that it flows into the next paragraph on research tooling better
+a research group can develop their analysis on their own OMOP CDM instance and then bundle together their study.
+% EXPLAIN: What is a study package?
+% EXPLAIN: What data assets?
+Once this packaging is done, they can pass along the study package to another site to collaborate on the same analysis but on different data assets.
+% EXPLAIN: What is a collaborator network?
+Data confidentiality is preserved in this approach as data from individual participating sites remains exclusive to each site, with only aggregate results shared across a given collaborator network.
+As such, no risk for re-identification comes about through this network collaboration and has ushered in the possibility for similar research approaches to be taken across the globe.
+\cite{ohdsi2019book}
 
+% TODO: Develop this paragraph more about research software
+As the OMOP CDM has become a cornerstone in running these widely distributed network studies, special attention to developing tools around handling OMOP CDM data is necessary to meaningfully contribute within the observational health research space.
 OHDSI supports the development of open-source software tools in a centralized ecosystem called HADES (Health Analytics Data-to-Evidence Suite) for particularly managing and working with OMOP CDM formatted data.
@@
 
-\subsubsection{JuliaHealth's Role within Observational Health Research}
+\subsection{JuliaHealth and Observational Health Research}
 
@@
 
-\section{Describing Common Observational Health Research Tasks within the Julia Ecosystem}
+\section{Accomplishing Common Observational Health Research Tasks within Julia Ecosystem}
 

### 2023-12-15 paper.tex
Small tweak

@@
 % TODO: Develop this paragraph more about research software
-As the OMOP CDM has become a cornerstone in running these widely distributed network studies, special attention to developing tools around handling OMOP CDM data is necessary to meaningfully contribute within the observational health research space.
+As the OMOP CDM has become ubiquitous within observational health research, developing research software to handle and manage OMOP CDM data is necessary to meaningfully contribute within the observational health research space.
 OHDSI supports the development of open-source software tools in a centralized ecosystem called HADES (Health Analytics Data-to-Evidence Suite) for particularly managing and working with OMOP CDM formatted data.

### 2023-12-22 paper.tex
Further reviews and edits

@@
 
-TODO: Review the abstract after reviewing the paper fully
+%TODO: Review the abstract after reviewing the paper fully
 Observational health continues to be a growing field in health informatics research as electronic health records (EHR), patient medical claims, and other ancilliary patient data source become more readily computable and accessible to researchers.
@@
 
-% EXPLAIN: What is observational health research?
+% EXPLAIN: What is observational health research? Done
 % EXPLAIN: What sort of innovations?
 Over the past $10$ years, there have been several innovations within the observational health research community.
+Observational health research is an area of study that leverages authentic health data to gain insights into the health system without interference.
+Data are collected by researchers as they follow groups of people over a specified period.
+Depending on the study, these groups may comprise healthy individuals, individuals with a specific disease, or those at high risk of developing a particular condition—such as individuals with a family history—while also taking into consideration the imperative of safeguarding their privacy.
+Observational studies can contribute to answering questions as what genes cause a specific disease to grow, patterns and trends for a particular disease, experiences of past patients, behaviours increasing the risk of developing a certain disease.
 
@@
 
-\section{Accomplishing Common Observational Health Research Tasks within Julia Ecosystem}
+\section{Accomplishing Common Observational Health Research Tasks within Julia}
 
-Achieving the notion of a network study in an OHDSI-sense is a highly valuable goal to multiple communities.
-JuliaHealth now has capacity to do just the same and can harness the best of OHDSI and the best of Julia to achieve novel observational health workflows.
-The JuliaHealth observational health subecosystem (referred to as OHS for the remainder of this paper) supports a variety of use cases designed to support and enable investigations into observational health data.
-This paper summarizes some of the current capacity OHS possesses to conduct such endeavors.
-In particular, the following supported areas will be discussed:
+JuliaHealth has capacity through the OHS to perform observational health workflows via a variety of tools designed to support and enable novel investigations.
+This paper summarizes and highlights some of the current capacity OHS provides such as:
 
@@
 \item \textbf{Using Phenotype Definitions} - brief overview of phenotype definitions and how to work with them
-\item \textbf{Rapid Cohort Iteration and Creation} - how to use JuliaHealth tools to iteratively develop patient cohorts to explore
-\item \textbf{Working with Patient Databases} - current approaches to working with different patient databases
-\item \textbf{Ecosystem Interoperability} - how one can bridge between JuliaHealth's ecosystem to other research ecosystems
+\item \textbf{Rapid Cohort Iteration and Creation} - how to iteratively develop patient cohorts to explore
+\item \textbf{Working with Patient Databases} - approaches to working with different patient databases
+\item \textbf{Ecosystem Interoperability} - extending JuliaHealth's ecosystem to other research ecosystems
 \end{enumerate}
@@
 
-One of the premier tools within HADES is the ATLAS Shiny application which was developed to "support the design and execution of observational analyses to generate real world evidence from patient level observational data."
+One of the premier tools within the OHDSI HADES ecosystem is the ATLAS Shiny application which was developed to develop patient phenotype definitions within observational health research contexts (amongst other use cases).
+These phenotype definitions have varying levels of complexity where they can be very simple (such as finding all patients with one health condition or diagnosis) or grow to be extremely complex (defining controlled, uncontrolled, and indeterminate patient cohorts based on medications, lab results, etc.)
+% TODO: Cite schema of the definition from GitHub
+The computable phenotype definitions that ATLAS generates based on phenotype definitions are in a variety of serialization formats such as JSON or prepared SQL expressions available in multiple SQL flavors.
 
-Broadly, ATLAS has become the de facto tool to develop patient phenotype definitions within observational health research contexts.
-These phenotype definitions have varying levels of complexity where they can be very simple (finding all patients with one health condition or diagnosis) or grow to be extremely complex (defining controlled, uncontrolled, and indeterminate patient cohorts based on medications, lab results, etc.)
-The computable phenotype definitions that ATLAS generates based on phenotype definitions are in a variety of serialization formats such as JSON (see \ref{appendix:json}) or prepared SQL expressions available in multiple SQL flavors (see \ref{appendix:postgresql}).
-
+% TODO: Cite package
+% TODO: Reframe this to more say how ATLAS can be used to build cohorts
+% TODO: Explain how this package to accomplish similar tasks
 The Julia package, OHDSICohortExpressions.jl, exists specifically to ingest these serialized expressions of computable phenotype definitions and create patient cohorts that match the given criterion in the definition.
+% TODO: Change this list into an enumeration
 How this tool works is to 1) read and verify the serialized computable phenotype definition 2) reinterpret the serialization into the FunSQL.jl Domain Specific Language 3) return a prepared SQL statement against a given SQL database that queries an OMOP CDM database for the desired patient population(s) of interest.

### 2023-12-23 paper.tex
Added additional todos and experimented with code listings

@@
 
+% TODO: Add a paragraph here about what is a phenotype definition
+% TODO: Cite my paper here
+
+% TODO: Segue this better from phenotype definitions
 One of the premier tools within the OHDSI HADES ecosystem is the ATLAS Shiny application which was developed to develop patient phenotype definitions within observational health research contexts (amongst other use cases).
@@
 % TODO: Cite schema of the definition from GitHub
+% EXPLAIN: Better introduce what SQL is
 The computable phenotype definitions that ATLAS generates based on phenotype definitions are in a variety of serialization formats such as JSON or prepared SQL expressions available in multiple SQL flavors.
@@
 How this tool works is to 1) read and verify the serialized computable phenotype definition 2) reinterpret the serialization into the FunSQL.jl Domain Specific Language 3) return a prepared SQL statement against a given SQL database that queries an OMOP CDM database for the desired patient population(s) of interest.
+% EXPLAIN: What is a prepared statement?
+% EXPLAIN: What is a cohort in this context?
 Once these prepared statements are made, researchers can execute the generated SQL against their OMOP CDM instance to generate cohorts for future analyses.
@@
 
+% TODO: Reword this sentence
 Although OHDSICohortExpressions.jl is sufficient to translate existing phenotype definitions generated in ATLAS to executable queries, oftentimes, there is a tricky problem of how to best iterate on phenotype definitions.
 \cite{zelkoDevelopingRobustComputable2023}
+% EXPLAIN: Better contextualize why this is important -- why not just create a new phenotype definition?
 Often, one has to take an existing definition and either revise to a more narrow cohort or write manual code to build more niche cohorts after an initial definition is created.
+% EXPLAIN: How so?
+% TODO: Review this sentence; maybe cut it
 This refactoring of a phenotype definition can lead to conflating ideas within conversation between expert clinical reviewers, analysts, and health researchers.
+% TODO: Move this to the previous section on phenotype definitions
 This broadly can be the notions of what we are exploring (the phenotype definition; often the responsibility of a clinical reviewer), what is possible to compute upon (the computable phenotype definition; often the responsibility of an analyst), and the goals of the overall study.
-For situations such as these, the tool OMOPCDMCohortCreator.jl was created.
 
+% TODO: Present this more as a general introduction to what tools are here for this
+% TODO: Talk about composition with DataFrames or Tidier
 OMOPCDMCohortCreator.jl allows one to iteratively build a patient population quickly and easily.
 One can use patient pools developed from pre-existing cohorts defined off of phenotype definitions as a starting point in exploring patients or create new patient populations quickly without the need for a well-defined phenotype definition.
+% TODO: Mention the ability to satisfy research requests
 This can enable analysts to more quickly understand what is inside their data assets, do "quick" analysis of a particular smaller research question, and enable a more distinct separation of work between a clinical expert reviewer and analyst teams.
-It is highly recommended to still create phenotype definitions about particular populations one explores.
 
-% TODO: Mention capr here somehow
+% TODO: Mention capr
 
@@
 
+% TODO: Mention JuliaCon Birds of a Feather discussion
 One of the biggest styming factors within not only the Julia ecosystem, but observational health research in general has been the difficulty of working with different databases.
 Although the OMOP CDM has been great for standardizing \textit{how} real world data is handled, there is a lack of consensus on \textit{what} is needed to handle these data assets.
+% TODO: Make this less verbose
 For a variety of reasons, different groups will choose different database architectures (perhaps PostgreSQL for improved permission support, DuckDB for performance, or even SQLite for ease of use) but all are subtly different enough to create headaches in seamlessly working across different research sites.
 
+% TODO: Change the phrasing of this
 The way that the Julia ecosystem handles this is via the creation of a common database interface called the DBInterface.
+% EXPLAIN: What is dispatch
 Specific Julia database packages can implement dispatches for this interace allowing users to have a much better experience working across databases.
-In spite of the such benefits, for some users (and especially within the context of observational health studies), it can beneficial to have a more prescriptive interface where one just needs to pass perhaps a password, username, database name, and schema.
-For that reason, we DBConnector.jl was created.
+In spite of such benefits, for some users (and especially within the context of observational health studies), it can beneficial to have a more prescriptive interface where one just needs to pass perhaps a password, username, database name, and schema.
 
+% TODO: Better introduce this
+For that reason, DBConnector.jl was created.
+% EXPLAIN: Contextualize this a bit better
 Although this is not an explicit JuliaHealth package, it was a pain point that was experienced throughout the process of observational health research across multiple partner sites.
@@
 DBConnector unifies Julia database packages that implement the DBInterface to achieve this level of simplified connection making.
+DBConnector.jl was inspired by the package \textit{DatabaseConnector} (Link: https://github.com/OHDSI/DatabaseConnector)
 
+% TODO: Combine with the previous paragraph and state that there are two approaches
+% TODO: Make explicit DBConnector's path
+% TODO: Make explicit the path of choosing general packages
 Finally, if someone needs more fine-grained control over a particular database connection, they can choose to eschew this package.
@@
 
-DBConnector.jl was inspired by the package \textit{DatabaseConnector} (Link: https://github.com/OHDSI/DatabaseConnector)
-
-% TODO: Add database connector example
-% TODO: Reference database connector example
-
+% TODO: Review if this section is actually needed
 \subsection{Ecosystem Interoperability}
@@
 
-With the JuliaHealth tools that have been created, we can now demonstrate an observational health study workflow that would be used in practice.
+% TODO: Reference code repository
+Using the JuliaHealth tools developed, we enumerate an example observational health study workflow that would be used in practice.\footnotemark
 For this paper, we used two datasets: a synthetic patient dataset called \textit{Eunomia} and the MIMIC III patient dataset which had been converted into the OMOP CDM to illustrate the workflow provided by the Observational Health sub-ecosystem.
@@
 
-% TODO: Add tutorial to separate repository
+% TODO: Add link to code repository
+\footnotetext{The code snippets presented in the body of the paper may omit certain intermediate steps. 
+Please review the associated code repository for a full working tutorial.}
 
@@
 
+\begin{listing}[!ht]
 \begin{minted}[breaklines,escapeinside=||,mathescape=true, numbersep=3pt, gobble=2, frame=lines, fontsize=\scriptsize, framesep=2mm]{julia}
@@
 \end{minted}
+\caption{This illustrates how to connect to the Eunomia database}
+\label{listing:connection}
+\end{listing}
 

### 2023-12-26 paper.tex
Finished full review of JuliaCon paper

@@
 Using the JuliaHealth tools developed, we enumerate an example observational health study workflow that would be used in practice.\footnotemark
+% TODO: Enumerate these two datasets
+% EXPLAIN: In enumeration talk about what Eunomia is 
+% EXPLAIN: In enumeration discuss MIMIC III
+% TODO: Cite MIMIC III, MIT LCP, and OMOP-ification process
 For this paper, we used two datasets: a synthetic patient dataset called \textit{Eunomia} and the MIMIC III patient dataset which had been converted into the OMOP CDM to illustrate the workflow provided by the Observational Health sub-ecosystem.
+% TODO: Mention in footnote explicitly that reproducing workflow in Eunomia is possible
+% TODO: Mention in footnote that explicitly reproducing workflow in MIMIC III is possible but not readily accessible
 Reproducing the workflow using the \textit{Eunomia} dataset is made readily available at the repository here: \cite{schuemieEunomia2023}
@@
 \subsection{Connecting to a Database}
+
+% EXPLAIN: What is this doing here?
+% TODO: Update DBConnector
+% TODO: Maybe remove package imports and mention them within the caption instead
 
@@
 
+% TODO: Review if this subsection is needed
 \subsection{Bridging between OHDSI and JuliaHealth Communities}
@@
 
+% EXPLAIN: What is this doing here?
 After having a strep throat phenotype definition defined using ATLAS, we can then execute the resulting computable phenotype definition against the \textit{Eunomia} database.
 
+% TODO: Add figure, caption, and label details
+% TODO: Maybe remove package imports and mention them within the caption instead
 \begin{minted}[breaklines,escapeinside=||,mathescape=true, numbersep=3pt, gobble=2, frame=lines, fontsize=\scriptsize, framesep=2mm]{julia}
@@
 
+% TODO: Move this to a caption
 A quick query of the cohort table (where the resulting subjects were stored) shows $\approx 1600$ from the dataset match this criteria.
@@
 
+% EXPLAIN: What is this section doing?
+
+% TODO: Add figure, caption, and label details
+% TODO: Maybe remove package imports and mention them within the caption instead
 \begin{minted}[breaklines,escapeinside=||,mathescape=true, numbersep=3pt, gobble=2, frame=lines, fontsize=\scriptsize, framesep=2mm]{julia}
@@
 \subsection{Calculating Crude Prevalence Rate}
+
+% EXPLAIN: What is this section doing?
+
+% TODO: Add figure, caption, and label details
+% TODO: Maybe remove package imports and mention them within the caption instead
 
@@
 Using this analytical approach, tabulation can be quickly conducted across patients.
+
+% EXPLAIN: Move this to the caption of the table
 Additionally, throughout the process, one can perform relevant analysis methods.
@@
 
+% EXPLAIN: What is this section doing?
+
+% TODO: Add figure, caption, and label details
 \begin{table}[!ht]
@@
 
-As seen in Table 1 and Table 2, informative results can be rapidly generated by this workflow relevant to observational health research.
+% TODO: Create table for MIMIC III results
+% TODO: Create reference to Eunomia table
+As seen in Table 1, informative results can be rapidly generated by this workflow relevant to observational health research.
 Although the values shown in Table 2 are synthetic and the crude prevalence values themselves are meaningless, we can take a this approach and apply it directly to the MIMIC III dataset.
@@
 
+% EXPLAIN: What is the overview of this section?
+% EXPLAIN: What does each topic do?
+% TODO: Convert this to an enumeration
+
     1. Modularization of code and chaining
     2. Distributed computing
+    3. Automatic generation of SQL from health packages
 
+% TODO: Rename section
 \subsubsection{Modularization of Code and Chaining}
 
+% TODO: Make this more explicit about utilizing Julia's composition
 An exciting emergent aspect of the JuliaHealth ecosystem that came about as a result of developing OMOPCDMCohortCreator.jl was the prospect of radical modularization and chaining of package functionalities together.
+% TODO: Tie into currying more directly
+% TODO: Check if I am mentioning currying correctly
+We can leverage a unique property of Julia in that Julia supports partial evaluation of functions (also known as currying) to simplify repeated function calls to the same function over and over again.
+% TODO: Rewrite this part in light of the previous sections 
 For example, taking our workflow example from the earlier section where we characterized patients by race, gender, and age group, we can do a few "tricks" to modularize these characterization functions as follows:
 
-We can leverage a unique property of Julia in that Julia supports partial evaluation of functions (also known as currying) to simplify repeated function calls to the same function over and over again.
+% TODO: Rework this section further
 First, we'll take our original functions and fix a connection object to each of our cofactor functions.
@@
 
+% TODO: Maybe move this into the tutorial code
+% TODO: Add figure, caption, and label details
 \begin{minted}[breaklines,escapeinside=||,mathescape=true, numbersep=3pt, gobble=2, frame=lines, fontsize=\scriptsize, framesep=2mm]{julia}
@@
 
+% TODO: Add this to previous code caption
 NOTE: The "C" prefix for each of our curried functions denotes that it is curried.
 
+% EXPLAIN: Refactor this paragraph better or put most of it into the caption for the next code block
 Second, as OMOPCDMCohortCreator.jl is built on top of DataFrames.jl and one has the option of creating a DataFrame with the result from every OCC function, we can use the package, Chain.jl.
@@
 
+% TODO: Do we even need chaining here or could we use direct composition of functions?
 \begin{minted}[breaklines,escapeinside=||,mathescape=true, numbersep=3pt, gobble=2, frame=lines, fontsize=\scriptsize, framesep=2mm]{julia}
@@
 
+% TODO: Can this be simplified?
 \begin{minted}[breaklines,escapeinside=||,mathescape=true, numbersep=3pt, gobble=2, frame=lines, fontsize=\scriptsize, framesep=2mm]{julia}
@@
 In this way, future researchers using tools such as OMOPCDMCohortCreator.jl can create these more convenient abstractions.
+% TODO: Move this into the future discussion section
 In fact, it could be a point of exploration in the future to build on top of OMOPCDMCohortCreator.jl even more abstracted and reusable functionalities to simplify ease of use while also enabling more readily auditable functionality.
+
+% TODO: Move this into its own section
 Finally, as these functions have now been fully modularized, these can now be readily distributed using Julia's distributed computing methods:
 
+% TODO: Redo example; I do not feel it is good
 \begin{minted}[breaklines,escapeinside=||,mathescape=true, numbersep=3pt, gobble=2, frame=lines, fontsize=\scriptsize, framesep=2mm]{julia}
@@
 
+% TODO: Rephrase this sentence
 One crucial aspect of the OHS is that it is unavoidable to work with databases that contain patient health information.
+% TODO: Rephrase this sentence better
+% TODO: Cite ANSI SQL
 For better or worse, the tool that is used to best interact with these databases are SQL dialects that loosely follow consistently the ANSI SQL guidance.
+% EXPLAIN: Highlight this more
+% EXPLAIN: In what scenarios might using Julia may not be possible
 To accomodate for this, several of the core tools developed in this sub-ecosystem generate SQL that can be used in the case where Julia may not be available in a given environment housing patient health data.
+% EXPLAIN: Introduce this better
 As an example, \textit{OMOPCDMCohortCreator.jl} functions build upon \textit{FunSQL.jl} allowing one to use a function such `GetPatientAgeGroup` without a connection object which would result in SQL.
 
+% TODO: Add figure, caption, and label details
 \begin{minted}[breaklines,escapeinside=||,mathescape=true, numbersep=3pt, gobble=2, frame=lines, fontsize=\scriptsize, framesep=2mm]{julia}
@@
 
+% TODO: Add figure, caption, and label details
 \begin{minted}[breaklines,escapeinside=||,mathescape=true, numbersep=3pt, gobble=2, frame=lines, fontsize=\scriptsize, framesep=2mm]{julia}
@@
 
+% TODO: Combine this section with the future directions section
 \subsection{Strong Composition within JuliaHealth}
@@
 
-As shown in this work, there is a rich opportunity for the JuliaHealth organization to become a strong presence within observational health research. 
+% TODO: Rephrase this introductory paragraph
 JuliaHealth is now sufficiently at a stage of maturity where potential future researchers can leverage JuliaHealth's tools to conduct further future research or build upon existing architecture to target specific needs.
@@
 
+% TODO: This is where to introduce the paradox of composition
 Due to the strong composition that exists within the JuliaHealth and broader Julia ecosystem, solutions may already exist for solving various health informatics related research.
@@
 
+% TODO: Combine this paragraph with the previous paragraph
+% TODO: Shorten this sentiment
 Additionally, due to the composable and modular nature of the OHS, research analyses and visualizations can be decoupled and interchanged with one another.
@@
 
+% TODO: Make this more about the flexible nature of the JuliaHealth organization
 Furthermore, again exploiting the composition within the JuliaHealth ecosystem, as mentioned in the context of JuliaHealth and machine learning applications, future efforts could focus on developing fairness and bias auditing tools in patient datasets.
 These tools could consider factors commonly explored in health equity literature like data completeness, demographic variables, and fairness algorithms.
-
-Moreover, there's potential to advance clinical informatics by developing a flexible clinical type system, supporting computable phenotype definitions and medical ontologies.
-The ecosystem can also benefit from tools to audit the fairness of phenotype definitions.
-Finally, as the landscape of data expands, integrating non-traditional sources like social media and environmental data into standardized patient records will be a critical frontier.
-In these ways, JuliaHealth is primed to drive innovation in observational health informatics, pushing the boundaries of data-driven healthcare research.
 
@@
 Although it is an ecosystem that is relatively new, its robustness allows for novel investigations to be pursued within this domain of research. 
+% TODO: Mention how this allows JuliaHealth to solve more problems than what is present within JuliaHealth
 Additionally, due to the strongly compositional nature of the Julia ecosystem, these JuliaHealth tools can readily compose with tooling to support in-depth statistical analyses, mathematical modeling, and visualization capacities.
 
-The deep dive into the JuliaHealth ecosystem unravels its arsenal of tools tailored for observational health research, meticulously designed to navigate the complexities of handling OMOP CDM data.
-Notably, the sub-ecosystem within JuliaHealth presents mature, internationally federated research capabilities, building upon existing innovations like OHDSI's HADES, while offering novel tools such as OMOPCDMCohortCreator.jl.
-In considering future directions, the paper posits the promising avenues for JuliaHealth, envisioning expanded documentation, the development of visualization tools, and integration of machine learning applications for fairness auditing.
-The ecosystem's potential to revolutionize clinical informatics by integrating non-traditional data sources, auditing phenotype definitions, and fostering innovation in data-driven healthcare research is paramount.
+% TODO: Rework this
 In essence, the JuliaHealth ecosystem stands as a catalyst for transformative observational health research, poised to push the boundaries of innovation, collaboration, and standardization in the pursuit of better healthcare outcomes.
@@
   \item Kristin Kostka for being an early champion in trying novel ideas
+  \item The JuliaHealth community and contributors for fearlessly trying to push boundaries on what observational health informatics can truly be
   \item Dilum Aluthge for helping this work find a home in JuliaHealth
   \item Adam Black for thinking through potential approaches
+  \item Clark Evans and Kirill Simonov for being the pioneers of this work in Julia
   \item The Roux Institute's OHDSI Center for their encouragement
-  \item The JuliaHealth community and contributors for fearlessly trying to push boundaries on what observational health informatics can truly be
 
@@
 \section{References}
+
+% TODO: Review and remove appendix
 

### 2023-12-26 reference.tex
Fixing up todos

% JuliaCon proceedings template
\documentclass{juliacon}
\setcounter{page}{1}

\begin{document}

\input{header}

\maketitle

\begin{abstract}

preparation system and the \verb|juliacon|  class file.

\end{abstract}

\section{Introduction}

The \LaTeX{} document preparation system is a special version of the
\TeX; typesetting program where-in a collection of \TeX; commands
are added to \LaTeX{} to simplify typesetting. Importantly, it allows
the author to concentrate on the logical structure of the document
rather than its visual layout.\vskip 6pt
Moreover, \LaTeX{} provides a consistent and comprehensive document
preparation interface. There are simple-to-use commands for
generating a table of contents, lists of figures and/or tables, and indexes.
\LaTeX{} can automatically number list entries, equations, figures,
tables, and footnotes, as well as articles, sections, and subsections.
Using this numbering system, bibliographic citations, page
references, and cross references to any other numbered entity (e.g.
article, section, equation, figure, list entry, etc.) become quite simple
and straightforward. The use of \LaTeX{} document classes allows
a simple change of class to transform the appearance of your document.\vskip 6pt
\LaTeX{} is a powerful tool for managing long and complex documents.
In particular, partial processing enables long documents to
be produced article by article without losing sequential information.
The use of document classes allows a simple change of style
(or style option) to transform the appearance of your document.

\section{The JuliaCon Article Class}
\label{sec:documentclass}
%
The juliacon class file preserves the standard LATEX{} interface such
that any document that can be produced using the standard LATEX{}
article class can also be produced with the class file.\vskip 6pt
It is likely that the make up will change after file submission. For
this reason, we ask you to ignore details such as slightly long lines,
page stretching, or figures falling out of synchronization, as these
details can be dealt with at a later stage.\vskip 6pt
Use should be made of symbolic references (\verb|\ref| ) in order to
protect against late changes of order, etc.

\section{USING THE JuliaCon Article CLASS FILE}

If the file \verb|juliacon.cls|  is not already in the appropriate system directory
for \LaTeX{} files, either arrange for it to be put there or copy
it to your working directory. The \verb|juliacon|  document class is implemented
as a complete class, not a document style option. In order to
use the \verb|juliacon  document class, replace \verb|article|  by \verb|juliacon|  in the
\verb|\documentclass|  command at the beginning of your document:
\vskip 6pt
\begin{centering}
    \verb|\documentclass{article}|  \end{centering}
\vskip 6pt
replace by
\vskip 6pt
 \verb|\documentclass{juliacon}|  \vskip 6pt
In general, the following standard document \verb|style|  options should
{ \itshape not} be used with the {\footnotesize \itshape article} class file:
\begin{enumerate}
\item[(1)] \verb|10pt|,  \verb|11pt|,  \verb|12pt|   ? unavailable;
\item[(2)] \verb|twoside|  (no associated style file) ? \verb|twoside|  is the default;
\item[(3)] \verb|fleqn|, \verb|leqno|, \verb|titlepage| ? should not be used;
\end{enumerate}

\section{Additional Document Style Options}
\label{sec:additional_doc}
%
The following additional style option is available with the \verb|juliacon|  class file:
\vskip 6pt
Please place any additional command definitions at the very start of
the \LaTeX{} file, before the \verb|\begin{document}|. For example, user-defined
\verb|\def|  and \verb|\newcommand|   commands that define macros for
technical expressions should be placed here. Other author-defined
macros should be kept to a minimum.
\vskip 6pt
Commands that differ from the standard \LaTeX{} interface, or that
are provided in addition to the standard interface, are explained in
this guide. This guide is not a substitute for the \LaTeX{} manual itself.
Authors planning to submit their papers in \LaTeX{} are advised to use
\verb|juliacon.cls|  as early as possible in the creation of their files.

%
%
%
%
\begin{table*}[t]
\tabcolsep22pt
\tbl{If necessary, the tables can be extended both columns.}{
\begin{tabular}{|l|l|c|c|}\hline
Label & \multicolumn{1}{c|}{Description}
& Number of Users &
Number of Queries\\\hline
Test 1 & Training Data &
\smash{\raise-7pt\hbox{70}} & 104\\
\cline{1-2}\cline{4-4}
Test 2 & Testing Data I & & 105\\\hline
Test 3 & Testing Data II & 30 & 119\\\hline
& Total & 100 & 328\\\hline
\end{tabular}}
\label{tab:symbols}
\begin{tabnote}
This is an example of table footnote.
\end{tabnote}
\end{table*}
% \begin{figure*}[t]
% \centerline{\includegraphics[width=11cm]{juliagraphs.png}}
% \caption{If necessary, the images can be extended both columns.}
%   \label{fig:sample_image}
% \end{figure*}

\section{Additional features}
\label{sec:additional_faci}
In addition to all the standard \LaTeX{} design elements, the \verb|juliacon|  class file includes the following features:
In general, once you have used the additional \verb|juliacon.cls| facilities
in your document, do not process it with a standard \LaTeX{} class
file.

\subsection{Titles, Author's Name, and Affiliation}
\label{subsub:title_auth}
The title of the article, author's name, and affiliation are used at the
beginning of the article (for the main title). These can be produced
using the following code:

\begin{verbatim}
\title{ This is an example of article title} }
\author{
   \large 1st Author \\[-3pt]
   \normalsize 1st author's affiliation  \\[-3pt]
    \normalsize 1st line of address \\[-3pt]
    \normalsize 2nd line of address \\[-3pt]
    \normalsize	1st author's email address \\[-3pt]
  \and
   \large 2nd Author \\[-3pt]
   \normalsize 2nd author's affiliation  \\[-3pt]
    \normalsize 1st line of address \\[-3pt]
    \normalsize 2nd line of address \\[-3pt]
    \normalsize	2nd author's email address \\[-3pt]
\and
   \large 3rd Author \\[-3pt]
   \normalsize 3rd author's affiliation  \\[-3pt]
    \normalsize 1st line of address \\[-3pt]
    \normalsize 2nd line of address \\[-3pt]
    \normalsize	3rd author's email address \\[-3pt]
}
\maketitle
\end{verbatim}

\subsection{Writing Julia code}

A special environment is already defined for Julia code,
built on top of \textit{listings} and \textit{jlcode}.

\begin{verbatim}
\begin{lstlisting}[language = Julia]
using Plots

x = -3.0:0.01:3.0
y = rand(length(x))
plot(x, y)
\end{lstlisting}
\end{verbatim}
\begin{lstlisting}[language = Julia]
using Plots

x = -3.0:0.01:3.0
y = rand(length(x))
plot(x, y)
\end{lstlisting}

\subsection{Abstracts, Key words, term etc...}
\label{subsub:abs_key_etc}

At the beginning of your article, the title should be generated
in the usual way using the \verb|\maketitle|  command. For genaral tem and keywords use
\verb|\terms|,
\verb|\keywords|  commands respectively. The abstract should be enclosed
within an abstract environment, All these environment
can be produced using the following code:
\begin{verbatim}
\terms{Experimentation, Human Factors}

\keywords{Face animation, image-based modelling...}

\begin{abstract}
In this paper, we propose a new method for the
systematic determination of the model's base of
time varying delay system. This method based on
the construction of the classification data related
to the considered system. The number, the orders,
the time delay and the parameters of the local
models are generated automatically without any
knowledge about the full operating range of the
process. The parametric identification of the local
models is realized by a new recursive algorithm for
on line identification of systems with unknown time
delay. The proposed algorithm allows simultaneous
estimation of time delay and parameters of
discrete-time systems. The effectiveness of
the new method has been illustrated through
simulation.
\end{abstract}

\end{verbatim}

\section{Some guidelines}
\label{sec:some_guide}
The following notes may help you achieve the best effects with the
\verb|juliacon|  class file.

\subsection{Sections}
\label{subsub:sections}
\LaTeXe{}  provides four levels of section headings and they are all
defined in the \verb|juliacon|  class file:
\begin{itemize}
\item \verb|\section|
\item \verb|\subsection|
\item \verb|\subsubsection|
\item \verb|\paragraph|
\end{itemize}
Section headings are automatically converted to allcaps style.
\subsection{Lists}
\label{sec:lists}
%
The \verb|juliacon|   class file provides unnumbered lists using the
\verb|unnumlist|  environment for example,

\begin{unnumlist}
\item First unnumbered item which has no label and is indented from the
left margin.
\item Second unnumbered item.
\item Third unnumbered item.
\end{unnumlist}
The unnumbered list which has no label and is indented from the
left margin. was produced by:
\begin{verbatim}
\begin{unnumlist}
\item First unnumbered item...
\item Second unnumbered item...
\item Third unnumbered item...
\end{unnumlist}
\end{verbatim}

The \verb|juliacon|  class file also provides hyphen list using the
\verb|itemize|  environment for example,
\begin{itemize}
\item First unnumbered bulleted item which has no label and is indented
from the left margin.
\item Second unnumbered bulleted item.
\item Third unnumbered bulleted item which has no label and is indented
from the left margin.
\end{itemize}
was produced by:
\begin{verbatim}
\begin{itemize}
\item First item...
\item Second item...
\item Third item...
\end{itemize}
\end{verbatim}

Numbered list is also provided in acmtog class file using the
enumerate environment for example,
\begin{enumerate}
\item The attenuated and diluted stellar radiation.
\item Scattered radiation, and
\item Reradiation from other grains.
\end{enumerate}

was produced by:
\begin{verbatim}
\begin{enumerate}
\item The attenuated...
\item Scattered radiation, and...
\item Reradiation from other grains...
\end{enumerate}
\end{verbatim}
\subsection{Illustrations (or figures)}
\label{subsub:sec_Illus}
The \verb|juliacon|  class file will cope with most of the positioning of
your illustrations and you should not normally use the optional positional
qualifiers on the \verb|figure|  environment that would override
these decisions.
\vskip 6pt

%
\begin{figure}[t]
\centerline{\includegraphics[width=4cm]{juliagraphs.png}}
\caption{This is example of the image in a column.}
	\label{fig:sample_figure}
\end{figure}

The figure \ref{fig:sample_figure} is taken from the JuliaGraphs
organization \footnote{https://github.com/JuliaGraphs}.

Figure captions should be \emph{below} the figure itself, therefore the
\verb|\caption|  command should appear after the figure or space left for
an illustration. For example, Figure 1 is produced using the following
commands:

\begin{verbatim}
\begin{figure}
\centerline{\includegraphics[width=20pc]{Graphics.eps}}
\caption{An example of the testing process for a
binary tree. The globa null hypothesis is tested
first at level $\alpha$ (a), and the level of
individual variables is reached last (d). Note
that individual hypotheses can be tested at
level $\alpha/4$ and not $\alpha/8$ as one might
expect at first.}
\label{sample-figure_2}
\end{figure}
\end{verbatim}
Figures can be resized using first and second argument of
\verb|\includegraphics|   command. First argument is used for modifying
figure height and the second argument is used for modifying
figure width respectively.
\vskip 6pt
Cross-referencing of figures, tables, and numbered, displayed
equations using the \verb|\label|  and \verb|\ref|   commands is encouraged.
For example, in referencing Figure 1 above, we used
\verb|Figure~\ref{sample-figure}|

 \subsection{Tables}
\label{subsub:sec_Tab}
The \verb|juliacon|   class file will cope with most of the positioning of
your tables and you should not normally use the optional positional qualifiers on the table environment which would override these
decisions. Table captions should be at the top.
\begin{verbatim}
\begin{table}
\tbl{Tuning Set and Testing Set}{
\begin{tabular}{|l|l|c|c|}\hline
Label & \multicolumn{1}{c|}{Description}
& Number of Users &
Number of Queries\\\hline
Train70 & Training Data &
\smash{\raise-7pt\hbox{70}} & 104\\
\cline{1-2}\cline{4-4}
Test70 & Testing Data I & & 105\\\hline
Test30 & Testing Data II & 30 & 119\\\hline
& Total & 100 & 328\\\hline
\end{tabular}}
\end{table}
\end{verbatim}

\begin{table}
\tbl{Tuning Set and Testing Set}{
\begin{tabular}{|l|l|c|c|}\hline
Label & \multicolumn{1}{c|}{Description}
& Number of Users &
Number of Queries\\\hline
Test 1 & Training Data &
\smash{\raise-7pt\hbox{70}} & 104\\
\cline{1-2}\cline{4-4}
Test 2 & Testing Data I & & 105\\\hline
Test 3 & Testing Data II & 30 & 119\\\hline
& Total & 100 & 328\\\hline
\end{tabular}}
\end{table}
\subsection{Landscaping Pages}
\label{subsub:landscaping_pages}
If a table is too wide to fit the standard measure, it may be turned,
with its caption, to 90 degrees. Landscape tables cannot be produced
directly using the \verb|juliacon|   class file because \TeX{} itself cannot
turn the page, and not all device drivers provide such a facility.
The following procedure can be used to produce such pages.
\vskip 6pt
Use the package \verb|rotating|   in your document and change the coding
from
\begin{verbatim}
\begin{table}...\end{table}
to
\begin{sidewaystable}...\end{sidewaystable}
and for figures
\begin{figure}...\end{figure}
to
\begin{sidewaysfigure}...\end{sidewaysfigure}
\end{verbatim}

environments in your document to turn your table on the appropriate
page of your document. For instance, the following code prints
a page with the running head, a message half way down and the
table number towards the bottom.
\begin{verbatim}
\begin{sidewaystable}
\tbl{Landscape table caption to go here.}{...}
\label{landtab}
\end{sidewaystable}
\end{verbatim}

\subsection{Double Column Figure and Tables}
\label{subsub:double_fig_tab}
For generating the output of figures and tables in double column
we can use the following coding:

\begin{enumerate}
\item For Figures:
\begin{verbatim}
\begin{figure*}...\end{figure*}
\end{verbatim}
\item For landscape figures:
\begin{verbatim}
\begin{sidewaysfigure*}...\end{sidewaysfigure*}
\end{verbatim}
\item For Tables:
\begin{verbatim}
\begin{table*}...\end{table*}
\end{verbatim}
\item For landscape tables:
\begin{verbatim}
\begin{sidewaystable*}...\end{sidewaystable*}
\end{verbatim}
\end{enumerate}

\subsection{Typesetting Mathematics}
\label{subsub:type_math}
The \verb|juliacon| class file will set displayed mathematics with center to
the column width, provided that you use the \LaTeXe{} standard of
open and closed square brackets as delimiters.
The equation
\[
\sum_{i=1}^p \lambda_i = (S)
\]

was typeset using the acmtog class file with the commands

\begin{verbatim}
\[
\sum_{i=1}^p \lambda_i = (S)
\]
\end{verbatim}

For display equations, cross-referencing is encouraged. For example,
\begin{verbatim}
\begin{equation}
(n-1)^{-1} \sum^n_{i=1} (X_i - \overline{X})^2.
\label{eq:samplevar}
\end{equation}
Equation~(\ref{eq:samplevar}) gives the formula for
sample variance.
\end{verbatim}
The following output is generated with the above coding:
\begin{equation}
(n-1)^{-1} \sum^n_{i=1} (X_i - \overline{X})^2.
\label{eq:samplevar}
\end{equation}
Equation~(\ref{eq:samplevar}) gives the formula for
sample variance.

\subsection{Enunciations}
\label{subsub:enunciation}
The \verb|juliacon|   class file generates the enunciations with the help of
the following commands:
\begin{verbatim}
\begin{theorem}...\end{theorem}
\begin{strategy}...\end{strategy}
\begin{property}...\end{property}
\begin{proposition}...\end{proposition}
\begin{lemma}...\end{lemma}
\begin{example}...\end{example}
\begin{proof}...\end{proof}
\begin{definition}...\end{definition}
\begin{algorithm}...\end{algorithm}
\begin{remark}...\end{remark}
\end{verbatim}
The above-mentioned coding can also include optional arguments
such as
\begin{verbatim}
\begin{theorem}[...]. Example for theorem:
\begin{theorem}[Generalized Poincare Conjecture]
Four score and seven ... created equal.
\end{theorem}
\end{verbatim}

\begin{theorem}[Generalized Poincare Conjecture]
Four score and seven years ago our fathers brought forth,
upon this continent, a new nation, conceived in Liberty,
 and dedicated to the proposition that all men are
created equal.
\end{theorem}

\subsection{Extract}
\label{subsub:extract}
Extract environment should be coded within
\begin{verbatim}
\begin{extract}..\end{extract}
\end{verbatim}

\subsection{Balancing column at last page}
\label{subsub:Balance}
For balancing the both column length at last page use :
\begin{verbatim}
\vadjust{\vfill\pagebreak}
\end{verbatim}

%\vadjust{\vfill\pagebreak}

at appropriate place in your \TeX{} file or in bibliography file.

\section{Handling references}
\label{subsub:references}
References are most easily (and correctly) generated using the
BIBTEX, which is easily invoked via
\begin{verbatim}
\bibliographystyle{juliacon}
\bibliography{ref}
\end{verbatim}
When submitting the document source (.tex) file to external
parties, the ref.bib file should be sent with it.
\cite{bezanson2017julia}

\input{bib.tex}

\end{document}

% Inspired by the International Journal of Computer Applications template

### 2023-12-28 paper.tex
Fixing up todos
Resolved some revisions with citations

--- 2023-12-23 paper.tex
@@
 
-%TODO: Review the abstract after reviewing the paper fully
+% TODO::HARD Review the abstract after reviewing the paper fully
 Observational health continues to be a growing field in health informatics research as electronic health records (EHR), patient medical claims, and other ancilliary patient data source become more readily computable and accessible to researchers.
@@
 
-% EXPLAIN: What is observational health research? Done
-% EXPLAIN: What sort of innovations?
+% EXPLAIN::EASY What is observational health research?
+% EXPLAIN::MEDIUM What sort of innovations?
 Over the past $10$ years, there have been several innovations within the observational health research community.
@@
 
-% EXPLAIN: Why is this a crucial innovation?
-% EXPLAIN: What is meant by a transferable analysis
-% TODO: Cite: J.
-M.
-Overhage, P.
-B.
-Ryan, C.
-G.
-Reich, A.
-G.
-Hartzema, and P.
-E.
-Stang, “Validation of a common data model for active safety surveillance research,” Journal of the American Medical Informatics Association, vol.
-19, no.
-1, pp.
-54–60, 2012.
+% EXPLAIN::MEDIUM Why is this a crucial innovation?
+% EXPLAIN::MEDIUM What is meant by a transferable analysis?
 Arguably, the most crucial innovation that has emerged is the Observational Medical Outcomes Partnership Common Data Model (OMOP CDM) which was designed specifically to standardize observational health data to enable transferable analyses.
-% TODO: Add citation from OHDSI website's collaborator statistics
-% EXPLAIN: Why is mapping relevant here?
-The importance of this model has become apparent within open science organizations like the OHDSI (Observational Health Data Science and Informatics) network where members across $100+$ countries with access to various health systems have been able to successfully map their patient data to the OMOP CDM.
-% TODO: Add citation about successful collaborations
-% EXPLAIN: What is a network study?
-% TODO: Cite the quote "write once, run everywhere" 
+\cite{overhage2012validation}
+% EXPLAIN::MEDIUM Why is mapping relevant here?
+The importance of this model has become apparent within open science organizations like the OHDSI (Observational Health Data Science and Informatics) network where members across $100+$ countries \cite{sachsonOurJourney2023} with access to various health systems have been able to successfully map their patient data to the OMOP CDM.
+% TODO::EASY Add citation about successful collaborations
+% EXPLAIN::EASY What is a network study?
+% TODO::EASY Cite the quote "write once, run everywhere" 
 % http://www.sun.com/smi/Press/sunflash/1996-01/sunflash.960123.10561.xml} 
@@
 
-% TODO: This paragraph should be centered on the core technology of data standards -- in particular, the OMOP CDM
-% EXPLAIN: What is real world data? Use https://jacobzelko.com/10282021140730-real-world-evidence/
+% TODO::MEDIUM This paragraph should be centered on the core technology of data standards -- in particular, the OMOP CDM
+% EXPLAIN::MEDIUM What is real world data? Use https://jacobzelko.com/10282021140730-real-world-evidence/
 OMOP was created in 2009 by the US Food and Drug Administration to address concerns related to data type, study design, and data privacy concerns of "Real World Data" \cite{ohdsi2019book} \cite{FDARealWorldEvidence}
 What could arguably be considered the longest standing achievement of OMOP was the development of the OMOP Common Data Model (OMOP CDM) to house and ingest real world data.
-% TODO: Combine the next two sentences together
-% EXPLAIN: What does "person-centered" mean?
-% EXPLAIN: What is a patient event? Could you give an example?
+% TODO::EASY Combine the next two sentences together
+% EXPLAIN::EASY What does "person-centered" mean?
+% EXPLAIN::MEDIUM What is a patient event? Could you give an example?
 in a consistent format that has been optimized to assist in extracting meaningful population health information.
 In brief, the OMOP Common Data Model (CDM) is a "person-centered" model that has standardized common events in a given patient encounter with database tables that represent the unique structures found in patient events.
-% TODO: Develop this fragment into a full sentence
+% TODO::EASY Develop this fragment into a full sentence
 prioritizing data protection through the limiting of information that could endanger patient anonymity.
-% EXPLAIN: What is a health system? Could you give an example?
+% EXPLAIN:: What is a health system? Could you give an example?
 This creates a longitudinal view of all healthcare-related events per person recorded in a singular health system.
-% TODO: Make this more coherent to the rest of the paragraph
-% TODO: Explicitly mention that although the model prescribes a database schema, it does not need a specific database type
+% TODO::EASY Make this more coherent to the rest of the paragraph
+% TODO::EASY Explicitly mention that although the model prescribes a database schema, it does not need a specific database type
 The OMOP CDM itself does not require a specific technology to work with the data stored in this standard.
 
-% TODO: This paragraph should be centered on the core technology of the concept of network studies
-% TODO: Use this to address the notion of what "network studies" are
-% TODO: Make this initially less OHDSI specific; pull from https://jacobzelko.com/05122023034151-network-macarthur
+% TODO::MEDIUM This paragraph should be centered on the core technology of the concept of network studies
+% TODO::MEDIUM Use this to address the notion of what "network studies" are
+% TODO::MEDIUM Make this initially less OHDSI specific; pull from https://jacobzelko.com/05122023034151-network-macarthur
 In particular, one collaboration mechanism that OHDSI has been able to operationalize meaningfully while working with observational health data is the notion of "network studies".
@@
 The goal of OHDSI is to produce reproducible, robust, and reliable evidence by developing standardized data formats and research protocol designs to enable large-scale collaborative research and improve the transparency and scalability of observational health research.
-% TODO: Carry HADES discussion into the paragraph talking about research tooling
+% TODO::MEDIUM Carry HADES discussion into the paragraph talking about research tooling
 One core innovation brought about by OHDSI in the space of network science is how, given the HADES tools and procedures established by OHDSI, 
-% TODO: Make this more general such that it flows into the next paragraph on research tooling better
+% TODO::EASY Make this more general such that it flows into the next paragraph on research tooling better
 a research group can develop their analysis on their own OMOP CDM instance and then bundle together their study.
-% EXPLAIN: What is a study package?
-% EXPLAIN: What data assets?
+% EXPLAIN::EASY What is a study package?
+% EXPLAIN::EASY What data assets?
 Once this packaging is done, they can pass along the study package to another site to collaborate on the same analysis but on different data assets.
-% EXPLAIN: What is a collaborator network?
+% EXPLAIN::MEDIUM What is a collaborator network?
 Data confidentiality is preserved in this approach as data from individual participating sites remains exclusive to each site, with only aggregate results shared across a given collaborator network.
@@
 
-% TODO: Develop this paragraph more about research software
+% TODO::HARD Develop this paragraph more about research software
 As the OMOP CDM has become ubiquitous within observational health research, developing research software to handle and manage OMOP CDM data is necessary to meaningfully contribute within the observational health research space.
@@
 
-% TODO: Cite Discourse post on start of JuliaHealth
-JuliaHealth is an open-science collective dedicated to bringing the Julia programming language to various avenues of health research domains including, but not limited to, observational health, medical imaging, standards interoperability and more.
+JuliaHealth is an open-science collective dedicated to bringing the Julia programming language to various avenues of health research domains including, but not limited to, observational health, medical imaging, standards interoperability and more.\cite{aluthgeAnnouncingJuliaHealthOrganization2020}
 It leverages the strengths of the Julia language to introduce new methods into these spaces, augment existing workflows, and provide a community structure for health researchers within the global Julia ecosystem.
-% EXPLAIN: What is a subecosystem?
-% EXPLAIN: How does having more subecosystems enable novel research to be done?
+% EXPLAIN::EASY What is a subecosystem?
+% EXPLAIN:: How does having more subecosystems enable novel research to be done?
 Eventually, JuliaHealth will host a variety of subecosystems and associated resources that will enable novel research to be conducted within Julia.
@@
 One area of the subecosystems within JuliaHealth that has matured significantly over the years has been the JuliaHealth observational health subecosystem (OHS).
-It can be said that OHS had its inception in work done by Evans and Simonov with efforts such as \textit{FunOHDSI.jl}, \textit{OHDSICohortExpressions.jl}, and \textit{FunSQL.jl} around $2021$ \cite{kirill_simonov_2023_7705325}
-% TODO: Cite FunOHDSI
-% TODO: Cite OHDSICohortExpressions
-% TODO: Cite OHDSI Symposium poster on JuliaHealth
-% TODO: Register OHDSI Symposium poster to Zenodo
-% TODO: Cite JuliaCon 2023 presentation
-% TODO: Register JuliaCon 2023 presentation to Zenodo
+It can be said that OHS had its inception in work done by Evans and Simonov with efforts such as \textit{FunOHDSI.jl} \cite{evansFunOHDSIJl2022}, \textit{OHDSICohortExpressions.jl} \cite{evansOHDSICohortExpressionsJl2023}, and \textit{FunSQL.jl} around $2021$ \cite{kirill_simonov_2023_7705325}
+% TODO::EASY Cite OHDSI Symposium poster on JuliaHealth
+% TODO::HARD Register OHDSI Symposium poster to Zenodo
+% TODO::EASY Cite JuliaCon 2023 presentation
+% TODO::HARD Register JuliaCon 2023 presentation to Zenodo
 Since then, Zelko et.
@@
 have spearheaded most recent developments within OHS via the creation of a variety of tools and demonstrated their effectiveness in using these tools for novel observational health research endeavors \cite{zelko2022pilot}
-% EXPLAIN: What does this mean?
+% EXPLAIN::MEDIUM What does this mean?
 Through such efforts, OHS is becoming further capable of executing novel observational health research.
@@
 
-% TODO: Add a paragraph here about what is a phenotype definition
-% TODO: Cite my paper here
+% TODO::MEDIUM Add a paragraph here about what is a phenotype definition
+% TODO::EASY Cite my paper here
 
-% TODO: Segue this better from phenotype definitions
+% TODO::EASY Segue this better from phenotype definitions
 One of the premier tools within the OHDSI HADES ecosystem is the ATLAS Shiny application which was developed to develop patient phenotype definitions within observational health research contexts (amongst other use cases).
 These phenotype definitions have varying levels of complexity where they can be very simple (such as finding all patients with one health condition or diagnosis) or grow to be extremely complex (defining controlled, uncontrolled, and indeterminate patient cohorts based on medications, lab results, etc.)
-% TODO: Cite schema of the definition from GitHub
-% EXPLAIN: Better introduce what SQL is
+% TODO::EASY Create footnote about schema
+% TODO::EASY Add reference to schema in footnote
+% TODO::EASY Cite reference schema from GitHub in footnote; \cite{madrilATLASJSONSchema2022}
+% EXPLAIN::EASY Better introduce what SQL is
 The computable phenotype definitions that ATLAS generates based on phenotype definitions are in a variety of serialization formats such as JSON or prepared SQL expressions available in multiple SQL flavors.
 
-% TODO: Cite package
-% TODO: Reframe this to more say how ATLAS can be used to build cohorts
-% TODO: Explain how this package to accomplish similar tasks
+% TODO::MEDIUM Reframe this to more say how ATLAS can be used to build cohorts
+% TODO::EASY Explain how this package accomplishes similar tasks
 The Julia package, OHDSICohortExpressions.jl, exists specifically to ingest these serialized expressions of computable phenotype definitions and create patient cohorts that match the given criterion in the definition.
-% TODO: Change this list into an enumeration
-How this tool works is to 1) read and verify the serialized computable phenotype definition 2) reinterpret the serialization into the FunSQL.jl Domain Specific Language 3) return a prepared SQL statement against a given SQL database that queries an OMOP CDM database for the desired patient population(s) of interest.
-% EXPLAIN: What is a prepared statement?
-% EXPLAIN: What is a cohort in this context?
+\cite{evansOHDSICohortExpressionsJl2023}
+% TODO::MEDIUM Turn all items into complete thoughts in list
+How this tool works is to:
+
+\begin{enumerate}
+
+\item read and verify the serialized computable phenotype definition
+
+\item reinterpret the serialization into the FunSQL.jl Domain Specific Language
+
+\item return a prepared SQL statement against a given SQL database that queries an OMOP CDM database for the desired patient population(s) of interest.
+
+\end{enumerate}
+
+% EXPLAIN::MEDIUM What is a prepared statement?
+% EXPLAIN::MEDIUM What is a cohort in this context?
 Once these prepared statements are made, researchers can execute the generated SQL against their OMOP CDM instance to generate cohorts for future analyses.
@@
 
-% TODO: Reword this sentence
+% TODO::EASY Reword this sentence
 Although OHDSICohortExpressions.jl is sufficient to translate existing phenotype definitions generated in ATLAS to executable queries, oftentimes, there is a tricky problem of how to best iterate on phenotype definitions.
 \cite{zelkoDevelopingRobustComputable2023}
-% EXPLAIN: Better contextualize why this is important -- why not just create a new phenotype definition?
+% EXPLAIN:: Better contextualize why this is important -- why not just create a new phenotype definition?
 Often, one has to take an existing definition and either revise to a more narrow cohort or write manual code to build more niche cohorts after an initial definition is created.
-% EXPLAIN: How so?
-% TODO: Review this sentence; maybe cut it
+% EXPLAIN::MEDIUM How so?
+% TODO::MEDIUM Review this sentence; maybe cut it
 This refactoring of a phenotype definition can lead to conflating ideas within conversation between expert clinical reviewers, analysts, and health researchers.
-% TODO: Move this to the previous section on phenotype definitions
+% TODO::EASY Move this to the previous section on phenotype definitions
 This broadly can be the notions of what we are exploring (the phenotype definition; often the responsibility of a clinical reviewer), what is possible to compute upon (the computable phenotype definition; often the responsibility of an analyst), and the goals of the overall study.
 
-% TODO: Present this more as a general introduction to what tools are here for this
-% TODO: Talk about composition with DataFrames or Tidier
+% TODO::EASY Present this more as a general introduction to what tools are here for this
+% TODO::EASY Talk about composition with DataFrames or Tidier
 OMOPCDMCohortCreator.jl allows one to iteratively build a patient population quickly and easily.
 One can use patient pools developed from pre-existing cohorts defined off of phenotype definitions as a starting point in exploring patients or create new patient populations quickly without the need for a well-defined phenotype definition.
-% TODO: Mention the ability to satisfy research requests
+% TODO::MEDIUM Mention the ability to satisfy research requests
 This can enable analysts to more quickly understand what is inside their data assets, do "quick" analysis of a particular smaller research question, and enable a more distinct separation of work between a clinical expert reviewer and analyst teams.
-
-% TODO: Mention capr
 
@@
 
-% TODO: Mention JuliaCon Birds of a Feather discussion
-One of the biggest styming factors within not only the Julia ecosystem, but observational health research in general has been the difficulty of working with different databases.
+One of the biggest styming factors within not only the Julia ecosystem, but observational health research in general has been the difficulty of working with different databases.\cite{zelko2023julia}
 Although the OMOP CDM has been great for standardizing \textit{how} real world data is handled, there is a lack of consensus on \textit{what} is needed to handle these data assets.
-% TODO: Make this less verbose
+% TODO::EASY Make this less verbose
 For a variety of reasons, different groups will choose different database architectures (perhaps PostgreSQL for improved permission support, DuckDB for performance, or even SQLite for ease of use) but all are subtly different enough to create headaches in seamlessly working across different research sites.
 
-% TODO: Change the phrasing of this
+% TODO::EASY Change the phrasing of this
 The way that the Julia ecosystem handles this is via the creation of a common database interface called the DBInterface.
-% EXPLAIN: What is dispatch
+% EXPLAIN:: What is dispatch
 Specific Julia database packages can implement dispatches for this interace allowing users to have a much better experience working across databases.
@@
 
-% TODO: Better introduce this
+% TODO::MEDIUM Better introduce this
 For that reason, DBConnector.jl was created.
-% EXPLAIN: Contextualize this a bit better
+% EXPLAIN::MEDIUM Contextualize this a bit better
 Although this is not an explicit JuliaHealth package, it was a pain point that was experienced throughout the process of observational health research across multiple partner sites.
@@
 
-% TODO: Combine with the previous paragraph and state that there are two approaches
-% TODO: Make explicit DBConnector's path
-% TODO: Make explicit the path of choosing general packages
+% TODO::MEDIUM Combine with the previous paragraph and state that there are two approaches
+% TODO::EASY Make explicit DBConnector's path
+% TODO::EASY Make explicit the path of choosing general packages
 Finally, if someone needs more fine-grained control over a particular database connection, they can choose to eschew this package.
@@
 
-% TODO: Review if this section is actually needed
+% TODO::HARD Review if this section is actually needed
 \subsection{Ecosystem Interoperability}
@@
 
-% TODO: Reference code repository
+% TODO::EASY Reference code repository
 Using the JuliaHealth tools developed, we enumerate an example observational health study workflow that would be used in practice.\footnotemark
-For this paper, we used two datasets: a synthetic patient dataset called \textit{Eunomia} and the MIMIC III patient dataset which had been converted into the OMOP CDM to illustrate the workflow provided by the Observational Health sub-ecosystem.
+% EXPLAIN::EASY In enumeration talk about what Eunomia is 
+% EXPLAIN::EASY In enumeration discuss MIMIC III
+
+For this paper, we used two datasets: a synthetic patient dataset called \textit{Eunomia} and the MIMIC III patient dataset which had been converted into the OMOP CDM to illustrate the workflow provided by the Observational Health sub-ecosystem.\cite{physionet, hripcsakCharacterizingTreatmentPathways2016, johnson2016mimic, johnson2018mimic}
+
+\begin{enumerate}
+
+\item \textbf{Eunomia} -
+
+\item \textbf{MIMIC III} -
+
+\end{enumerate}
+
+% TODO::EASY Mention in footnote explicitly that reproducing workflow in Eunomia is possible
+% TODO::EASY Mention in footnote that explicitly reproducing workflow in MIMIC III is possible but not readily accessible
 Reproducing the workflow using the \textit{Eunomia} dataset is made readily available at the repository here: \cite{schuemieEunomia2023}
 
-% TODO: Add link to code repository
+% TODO::EASY Add link to code repository
 \footnotetext{The code snippets presented in the body of the paper may omit certain intermediate steps. 
@@
 \subsection{Connecting to a Database}
+
+% EXPLAIN::EASY What is this doing here?
+% TODO::HARD Update DBConnector
+% TODO::MEDIUM Maybe remove package imports and mention them within the caption instead
 
@@
 
+% TODO::HARD Review if this subsection is needed
 \subsection{Bridging between OHDSI and JuliaHealth Communities}
@@
 
+% EXPLAIN:: What is this doing here?
 After having a strep throat phenotype definition defined using ATLAS, we can then execute the resulting computable phenotype definition against the \textit{Eunomia} database.
 
+% TODO::HARD Maybe remove package imports and mention them within the caption instead
+\begin{listing}[!ht]
 \begin{minted}[breaklines,escapeinside=||,mathescape=true, numbersep=3pt, gobble=2, frame=lines, fontsize=\scriptsize, framesep=2mm]{julia}
@@
 \end{minted}
+% TODO::EASY Add a caption
+\caption{}
+% TODO::EASY Add a label
+\label{listing:}
+\end{listing}
 
+% TODO::EASY Move this to a caption
 A quick query of the cohort table (where the resulting subjects were stored) shows $\approx 1600$ from the dataset match this criteria.
@@
 
+% EXPLAIN:: What is this section doing?
+
+% TODO::EASY Maybe remove package imports and mention them within the caption instead
+\begin{listing}[!ht]
 \begin{minted}[breaklines,escapeinside=||,mathescape=true, numbersep=3pt, gobble=2, frame=lines, fontsize=\scriptsize, framesep=2mm]{julia}
@@
 \end{minted}
+% TODO::EASY Add a caption
+\caption{}
+% TODO::EASY Add a label
+\label{listing:}
+\end{listing}
 
@@
 
+% EXPLAIN::MEDIUM What is this section doing?
+
+% TODO::EASY Maybe remove package imports and mention them within the caption instead
+
+\begin{listing}[!ht]
 \begin{minted}[breaklines,escapeinside=||,mathescape=true, numbersep=3pt, gobble=2, frame=lines, fontsize=\scriptsize, framesep=2mm]{julia}
@@
 \end{minted}
+% TODO::EASY Add a caption
+\caption{}
+% TODO::EASY Add a label
+\label{listing:}
+\end{listing}
 
 Using this analytical approach, tabulation can be quickly conducted across patients.
+
+% EXPLAIN::EASY Move this to the caption of the table
 Additionally, throughout the process, one can perform relevant analysis methods.
 Here, we'll conduct a crude prevalence calculation across the patient cohort.
+
+% EXPLAIN::MEDIUM What is this section doing?
 
@@
     \end{tabular}
+    % TODO::EASY Add a caption
+    \caption{}
+    % TODO::EASY Add a label
+    \label{table:}
 \end{table}
@@
 
-As seen in Table 1 and Table 2, informative results can be rapidly generated by this workflow relevant to observational health research.
+% TODO::HARD Create table for MIMIC III results
+% TODO::EASY Create reference to Eunomia table
+As seen in Table 1, informative results can be rapidly generated by this workflow relevant to observational health research.
 Although the values shown in Table 2 are synthetic and the crude prevalence values themselves are meaningless, we can take a this approach and apply it directly to the MIMIC III dataset.
@@
 
-    1. Modularization of code and chaining
-    2. Distributed computing
+% EXPLAIN::MEDIUM What is the overview of this section?
+% EXPLAIN::MEDIUM What does each topic do?
+\begin{enumerate}
 
+\item Modularization of code and chaining
+
+\item Distributed computing
+
+\item Automatic generation of SQL from health packages
+
+\end{enumerate}
+
+% TODO::EASY Rename section
 \subsubsection{Modularization of Code and Chaining}
 
+% TODO::MEDIUM Make this more explicit about utilizing Julia's composition
 An exciting emergent aspect of the JuliaHealth ecosystem that came about as a result of developing OMOPCDMCohortCreator.jl was the prospect of radical modularization and chaining of package functionalities together.
+% TODO::MEDIUM Tie into currying more directly
+% TODO::MEDIUM Check if I am mentioning currying correctly
+We can leverage a unique property of Julia in that Julia supports partial evaluation of functions (also known as currying) to simplify repeated function calls to the same function over and over again.
+% TODO::MEDIUM Rewrite this part in light of the previous sections 
 For example, taking our workflow example from the earlier section where we characterized patients by race, gender, and age group, we can do a few "tricks" to modularize these characterization functions as follows:
 
-We can leverage a unique property of Julia in that Julia supports partial evaluation of functions (also known as currying) to simplify repeated function calls to the same function over and over again.
+% TODO::EASY Rework this section further
 First, we'll take our original functions and fix a connection object to each of our cofactor functions.
@@
 
+% TODO::MEDIUM Maybe move this into the tutorial code
+\begin{listing}[!ht]
 \begin{minted}[breaklines,escapeinside=||,mathescape=true, numbersep=3pt, gobble=2, frame=lines, fontsize=\scriptsize, framesep=2mm]{julia}
@@
 \end{minted}
+% TODO::EASY Add a caption
+\caption{}
+% TODO::EASY Add a label
+\label{listing:}
+\end{listing}
 
+% TODO::EASY Add this to previous code caption
 NOTE: The "C" prefix for each of our curried functions denotes that it is curried.
 
+% EXPLAIN::EASY Refactor this paragraph better or put most of it into the caption for the next code block
 Second, as OMOPCDMCohortCreator.jl is built on top of DataFrames.jl and one has the option of creating a DataFrame with the result from every OCC function, we can use the package, Chain.jl.
@@
 
+% TODO::MEDIUM Do we even need chaining here or could we use direct composition of functions?
+\begin{listing}[!ht]
 \begin{minted}[breaklines,escapeinside=||,mathescape=true, numbersep=3pt, gobble=2, frame=lines, fontsize=\scriptsize, framesep=2mm]{julia}
@@
 \end{minted}
+% TODO::EASY Add a caption
+\caption{}
+% TODO::EASY Add a label
+\label{listing:}
+\end{listing}
 
+% TODO::MEDIUM Can this be simplified?
+\begin{listing}[!ht]
 \begin{minted}[breaklines,escapeinside=||,mathescape=true, numbersep=3pt, gobble=2, frame=lines, fontsize=\scriptsize, framesep=2mm]{julia}
@@
 \end{minted}
+% TODO::EASY Add a caption
+\caption{}
+% TODO::EASY Add a label
+\label{listing:}
+\end{listing}
 
 In this way, future researchers using tools such as OMOPCDMCohortCreator.jl can create these more convenient abstractions.
+% TODO::EASY Move this into the future discussion section
 In fact, it could be a point of exploration in the future to build on top of OMOPCDMCohortCreator.jl even more abstracted and reusable functionalities to simplify ease of use while also enabling more readily auditable functionality.
+
+% TODO::MEDIUM Move this into its own section
 Finally, as these functions have now been fully modularized, these can now be readily distributed using Julia's distributed computing methods:
 
+% TODO::HARD Redo example; I do not feel it is good
+\begin{listing}[!ht]
 \begin{minted}[breaklines,escapeinside=||,mathescape=true, numbersep=3pt, gobble=2, frame=lines, fontsize=\scriptsize, framesep=2mm]{julia}
@@
 
+% TODO::EASY Rephrase this sentence
 One crucial aspect of the OHS is that it is unavoidable to work with databases that contain patient health information.
+% TODO::EASY Rephrase this sentence better
+% TODO::EASY Cite ANSI SQL
 For better or worse, the tool that is used to best interact with these databases are SQL dialects that loosely follow consistently the ANSI SQL guidance.
+% EXPLAIN::MEDIUM Highlight this more
+% EXPLAIN::MEDIUM In what scenarios might using Julia may not be possible
 To accomodate for this, several of the core tools developed in this sub-ecosystem generate SQL that can be used in the case where Julia may not be available in a given environment housing patient health data.
+% EXPLAIN::MEDIUM Introduce this better
 As an example, \textit{OMOPCDMCohortCreator.jl} functions build upon \textit{FunSQL.jl} allowing one to use a function such `GetPatientAgeGroup` without a connection object which would result in SQL.
 
+\begin{listing}[!ht]
 \begin{minted}[breaklines,escapeinside=||,mathescape=true, numbersep=3pt, gobble=2, frame=lines, fontsize=\scriptsize, framesep=2mm]{julia}
@@
 \end{minted}
+% TODO::EASY Add a caption
+\caption{}
+% TODO::EASY Add a label
+\label{listing:}
+\end{listing}
 
+\begin{listing}[!ht]
 \begin{minted}[breaklines,escapeinside=||,mathescape=true, numbersep=3pt, gobble=2, frame=lines, fontsize=\scriptsize, framesep=2mm]{julia}
@@
 \end{minted}
+% TODO::EASY Add a caption
+\caption{}
+% TODO::EASY Add a label
+\label{listing:}
+\end{listing}
 
+% TODO::MEDIUM Combine this section with the future directions section
 \subsection{Strong Composition within JuliaHealth}
@@
 
-As shown in this work, there is a rich opportunity for the JuliaHealth organization to become a strong presence within observational health research. 
+% TODO::EASY Rephrase this introductory paragraph
 JuliaHealth is now sufficiently at a stage of maturity where potential future researchers can leverage JuliaHealth's tools to conduct further future research or build upon existing architecture to target specific needs.
@@
 
+% TODO::MEDIUM This is where to introduce the paradox of composition
 Due to the strong composition that exists within the JuliaHealth and broader Julia ecosystem, solutions may already exist for solving various health informatics related research.
@@
 
+% TODO::MEDIUM Combine this paragraph with the previous paragraph
+% TODO::EASY Shorten this sentiment
 Additionally, due to the composable and modular nature of the OHS, research analyses and visualizations can be decoupled and interchanged with one another.
@@
 
+% TODO::EASY Make this more about the flexible nature of the JuliaHealth organization
 Furthermore, again exploiting the composition within the JuliaHealth ecosystem, as mentioned in the context of JuliaHealth and machine learning applications, future efforts could focus on developing fairness and bias auditing tools in patient datasets.
 These tools could consider factors commonly explored in health equity literature like data completeness, demographic variables, and fairness algorithms.
-
-Moreover, there's potential to advance clinical informatics by developing a flexible clinical type system, supporting computable phenotype definitions and medical ontologies.
-The ecosystem can also benefit from tools to audit the fairness of phenotype definitions.
-Finally, as the landscape of data expands, integrating non-traditional sources like social media and environmental data into standardized patient records will be a critical frontier.
-In these ways, JuliaHealth is primed to drive innovation in observational health informatics, pushing the boundaries of data-driven healthcare research.
 
@@
 Although it is an ecosystem that is relatively new, its robustness allows for novel investigations to be pursued within this domain of research. 
+% TODO::MEDIUM Mention how this allows JuliaHealth to solve more problems than what is present within JuliaHealth
 Additionally, due to the strongly compositional nature of the Julia ecosystem, these JuliaHealth tools can readily compose with tooling to support in-depth statistical analyses, mathematical modeling, and visualization capacities.
 
-The deep dive into the JuliaHealth ecosystem unravels its arsenal of tools tailored for observational health research, meticulously designed to navigate the complexities of handling OMOP CDM data.
-Notably, the sub-ecosystem within JuliaHealth presents mature, internationally federated research capabilities, building upon existing innovations like OHDSI's HADES, while offering novel tools such as OMOPCDMCohortCreator.jl.
-In considering future directions, the paper posits the promising avenues for JuliaHealth, envisioning expanded documentation, the development of visualization tools, and integration of machine learning applications for fairness auditing.
-The ecosystem's potential to revolutionize clinical informatics by integrating non-traditional data sources, auditing phenotype definitions, and fostering innovation in data-driven healthcare research is paramount.
+% TODO::MEDIUM Rework this
 In essence, the JuliaHealth ecosystem stands as a catalyst for transformative observational health research, poised to push the boundaries of innovation, collaboration, and standardization in the pursuit of better healthcare outcomes.
@@
 \section{References}
+
+% TODO::HARD Review and probably remove appendix
 

### 2023-12-28 header.tex
New title details

@@
 
-\title{Developing Observational Health Research Capacity in the Julia Ecosystem}
+\title{Introducing the JuliaHealth Observational Health Research Subecosystem}
 
@@
 \hypersetup{
-pdftitle = {Developing Observational Health Research Capacity in the Julia Ecosystem},
+pdftitle = {Introducing the JuliaHealth Observational Health Research Subecosystem},
 pdfsubject = {JuliaCon 2019 Proceedings},

### 2023-12-30 paper.tex
Added some figure captions, enumerations, and formatting options

@@
 \documentclass{juliacon}
+
+\usepackage[labelfont=bf]{caption}
+\captionsetup[table]{skip=4pt}
+\captionsetup{font=scriptsize}
 
@@
 Using the JuliaHealth tools developed, we enumerate an example observational health study workflow that would be used in practice.\footnotemark
-% EXPLAIN::EASY In enumeration talk about what Eunomia is 
-% EXPLAIN::EASY In enumeration discuss MIMIC III
 
-For this paper, we used two datasets: a synthetic patient dataset called \textit{Eunomia} and the MIMIC III patient dataset which had been converted into the OMOP CDM to illustrate the workflow provided by the Observational Health sub-ecosystem.\cite{physionet, hripcsakCharacterizingTreatmentPathways2016, johnson2016mimic, johnson2018mimic}
+For this paper, we used two datasets:
 
@@
 
-\item \textbf{Eunomia} -
+% TODO::EASY Cite https://github.com/synthetichealth/synthea
+\item \textbf{Eunomia} - this is a synthetic patient dataset created from the open source synthetic patient generator, \textit{Synthea}, and hosted within the HADES and JuliaHealth ecosystems.
 
-\item \textbf{MIMIC III} -
+\item \textbf{MIMIC III} - a deidentified datasets consisting of inpatient visits from Beth Israel Deaconness Hospital in Boston, Massachussetts.
+The particular version of the dataset that we used was transformed into the OMOP CDM using tools from MIT LCP and PhysioNet.
+\cite{physionet, hripcsakCharacterizingTreatmentPathways2016, johnson2016mimic, johnson2018mimic}
 
@@
 % TODO::HARD Update DBConnector
-% TODO::MEDIUM Maybe remove package imports and mention them within the caption instead
-
 \begin{listing}[!ht]
@@
   Eunomia
-import OMOPCDMCohortCreator as OCC
 
@@
 \end{minted}
-\caption{This illustrates how to connect to the Eunomia database}
+\caption{\textbf{Configuring OMOP CDM Database Connection.} Using \textit{DBConnector.jl}, connection to the \textit{Eunomia} SQLite database can be made.
+Then, \textit{OMOPCDMCohortCreator.jl} (occ) is used to generate connection details used internally by the package for the rest of the session.}
 \label{listing:connection}
@@
 \begin{minted}[breaklines,escapeinside=||,mathescape=true, numbersep=3pt, gobble=2, frame=lines, fontsize=\scriptsize, framesep=2mm]{julia}
-using JSON3 
-using OHDSICohortExpressions
-
-import DBInterface as DBI
-
-cohort = JSON3.read("strep_throat.json")
-cohort_expression = cohort[:items][1][:expression]
-
-model = Model(cdm_version = v"5.3.1", 
+model = oce.Model(cdm_version = v"5.3.1", 
               cdm_schema = "main",
@@
 
-sql = translate(cohort_expression, 
+sql = oce.translate(cohort_expression, 
                 dialect = :sqlite, 
@@
                 cohort_definition_id = 1);
-
-for query in split(sql, ";")[1:end-1]
-  DBI.execute(conn, query)
-end
 \end{minted}
-% TODO::EASY Add a caption
-\caption{}
-% TODO::EASY Add a label
-\label{listing:}
+\caption{\textbf{Creating a Cohort Using a Computable Phenotype Definition.} \textit{OHDSICohortExpressions.jl} (oce) ingests a computable phenotype definition to generate a cohort of patients.
+The JSON is translated into an intermediate SQL Syntax Tree, instantiated in SQL (SQLite) targeting an OMOP CDM v5.3.1 database, and then populates the database's cohort table.}
+\label{listing:cohort_creation}
 \end{listing}
@@
 
-% TODO::EASY Maybe remove package imports and mention them within the caption instead
 \begin{listing}[!ht]
 \begin{minted}[breaklines,escapeinside=||,mathescape=true, numbersep=3pt, gobble=2, frame=lines, fontsize=\scriptsize, framesep=2mm]{julia}
-C = GetCohortSubjects(1, conn)
-rename!(C, :subject_id => :person_id)
+patients = occ.GetCohortSubjects(1, conn)
 
-C_race = occ.GetPatientRace(C.person_id, conn)
-C_gender = occ.GetPatientGender(C.person_id, conn)
+patients_race = occ.GetPatientRace(C.person_id, conn)
+patients_gender = occ.GetPatientGender(C.person_id, conn)
 
-C_age_group = occ.GetPatientAgeGroup(
+patients_age_group = occ.GetPatientAgeGroup(
   C.person_id, 
@@
 \end{minted}
-% TODO::EASY Add a caption
-\caption{}
-% TODO::EASY Add a label
-\label{listing:}
+\caption{\textbf{Find Demographic Characteristics of Cohort.} Demographic properties of a cohort can be quickly queried from a database.}
+\label{listing:demographics}
 \end{listing}
@@
 
-% TODO::EASY Maybe remove package imports and mention them within the caption instead
-
 \begin{listing}[!ht]
 \begin{minted}[breaklines,escapeinside=||,mathescape=true, numbersep=3pt, gobble=2, frame=lines, fontsize=\scriptsize, framesep=2mm]{julia}
-import DataFrames as DF
+strep_df = df.combine(patient_groups, df.nrow => :count)
 
-C_features = DF.outerjoin(C_race,
-                          C_gender,
-                          C_age_group;
-                          on = :person_id, 
-                          matchmissing = :equal)
+# Creator code to process data and generate counts...
 
-C_features = C_features[:, DF.Not(:person_id)]
-
-C_groups = DF.groupby(C_features, 
-                      [
-                        :race_concept_id, 
-                        :gender_concept_id, 
-                        :age_group
-                      ])
-
-C_groups = DF.combine(C_groups, DF.nrow => :count)
+strep_df.prev = strep_df.count ./ strep_df.total_count 
 \end{minted}
-% TODO::EASY Add a caption
-\caption{}
-% TODO::EASY Add a label
-\label{listing:}
+\caption{\textbf{Calculating crude prevalence rates.} Once demographic information is extracted, processing can be performed using tools such as \textit{DataFrames.jl} (df) to calculate metrics such as prevalence rates in a very straightforward manner.}
+\label{listing:prevalence}
 \end{listing}
@@
 Using this analytical approach, tabulation can be quickly conducted across patients.
-
-% EXPLAIN::EASY Move this to the caption of the table
-Additionally, throughout the process, one can perform relevant analysis methods.
-Here, we'll conduct a crude prevalence calculation across the patient cohort.
 
@@
     % TODO::EASY Add a caption
-    \caption{}
-    % TODO::EASY Add a label
-    \label{table:}
+    \caption{\textbf{Example of a Strep Throat Prevalence Calculation.} This illustrates the sort of results one can quickly generate using the OHS.
+Note, that although the prevalence values in this table are based on synthetic data and are not realistic, the output looks very similar to what one could expect to generate.}
+    \label{table:eunomia_prevalence}
 \end{table}
@@
 CharacterizePatients(patients) = @chain patients begin
-    _[:, DF.Not(:person_id)]
-    DF.groupby(_, names(_))
-    DF.combine(_, DF.nrow => :count)
+    _[:, df.Not(:person_id)]
+    df.groupby(_, names(_))
+    df.combine(_, df.nrow => :count)
     occ.ExecuteAudit

### 2023-12-31 paper.tex
Wrapped up some more todos

@@
 
-% TODO::MEDIUM Maybe move this into the tutorial code
-\begin{listing}[!ht]
-\begin{minted}[breaklines,escapeinside=||,mathescape=true, numbersep=3pt, gobble=2, frame=lines, fontsize=\scriptsize, framesep=2mm]{julia}
-import Base:
-  Fix2
-
-import OMOPCDMCohortCreator:
-  GetPatientRace,
-  GetPatientGender,
-  GetPatientAgeGroup
-
-CGetPatientRace = Fix2(GetPatientRace, c)
-CGetPatientGender = Fix2(GetPatientGender, c)
-CGetPatientAgeGroup = Fix2(GetPatientAgeGroup, c)
-\end{minted}
-% TODO::EASY Add a caption
-\caption{}
-% TODO::EASY Add a label
-\label{listing:}
-\end{listing}
-
-% TODO::EASY Add this to previous code caption
-NOTE: The "C" prefix for each of our curried functions denotes that it is curried.
-
 % EXPLAIN::EASY Refactor this paragraph better or put most of it into the caption for the next code block
@@
 
-% TODO::MEDIUM Do we even need chaining here or could we use direct composition of functions?
 \begin{listing}[!ht]
 \begin{minted}[breaklines,escapeinside=||,mathescape=true, numbersep=3pt, gobble=2, frame=lines, fontsize=\scriptsize, framesep=2mm]{julia}
-import Chain:
-  @chain
+FGetPatientRace = Fix2(occ.GetPatientRace, conn)
+FGetPatientGender = Fix2(occ.GetPatientGender, conn)
+FGetPatientAgeGroup = Fix2(occ.GetPatientAgeGroup, conn)
 
 CofactorPatients(patients) = @chain patients begin
-    CGetPatientRace_fixed
-    CGetPatientGender_fixed
-    CGetPatientAgeGroup_fixed
+    FGetPatientRace
+    FGetPatientGender
+    FGetPatientAgeGroup
 end
-\end{minted}
-% TODO::EASY Add a caption
-\caption{}
-% TODO::EASY Add a label
-\label{listing:}
-\end{listing}
 
-% TODO::MEDIUM Can this be simplified?
-\begin{listing}[!ht]
-\begin{minted}[breaklines,escapeinside=||,mathescape=true, numbersep=3pt, gobble=2, frame=lines, fontsize=\scriptsize, framesep=2mm]{julia}
 CharacterizePatients(patients) = @chain patients begin
@@
 end
-\end{minted}
-% TODO::EASY Add a caption
-\caption{}
-% TODO::EASY Add a label
-\label{listing:}
-\end{listing}
 
-This gives one the ability to simply reason about an analysis in a very straightforward and readable manner:
-
-\begin{listing}[!ht]
-\begin{minted}[breaklines,escapeinside=||,mathescape=true, numbersep=3pt, gobble=2, frame=lines, fontsize=\scriptsize, framesep=2mm]{julia}
 RunStudy(patients) = @chain patients begin
@@
 \end{minted}
-% TODO::EASY Add a caption
-\caption{}
-% TODO::EASY Add a label
-\label{listing:}
+\caption{\textbf{Modularizing Study Code.} Using Julia's partial application abilities, connection objects can be fixed allowing study code to be readily modularized.
+Using the package, Chain.jl, occ and processing functions can be composed together to create modular study code.}
+\label{listing:modular_study}
 \end{listing}
@@
 \end{minted}
-% TODO::EASY Add a caption
-\caption{}
-% TODO::EASY Add a label
-\label{listing:}
+\caption{\textbf{Distributed Computation of a Study.} With occ, study code can be easily abstracted and modularized.
+Using Julia's native Distributed package, each module of a study can then be parallelized over patient cohorts.}
+\label{listing:study_parallelization}
 \end{listing}
@@
 \begin{listing}[!ht]
-\begin{minted}[breaklines,escapeinside=||,mathescape=true, numbersep=3pt, gobble=2, frame=lines, fontsize=\scriptsize, framesep=2mm]{julia}
-  GetPatientAgeGroup([1]; age_groupings = [[0, 40], [41, 80]])
+\begin{minted}[breaklines,escapeinside=||,mathescape=true, numbersep=3pt, gobble=2, frame=topline, fontsize=\scriptsize, framesep=2mm]{julia}
+# Julia Input
+occ.GetPatientAgeGroup([1]; age_groupings = [
+                                              [0, 40], 
+                                              [41, 80]
+                                            ])
+
 \end{minted}
-% TODO::EASY Add a caption
-\caption{}
-% TODO::EASY Add a label
-\label{listing:}
-\end{listing}
 
-\begin{listing}[!ht]
-\begin{minted}[breaklines,escapeinside=||,mathescape=true, numbersep=3pt, gobble=2, frame=lines, fontsize=\scriptsize, framesep=2mm]{julia}
+\begin{minted}[breaklines,escapeinside=||,mathescape=true, numbersep=3pt, gobble=2, frame=bottomline, fontsize=\scriptsize, framesep=2mm]{sql}
+# SQL Output
 SELECT
@@
 \end{minted}
-% TODO::EASY Add a caption
-\caption{}
-% TODO::EASY Add a label
-\label{listing:}
+\caption{\textbf{Producing SQL from Julia Expression.} When occ functions are not passed a connection object, they can produce SQL representing the underyling query the Julia expression is executing.}
+\label{listing:julia_sql}
 \end{listing}
@@
 
-\section{References}
-
-% TODO::HARD Review and probably remove appendix
-
-% APPENDIX
-\appendix
-
-\section{JSON Serialization of Computable Phenotype Definition}\label{appendix:json}
-
-An example of the JSON serialization of computable phenotype definitions can be found within the paper repository here:
-
-In particular, this example seeks to identify individuals with Type 2 Diabetes Mellitus within a given OMOP CDM.
-
-This was generated using ATLAS Version 2.12.1 as well as 
-WebAPI Version 2.12.1.
-A generic definition of the schema that was created for ATLAS 2.7.1 can be explored here: https://github.com/OHDSIBr/ATLAS-JSON-Schema/tree/2.7.4
-
-\section{PostgreSQL Serialization of Computable Phenotype Definition}\label{appendix:postgresql}
-
-An example of a PostgreSQL SQL serialization of computable phenotype definitions can be found within the paper repository here:
-
-In particular, this example seeks to identify individuals with Type 2 Diabetes Mellitus within a given OMOP CDM.
-
-This was generated using ATLAS Version 2.12.1 as well as 
-WebAPI Version 2.12.1.
-
 \input{bib.tex}

### 2023-12-31 header.tex
Renamed paper and added author information

@@
 
-\title{Introducing the JuliaHealth Observational Health Research Subecosystem}
+\title{The JuliaHealth Observational Health Research Subecosystem}
 
-\author[1]{Jacob S. Zelko}
-\author[3]{Fareeda Abdelazeez}
+\author[5]{Jacob S. Zelko}
 \author[1, 2]{Malina Hy}
 \author[1, 2]{Varshini Chinta}
-\author[3]{Jay Sanjay}
+\author[3]{Fareeda Abdelazeez}
+\author[4]{Jay Sanjay}
 \affil[1]{Georgia Tech Research Institute}
 \affil[2]{Georgia Institute of Technology}
-\affil[3]{Foobar}
+\affil[3]{Zewail University}
+\affil[4]{Foobar}
+\affil[5]{Northeastern University}
 
@@
 \hypersetup{
-pdftitle = {Introducing the JuliaHealth Observational Health Research Subecosystem},
+pdftitle = {The JuliaHealth Observational Health Research Subecosystem},
 pdfsubject = {JuliaCon 2019 Proceedings},
-pdfauthor = {Jacob S. Zelko, Fareeda Abdelazeez, Malina Hy, Varshini Chinta, Jay Sanjay},
+pdfauthor = {Jacob S. Zelko, Malina Hy, Varshini Chinta, Fareeda Abdelazeez, Jay Sanjay},
 pdfkeywords = {Observational Health, OMOP Common Data Model, Database Management, Characterization},

### 2024-01-01 paper.tex
Fixed up some more todos

@@
 \footnotetext{The code snippets presented in the body of the paper may omit certain intermediate steps. 
-Please review the associated code repository for a full working tutorial.}
+Please review the associated code repository for a full working tutorial located here: https://github.com/TheCedarPrince/2022-JuliaCon-Proceedings-Submission}
 
@@
 
-% TODO::EASY Move this to a caption
-A quick query of the cohort table (where the resulting subjects were stored) shows $\approx 1600$ from the dataset match this criteria.
-
 \subsection{Characterizing Patient Populations}
@@
     \end{tabular}
-    % TODO::EASY Add a caption
     \caption{\textbf{Example of a Strep Throat Prevalence Calculation.} This illustrates the sort of results one can quickly generate using the OHS.
@@
 % TODO::HARD Create table for MIMIC III results
-% TODO::EASY Create reference to Eunomia table
-As seen in Table 1, informative results can be rapidly generated by this workflow relevant to observational health research.
+As seen in \ref{table:eunomia_prevalence}, informative results can be rapidly generated by this workflow relevant to observational health research.
 Although the values shown in Table 2 are synthetic and the crude prevalence values themselves are meaningless, we can take a this approach and apply it directly to the MIMIC III dataset.

### 2024-01-01 paper.tex
Fixed up some more TODOs

@@
 % EXPLAIN::EASY What is a network study?
-% TODO::EASY Cite the quote "write once, run everywhere" 
-% http://www.sun.com/smi/Press/sunflash/1996-01/sunflash.960123.10561.xml} 
-This has led to successful collaboration relationships across countries where one research group can develop a network study and then disseminate the study across collaborators leading to the idea that you can "write once, run everywhere."
+This has led to successful collaboration relationships across countries where one research group can develop a network study and then disseminate the study across collaborators leading to the idea that you can "write once, run everywhere."\footnotemark
+
+\footnotetext{This quote is attributed to Sun Microsystems when discussing the Java programming language.}
 
@@
 % TODO::EASY Create footnote about schema
-% TODO::EASY Add reference to schema in footnote
-% TODO::EASY Cite reference schema from GitHub in footnote; \cite{madrilATLASJSONSchema2022}
+% TODO::EASY Cite reference schema from GitHub in footnote; 
 % EXPLAIN::EASY Better introduce what SQL is
-The computable phenotype definitions that ATLAS generates based on phenotype definitions are in a variety of serialization formats such as JSON or prepared SQL expressions available in multiple SQL flavors.
+The computable phenotype definitions that ATLAS generates based on phenotype definitions are in a variety of serialization formats such as JSON or prepared SQL expressions available in multiple SQL flavors.\cite{madrilATLASJSONSchema2022}
 
@@
 
-% TODO::EASY Reference code repository
 Using the JuliaHealth tools developed, we enumerate an example observational health study workflow that would be used in practice.\footnotemark
@@
 
-% TODO::EASY Cite https://github.com/synthetichealth/synthea
 \item \textbf{Eunomia} - this is a synthetic patient dataset created from the open source synthetic patient generator, \textit{Synthea}, and hosted within the HADES and JuliaHealth ecosystems.
+\cite{walonoski2018synthea}
 
@@
 
-% TODO::EASY Mention in footnote explicitly that reproducing workflow in Eunomia is possible
-% TODO::EASY Mention in footnote that explicitly reproducing workflow in MIMIC III is possible but not readily accessible
 Reproducing the workflow using the \textit{Eunomia} dataset is made readily available at the repository here: \cite{schuemieEunomia2023}
 
-% TODO::EASY Add link to code repository
 \footnotetext{The code snippets presented in the body of the paper may omit certain intermediate steps. 
-Please review the associated code repository for a full working tutorial located here: https://github.com/TheCedarPrince/2022-JuliaCon-Proceedings-Submission}
+Please review the associated code repository for a full working tutorial located here: https://github.com/TheCedarPrince/2022-JuliaCon-Proceedings-Submission.
+As a further note, the code to fully reproduce a sample analysis with \textit{Eunomia} is available within the code repository. 
+As MIMIC III is not publicly available, a sample analysis using MIMIC III is available with some information stripped for privacy considerations.}
 
@@
 
-\item Modularization of code and chaining
+\item Modularization
 
-\item Distributed computing
+\item Distributed Computing
 
-\item Automatic generation of SQL from health packages
+\item Automatic SQL Generation
 
@@
 
-% TODO::EASY Rename section
-\subsubsection{Modularization of Code and Chaining}
+\subsubsection{Modularization}
 
@@
 
-\subsubsection{Auto-Generating SQL from JuliaHealth Commands}
+\subsubsection{Automatic SQL Generation}
 
@@
 % TODO::EASY Rephrase this sentence better
-% TODO::EASY Cite ANSI SQL
-For better or worse, the tool that is used to best interact with these databases are SQL dialects that loosely follow consistently the ANSI SQL guidance.
+For better or worse, the tool that is used to best interact with these databases are SQL dialects that loosely follow consistently the ANSI SQL guidance.\cite{kelechavaSQLStandardISO2018}
 % EXPLAIN::MEDIUM Highlight this more

### 2024-01-06 paper.tex
Resolving additional issues and citation details

@@
 The importance of this model has become apparent within open science organizations like the OHDSI (Observational Health Data Science and Informatics) network where members across $100+$ countries \cite{sachsonOurJourney2023} with access to various health systems have been able to successfully map their patient data to the OMOP CDM.
-% TODO::EASY Add citation about successful collaborations
 % EXPLAIN::EASY What is a network study?
-This has led to successful collaboration relationships across countries where one research group can develop a network study and then disseminate the study across collaborators leading to the idea that you can "write once, run everywhere."\footnotemark
+This has led to successful collaboration relationships \cite{roseFindingAnswersBig2008} across countries where one research group can develop a network study and then disseminate the study across collaborators leading to the idea that you can "write once, run everywhere."\footnotemark
 
@@
 % EXPLAIN::EASY What is a study package?
-% EXPLAIN::EASY What data assets?
-Once this packaging is done, they can pass along the study package to another site to collaborate on the same analysis but on different data assets.
+Once this packaging is done, they can pass along the study package to another site to collaborate on the same analysis but execute the study package on their unique OMOP CDM formatted dataset.
 % EXPLAIN::MEDIUM What is a collaborator network?
@@
 These phenotype definitions have varying levels of complexity where they can be very simple (such as finding all patients with one health condition or diagnosis) or grow to be extremely complex (defining controlled, uncontrolled, and indeterminate patient cohorts based on medications, lab results, etc.)
-% TODO::EASY Create footnote about schema
-% TODO::EASY Cite reference schema from GitHub in footnote; 
-% EXPLAIN::EASY Better introduce what SQL is
-The computable phenotype definitions that ATLAS generates based on phenotype definitions are in a variety of serialization formats such as JSON or prepared SQL expressions available in multiple SQL flavors.\cite{madrilATLASJSONSchema2022}
+The computable phenotype definitions that ATLAS generates based on phenotype definitions are in a variety of serialization formats such as JSON schemas or prepared SQL expressions available in multiple SQL flavors.\footnotemark
+
+\footnotetext{Although the JSON schema is not made formally available within OHDSI, a community made copy of the schema can be found here: \cite{madrilATLASJSONSchema2022}}
 
@@
 \cite{evansOHDSICohortExpressionsJl2023}
-% TODO::MEDIUM Turn all items into complete thoughts in list
 How this tool works is to:
@@
 
-\item read and verify the serialized computable phenotype definition
+\item Read and verify serialized computable phenotype definitions.
 
-\item reinterpret the serialization into the FunSQL.jl Domain Specific Language
+\item Reinterpret serialization into the FunSQL.jl Domain Specific Language.
 
-\item return a prepared SQL statement against a given SQL database that queries an OMOP CDM database for the desired patient population(s) of interest.
+\item Return a SQL statement for an OMOP CDM database to generate a given cohort.
 
@@
 Often, one has to take an existing definition and either revise to a more narrow cohort or write manual code to build more niche cohorts after an initial definition is created.
-% EXPLAIN::MEDIUM How so?
-% TODO::MEDIUM Review this sentence; maybe cut it
-This refactoring of a phenotype definition can lead to conflating ideas within conversation between expert clinical reviewers, analysts, and health researchers.
 % TODO::EASY Move this to the previous section on phenotype definitions
@@
 
-% TODO::HARD Review if this subsection is needed
-\subsection{Bridging between OHDSI and JuliaHealth Communities}
-
-Briding between tools used in the OHDSI and JuliaHealth communities is now exceptionally straightforward.
-
-% \includegraphics{./atlas_gui.png}
-
 \subsection{Using a Cohort Definition To Build an Initial Cohort}
@@
 
-% TODO::HARD Maybe remove package imports and mention them within the caption instead
 \begin{listing}[!ht]

### 2024-01-06 paper.tex
Finished introduction

@@
 
-% TODO::HARD Review the abstract after reviewing the paper fully
+% TODO::HARD Review the abstract after reviewing the paper fully/***//***//***//***//***//***/
 Observational health continues to be a growing field in health informatics research as electronic health records (EHR), patient medical claims, and other ancilliary patient data source become more readily computable and accessible to researchers.
@@
 
-% EXPLAIN::EASY What is observational health research?
-% EXPLAIN::MEDIUM What sort of innovations?
-Over the past $10$ years, there have been several innovations within the observational health research community.
-Observational health research is an area of study that leverages authentic health data to gain insights into the health system without interference.
-Data are collected by researchers as they follow groups of people over a specified period.
-Depending on the study, these groups may comprise healthy individuals, individuals with a specific disease, or those at high risk of developing a particular condition—such as individuals with a family history—while also taking into consideration the imperative of safeguarding their privacy.
-Observational studies can contribute to answering questions as what genes cause a specific disease to grow, patterns and trends for a particular disease, experiences of past patients, behaviours increasing the risk of developing a certain disease.
+% NOTE: I made this definition up on the fly using my expertise and familiarity with the space! :D /***//***//***//***//***//***/
+Observational health research is the study and analysis of historical real world data to explore various cohorts found within such datasets to generate real world evidence to support applications within health-related domains.
+Research within this space can investigate questions such as population health trends for a particular disease, health disparities within patient populations, and cost costs of care related to particular treatment procedures.
+Due to the untapped potential of this space, over the past $20$ years, there have been several developments within the observational health research community to enable more effective research to be conducted within this field (such as the creation of interoperability standards and tool creation \cite{benderHL7FHIRAgile2013; mandelSMARTFHIRStandardsbased2016; overhage2012validation}).
 
-% EXPLAIN::MEDIUM Why is this a crucial innovation?
-% EXPLAIN::MEDIUM What is meant by a transferable analysis?
-Arguably, the most crucial innovation that has emerged is the Observational Medical Outcomes Partnership Common Data Model (OMOP CDM) which was designed specifically to standardize observational health data to enable transferable analyses.
+Arguably, the most crucial innovation that has emerged in observational health research within the last decade is the development and adoption of the Observational Medical Outcomes Partnership Common Data Model (OMOP CDM) which was designed specifically to standardize real world health data.
 \cite{overhage2012validation}
-% EXPLAIN::MEDIUM Why is mapping relevant here?
-The importance of this model has become apparent within open science organizations like the OHDSI (Observational Health Data Science and Informatics) network where members across $100+$ countries \cite{sachsonOurJourney2023} with access to various health systems have been able to successfully map their patient data to the OMOP CDM.
-% EXPLAIN::EASY What is a network study?
-This has led to successful collaboration relationships \cite{roseFindingAnswersBig2008} across countries where one research group can develop a network study and then disseminate the study across collaborators leading to the idea that you can "write once, run everywhere."\footnotemark
+The importance of this model has become apparent within open science organizations like the OHDSI (Observational Health Data Science and Informatics) network where members across $100+$ countries \cite{sachsonOurJourney2023} with access to real world data have been able to successfully transform their datasets to the OMOP CDM.
+This has led to successful collaboration relationships \cite{roseFindingAnswersBig2008} across countries and institutions where one institutional research group can develop an analysis centered around the OMOP CDM.
+Then, the network of institutions who have adopted the OMOP CDM can package and share this analysis amongst themselves (this is colloquially known as a "network study" within the context of observational health research) to enable research on an international scale.
+The network study idea is particularly salient to this field as it leads to the idea that you can "write [a study] once [and] run [it] everywhere."\footnotemark
 
-\footnotetext{This quote is attributed to Sun Microsystems when discussing the Java programming language.}
+\footnotetext{This paraphrased quote is attributed to Sun Microsystems when discussing the Java programming language.}
 
@@
 
-Modern observational health research centers around a variety of core technologies.
-A few of these technologies highlighted in this paper are:
+Modern observational health research utilizes several different technologies.
 
-\begin{enumerate}
+\textit{Data standards within observational health research.} Although other standards exist within observational health, the OMOP CDM will be discussed and focused on for the remainder of this paper.
+To give a brief history of the OMOP CDM, OMOP was convened in 2009 by the US Food and Drug Administration to address concerns related to data type, study design, and data privacy concerns of utilizing "Real World Data" \cite{ohdsi2019book} \cite{FDARealWorldEvidence}
+Their collective efforts resulted in the OMOP Common Data Model (OMOP CDM) to house and ingest real world data that is used in observational health research.
+The OMOP CDM is a "person-centered" model (i.e.
+any data found in a database formatted to the OMOP CDM can be traced back to an individual patient) that standardizes patient events to a plethora of database tables to represent the unique dynamics and structures found in individual patient events (for more details on these relationships, please refer to \ref{fig:cdm}).
+These events would be any recorded interaction a patient has with a healthcare system and they can consist of situations such as a patient visiting an inpatient emergency room, a drug prescription, a patient condition diagnosis, medical procedures, and more.
 
-\item Data standards within observational health research
-\item Collaborative networks and studies using observational health data
-\item Research software to support the analysis of observational health data
+\begin{figure}[!ht]
+\centering
+\includegraphics[scale = 0.6]{omop_cdm.png}
+\caption{\textbf{High Level Relationship Schema of the OMOP CDM Version 5.4.} This is the schema that is prescribed by the OMOP CDM.
+It should be noted OMOP CDM itself does not require a specific technology to work with the data stored in this standard.}
+\label{fig:cdm}
+\end{figure}
 
-\end{enumerate}
+% TODO::HARD Develop this paragraph more about research software/***//***//***//***//***/
+\textit{Research software to support the analysis of observational health data.} As the OMOP CDM has become ubiquitous within observational health research, developing research software to handle and manage OMOP CDM data is necessary to meaningfully contribute within the observational health research space.
+OHDSI supports the development of open-source software tools in a centralized ecosystem called HADES (Health Analytics Data-to-Evidence Suite) for particularly managing and working with OMOP CDM formatted data.
+The HADES ecosystem has been in existence for $\approx 10$ years and has given rise to improvements in the quality and reliability of observational health research as well as motivating novel collaboration mechanisms.
 
-In particular, the data standard that will be most highlighted for the remainder of this paper is the Observational Medical Outcomes Partnership Common Data Model (OMOP CDM).
-Subsequent discussion will be focused on the opportunities that this specific data model has given rise to across observational health as well as research communities such as the Observational Health Data Sciences and Informatics (OHDSI) open science consortium.
-
-% TODO::MEDIUM This paragraph should be centered on the core technology of data standards -- in particular, the OMOP CDM
-% EXPLAIN::MEDIUM What is real world data? Use https://jacobzelko.com/10282021140730-real-world-evidence/
-OMOP was created in 2009 by the US Food and Drug Administration to address concerns related to data type, study design, and data privacy concerns of "Real World Data" \cite{ohdsi2019book} \cite{FDARealWorldEvidence}
-What could arguably be considered the longest standing achievement of OMOP was the development of the OMOP Common Data Model (OMOP CDM) to house and ingest real world data.
-% TODO::EASY Combine the next two sentences together
-% EXPLAIN::EASY What does "person-centered" mean?
-% EXPLAIN::MEDIUM What is a patient event? Could you give an example?
-in a consistent format that has been optimized to assist in extracting meaningful population health information.
-In brief, the OMOP Common Data Model (CDM) is a "person-centered" model that has standardized common events in a given patient encounter with database tables that represent the unique structures found in patient events.
-% TODO::EASY Develop this fragment into a full sentence
-prioritizing data protection through the limiting of information that could endanger patient anonymity.
-% EXPLAIN:: What is a health system? Could you give an example?
-This creates a longitudinal view of all healthcare-related events per person recorded in a singular health system.
-% TODO::EASY Make this more coherent to the rest of the paragraph
-% TODO::EASY Explicitly mention that although the model prescribes a database schema, it does not need a specific database type
-The OMOP CDM itself does not require a specific technology to work with the data stored in this standard.
-
-% TODO::MEDIUM This paragraph should be centered on the core technology of the concept of network studies
-% TODO::MEDIUM Use this to address the notion of what "network studies" are
-% TODO::MEDIUM Make this initially less OHDSI specific; pull from https://jacobzelko.com/05122023034151-network-macarthur
-In particular, one collaboration mechanism that OHDSI has been able to operationalize meaningfully while working with observational health data is the notion of "network studies".
+% TODO::MEDIUM This paragraph should be centered on the core technology of the concept of network studies/***//***//***//***//***/
+% TODO::MEDIUM Use this to address the notion of what "network studies" are/***//***//***//***//***/
+% TODO::MEDIUM Make this initially less OHDSI specific; pull from https://jacobzelko.com/05122023034151-network-macarthur/***//***//***//***//***/
+\textit{Collaborative networks and studies using observational health data.} In particular, one collaboration mechanism that OHDSI has been able to operationalize meaningfully while working with observational health data is the notion of "network studies".
 These studies, are designed to be transparent and encompasses research endeavors being explored by members of OHDSI comprehensive of relevant documentation, analysis code, results, and study protocols describing the study's scope and objectives.
@@
 The goal of OHDSI is to produce reproducible, robust, and reliable evidence by developing standardized data formats and research protocol designs to enable large-scale collaborative research and improve the transparency and scalability of observational health research.
-% TODO::MEDIUM Carry HADES discussion into the paragraph talking about research tooling
+% TODO::MEDIUM Carry HADES discussion into the paragraph talking about research tooling/***//***//***//***//***/
 One core innovation brought about by OHDSI in the space of network science is how, given the HADES tools and procedures established by OHDSI, 
-% TODO::EASY Make this more general such that it flows into the next paragraph on research tooling better
+% TODO::EASY Make this more general such that it flows into the next paragraph on research tooling better/***//***//***//***//***/
 a research group can develop their analysis on their own OMOP CDM instance and then bundle together their study.
-% EXPLAIN::EASY What is a study package?
+% EXPLAIN::EASY What is a study package?/***//***//***//***//***/
 Once this packaging is done, they can pass along the study package to another site to collaborate on the same analysis but execute the study package on their unique OMOP CDM formatted dataset.
-% EXPLAIN::MEDIUM What is a collaborator network?
+% EXPLAIN::MEDIUM What is a collaborator network?/***//***//***//***//***/
 Data confidentiality is preserved in this approach as data from individual participating sites remains exclusive to each site, with only aggregate results shared across a given collaborator network.
@@
 \cite{ohdsi2019book}
-
-% TODO::HARD Develop this paragraph more about research software
-As the OMOP CDM has become ubiquitous within observational health research, developing research software to handle and manage OMOP CDM data is necessary to meaningfully contribute within the observational health research space.
-OHDSI supports the development of open-source software tools in a centralized ecosystem called HADES (Health Analytics Data-to-Evidence Suite) for particularly managing and working with OMOP CDM formatted data.
-The HADES ecosystem has been in existence for $\approx 10$ years and has given rise to improvements in the quality and reliability of observational health research as well as motivating novel collaboration mechanisms.
 
@@
 It leverages the strengths of the Julia language to introduce new methods into these spaces, augment existing workflows, and provide a community structure for health researchers within the global Julia ecosystem.
-% EXPLAIN::EASY What is a subecosystem?
-% EXPLAIN:: How does having more subecosystems enable novel research to be done?
+% EXPLAIN::EASY What is a subecosystem?/***//***//***//***//***/
+% EXPLAIN:: How does having more subecosystems enable novel research to be done?/***//***//***//***//***/
 Eventually, JuliaHealth will host a variety of subecosystems and associated resources that will enable novel research to be conducted within Julia.
@@
 It can be said that OHS had its inception in work done by Evans and Simonov with efforts such as \textit{FunOHDSI.jl} \cite{evansFunOHDSIJl2022}, \textit{OHDSICohortExpressions.jl} \cite{evansOHDSICohortExpressionsJl2023}, and \textit{FunSQL.jl} around $2021$ \cite{kirill_simonov_2023_7705325}
-% TODO::EASY Cite OHDSI Symposium poster on JuliaHealth
-% TODO::HARD Register OHDSI Symposium poster to Zenodo
-% TODO::EASY Cite JuliaCon 2023 presentation
-% TODO::HARD Register JuliaCon 2023 presentation to Zenodo
+% TODO::EASY Cite OHDSI Symposium poster on JuliaHealth/***//***//***//***//***/
+% TODO::HARD Register OHDSI Symposium poster to Zenodo/***//***//***//***//***//***/
+% TODO::EASY Cite JuliaCon 2023 presentation/***//***//***//***//***/
+% TODO::HARD Register JuliaCon 2023 presentation to Zenodo/***//***//***//***//***/
 Since then, Zelko et.
@@
 have spearheaded most recent developments within OHS via the creation of a variety of tools and demonstrated their effectiveness in using these tools for novel observational health research endeavors \cite{zelko2022pilot}
-% EXPLAIN::MEDIUM What does this mean?
+% EXPLAIN::MEDIUM What does this mean?/***//***//***//***//***/
 Through such efforts, OHS is becoming further capable of executing novel observational health research.
@@
 
-% TODO::MEDIUM Add a paragraph here about what is a phenotype definition
-% TODO::EASY Cite my paper here
+% TODO::MEDIUM Add a paragraph here about what is a phenotype definition/***//***//***//***//***/
+% TODO::EASY Cite my paper here/***//***//***//***//***/
 
-% TODO::EASY Segue this better from phenotype definitions
+% TODO::EASY Segue this better from phenotype definitions/***//***//***//***//***/
 One of the premier tools within the OHDSI HADES ecosystem is the ATLAS Shiny application which was developed to develop patient phenotype definitions within observational health research contexts (amongst other use cases).
@@
 
-% TODO::MEDIUM Reframe this to more say how ATLAS can be used to build cohorts
-% TODO::EASY Explain how this package accomplishes similar tasks
+% TODO::MEDIUM Reframe this to more say how ATLAS can be used to build cohorts/***//***//***//***//***/
+% TODO::EASY Explain how this package accomplishes similar tasks/***//***//***//***//***/
 The Julia package, OHDSICohortExpressions.jl, exists specifically to ingest these serialized expressions of computable phenotype definitions and create patient cohorts that match the given criterion in the definition.
@@
 
-% EXPLAIN::MEDIUM What is a prepared statement?
-% EXPLAIN::MEDIUM What is a cohort in this context?
+% EXPLAIN::MEDIUM What is a prepared statement?/***//***//***//***//***/
+% EXPLAIN::MEDIUM What is a cohort in this context?/***//***//***//***//***/
 Once these prepared statements are made, researchers can execute the generated SQL against their OMOP CDM instance to generate cohorts for future analyses.
@@
 
-% TODO::EASY Reword this sentence
+% TODO::EASY Reword this sentence/***//***//***//***//***/
 Although OHDSICohortExpressions.jl is sufficient to translate existing phenotype definitions generated in ATLAS to executable queries, oftentimes, there is a tricky problem of how to best iterate on phenotype definitions.
 \cite{zelkoDevelopingRobustComputable2023}
-% EXPLAIN:: Better contextualize why this is important -- why not just create a new phenotype definition?
+% EXPLAIN:: Better contextualize why this is important -- why not just create a new phenotype definition?/***//***//***//***//***/
 Often, one has to take an existing definition and either revise to a more narrow cohort or write manual code to build more niche cohorts after an initial definition is created.
-% TODO::EASY Move this to the previous section on phenotype definitions
+% TODO::EASY Move this to the previous section on phenotype definitions/***//***//***//***//***/
 This broadly can be the notions of what we are exploring (the phenotype definition; often the responsibility of a clinical reviewer), what is possible to compute upon (the computable phenotype definition; often the responsibility of an analyst), and the goals of the overall study.
 
-% TODO::EASY Present this more as a general introduction to what tools are here for this
-% TODO::EASY Talk about composition with DataFrames or Tidier
+% TODO::EASY Present this more as a general introduction to what tools are here for this/***//***//***//***//***/
+% TODO::EASY Talk about composition with DataFrames or Tidier/***//***//***//***//***/
 OMOPCDMCohortCreator.jl allows one to iteratively build a patient population quickly and easily.
 One can use patient pools developed from pre-existing cohorts defined off of phenotype definitions as a starting point in exploring patients or create new patient populations quickly without the need for a well-defined phenotype definition.
-% TODO::MEDIUM Mention the ability to satisfy research requests
+% TODO::MEDIUM Mention the ability to satisfy research requests/***//***//***//***//***/
 This can enable analysts to more quickly understand what is inside their data assets, do "quick" analysis of a particular smaller research question, and enable a more distinct separation of work between a clinical expert reviewer and analyst teams.
@@
 Although the OMOP CDM has been great for standardizing \textit{how} real world data is handled, there is a lack of consensus on \textit{what} is needed to handle these data assets.
-% TODO::EASY Make this less verbose
+% TODO::EASY Make this less verbose/***//***//***//***//***/
 For a variety of reasons, different groups will choose different database architectures (perhaps PostgreSQL for improved permission support, DuckDB for performance, or even SQLite for ease of use) but all are subtly different enough to create headaches in seamlessly working across different research sites.
 
-% TODO::EASY Change the phrasing of this
+% TODO::EASY Change the phrasing of this/***//***//***//***//***/
 The way that the Julia ecosystem handles this is via the creation of a common database interface called the DBInterface.
-% EXPLAIN:: What is dispatch
+% EXPLAIN:: What is dispatch/***//***//***//***//***/
 Specific Julia database packages can implement dispatches for this interace allowing users to have a much better experience working across databases.
@@
 
-% TODO::MEDIUM Better introduce this
+% TODO::MEDIUM Better introduce this/***//***//***//***//***/
 For that reason, DBConnector.jl was created.
-% EXPLAIN::MEDIUM Contextualize this a bit better
+% EXPLAIN::MEDIUM Contextualize this a bit better/***//***//***//***//***/
 Although this is not an explicit JuliaHealth package, it was a pain point that was experienced throughout the process of observational health research across multiple partner sites.
@@
 
-% TODO::MEDIUM Combine with the previous paragraph and state that there are two approaches
-% TODO::EASY Make explicit DBConnector's path
-% TODO::EASY Make explicit the path of choosing general packages
+% TODO::MEDIUM Combine with the previous paragraph and state that there are two approaches/***//***//***//***//***/
+% TODO::EASY Make explicit DBConnector's path/***//***//***//***//***/
+% TODO::EASY Make explicit the path of choosing general packages/***//***//***//***//***/
 Finally, if someone needs more fine-grained control over a particular database connection, they can choose to eschew this package.
@@
 
-% TODO::HARD Review if this section is actually needed
+% TODO::HARD Review if this section is actually needed/***//***//***//***//***/
 \subsection{Ecosystem Interoperability}
@@
 
-% EXPLAIN::EASY What is this doing here?
-% TODO::HARD Update DBConnector
+% EXPLAIN::EASY What is this doing here?/***//***//***//***//***/
+% TODO::HARD Update DBConnector/***//***//***//***//***/
 \begin{listing}[!ht]
@@
 
-% EXPLAIN:: What is this doing here?
+% EXPLAIN:: What is this doing here?/***//***//***//***//***/
 After having a strep throat phenotype definition defined using ATLAS, we can then execute the resulting computable phenotype definition against the \textit{Eunomia} database.
@@
 
-% EXPLAIN:: What is this section doing?
+% EXPLAIN:: What is this section doing?/***//***//***//***//***/
 
@@
 
-% EXPLAIN::MEDIUM What is this section doing?
+% EXPLAIN::MEDIUM What is this section doing?/***//***//***//***//***/
 
@@
 
-% TODO::HARD Create table for MIMIC III results
+% TODO::HARD Create table for MIMIC III results/***//***//***//***//***/
 As seen in \ref{table:eunomia_prevalence}, informative results can be rapidly generated by this workflow relevant to observational health research.
@@
 
-% EXPLAIN::MEDIUM What is the overview of this section?
-% EXPLAIN::MEDIUM What does each topic do?
+% EXPLAIN::MEDIUM What is the overview of this section?/***//***//***//***//***/
+% EXPLAIN::MEDIUM What does each topic do?/***//***//***//***//***/
 \begin{enumerate}
@@
 
-% TODO::MEDIUM Make this more explicit about utilizing Julia's composition
+% TODO::MEDIUM Make this more explicit about utilizing Julia's composition/***//***//***//***//***/
 An exciting emergent aspect of the JuliaHealth ecosystem that came about as a result of developing OMOPCDMCohortCreator.jl was the prospect of radical modularization and chaining of package functionalities together.
-% TODO::MEDIUM Tie into currying more directly
-% TODO::MEDIUM Check if I am mentioning currying correctly
+% TODO::MEDIUM Tie into currying more directly/***//***//***//***//***/
+% TODO::MEDIUM Check if I am mentioning currying correctly/***//***//***//***//***/
 We can leverage a unique property of Julia in that Julia supports partial evaluation of functions (also known as currying) to simplify repeated function calls to the same function over and over again.
-% TODO::MEDIUM Rewrite this part in light of the previous sections 
+% TODO::MEDIUM Rewrite this part in light of the previous sections /***//***//***//***//***/
 For example, taking our workflow example from the earlier section where we characterized patients by race, gender, and age group, we can do a few "tricks" to modularize these characterization functions as follows:
 
-% TODO::EASY Rework this section further
+% TODO::EASY Rework this section further/***//***//***//***//***/
 First, we'll take our original functions and fix a connection object to each of our cofactor functions.
@@
 
-% EXPLAIN::EASY Refactor this paragraph better or put most of it into the caption for the next code block
+% EXPLAIN::EASY Refactor this paragraph better or put most of it into the caption for the next code block/***//***//***//***//***/
 Second, as OMOPCDMCohortCreator.jl is built on top of DataFrames.jl and one has the option of creating a DataFrame with the result from every OCC function, we can use the package, Chain.jl.
@@
 In this way, future researchers using tools such as OMOPCDMCohortCreator.jl can create these more convenient abstractions.
-% TODO::EASY Move this into the future discussion section
+% TODO::EASY Move this into the future discussion section/***//***//***//***//***/
 In fact, it could be a point of exploration in the future to build on top of OMOPCDMCohortCreator.jl even more abstracted and reusable functionalities to simplify ease of use while also enabling more readily auditable functionality.
 
-% TODO::MEDIUM Move this into its own section
+% TODO::MEDIUM Move this into its own section/***//***//***//***//***/
 Finally, as these functions have now been fully modularized, these can now be readily distributed using Julia's distributed computing methods:
 
-% TODO::HARD Redo example; I do not feel it is good
+% TODO::HARD Redo example; I do not feel it is good/***//***//***//***//***/
 \begin{listing}[!ht]
@@
 
-% TODO::EASY Rephrase this sentence
+% TODO::EASY Rephrase this sentence/***//***//***//***//***/
 One crucial aspect of the OHS is that it is unavoidable to work with databases that contain patient health information.
-% TODO::EASY Rephrase this sentence better
+% TODO::EASY Rephrase this sentence better/***//***//***//***//***/
 For better or worse, the tool that is used to best interact with these databases are SQL dialects that loosely follow consistently the ANSI SQL guidance.\cite{kelechavaSQLStandardISO2018}
-% EXPLAIN::MEDIUM Highlight this more
-% EXPLAIN::MEDIUM In what scenarios might using Julia may not be possible
+% EXPLAIN::MEDIUM Highlight this more/***//***//***//***//***/
+% EXPLAIN::MEDIUM In what scenarios might using Julia may not be possible/***//***//***//***//***/
 To accomodate for this, several of the core tools developed in this sub-ecosystem generate SQL that can be used in the case where Julia may not be available in a given environment housing patient health data.
-% EXPLAIN::MEDIUM Introduce this better
+% EXPLAIN::MEDIUM Introduce this better/***//***//***//***//***/
 As an example, \textit{OMOPCDMCohortCreator.jl} functions build upon \textit{FunSQL.jl} allowing one to use a function such `GetPatientAgeGroup` without a connection object which would result in SQL.
@@
 
-% TODO::MEDIUM Combine this section with the future directions section
+% TODO::MEDIUM Combine this section with the future directions section/***//***//***//***//***/
 \subsection{Strong Composition within JuliaHealth}
@@
 
-% TODO::EASY Rephrase this introductory paragraph
+% TODO::EASY Rephrase this introductory paragraph/***//***//***//***//***/
 JuliaHealth is now sufficiently at a stage of maturity where potential future researchers can leverage JuliaHealth's tools to conduct further future research or build upon existing architecture to target specific needs.
@@
 
-% TODO::MEDIUM This is where to introduce the paradox of composition
+% TODO::MEDIUM This is where to introduce the paradox of composition/***//***//***//***//***/
 Due to the strong composition that exists within the JuliaHealth and broader Julia ecosystem, solutions may already exist for solving various health informatics related research.
@@
 
-% TODO::MEDIUM Combine this paragraph with the previous paragraph
-% TODO::EASY Shorten this sentiment
+% TODO::MEDIUM Combine this paragraph with the previous paragraph/***//***//***//***//***/
+% TODO::EASY Shorten this sentiment/***//***//***//***//***/
 Additionally, due to the composable and modular nature of the OHS, research analyses and visualizations can be decoupled and interchanged with one another.
@@
 
-% TODO::EASY Make this more about the flexible nature of the JuliaHealth organization
+% TODO::EASY Make this more about the flexible nature of the JuliaHealth organization/***//***//***//***//***/
 Furthermore, again exploiting the composition within the JuliaHealth ecosystem, as mentioned in the context of JuliaHealth and machine learning applications, future efforts could focus on developing fairness and bias auditing tools in patient datasets.
@@
 Although it is an ecosystem that is relatively new, its robustness allows for novel investigations to be pursued within this domain of research. 
-% TODO::MEDIUM Mention how this allows JuliaHealth to solve more problems than what is present within JuliaHealth
+% TODO::MEDIUM Mention how this allows JuliaHealth to solve more problems than what is present within JuliaHealth/***//***//***//***//***/
 Additionally, due to the strongly compositional nature of the Julia ecosystem, these JuliaHealth tools can readily compose with tooling to support in-depth statistical analyses, mathematical modeling, and visualization capacities.
 
-% TODO::MEDIUM Rework this
+% TODO::MEDIUM Rework this/***//***//***//***//***/
 In essence, the JuliaHealth ecosystem stands as a catalyst for transformative observational health research, poised to push the boundaries of innovation, collaboration, and standardization in the pursuit of better healthcare outcomes.

### 2024-01-17 paper.tex
Wrapped up several TODOs

@@
 
-% TODO::HARD Review the abstract after reviewing the paper fully/***//***//***//***//***//***/
+% TODO::HARD Review the abstract after reviewing the paper fully/***//***//***/
 Observational health continues to be a growing field in health informatics research as electronic health records (EHR), patient medical claims, and other ancilliary patient data source become more readily computable and accessible to researchers.
@@
 
-% NOTE: I made this definition up on the fly using my expertise and familiarity with the space! :D /***//***//***//***//***//***/
+% NOTE: I made this definition up on the fly using my expertise and familiarity with the space! :D /***//***//***/
 Observational health research is the study and analysis of historical real world data to explore various cohorts found within such datasets to generate real world evidence to support applications within health-related domains.
@@
 The importance of this model has become apparent within open science organizations like the OHDSI (Observational Health Data Science and Informatics) network where members across $100+$ countries \cite{sachsonOurJourney2023} with access to real world data have been able to successfully transform their datasets to the OMOP CDM.
-This has led to successful collaboration relationships \cite{roseFindingAnswersBig2008} across countries and institutions where one institutional research group can develop an analysis centered around the OMOP CDM.
-Then, the network of institutions who have adopted the OMOP CDM can package and share this analysis amongst themselves (this is colloquially known as a "network study" within the context of observational health research) to enable research on an international scale.
-The network study idea is particularly salient to this field as it leads to the idea that you can "write [a study] once [and] run [it] everywhere."\footnotemark
+This has led to exciting new venues to enable successful collaboration relationships \cite{roseFindingAnswersBig2008} across countries and institutions.
 
-\footnotetext{This paraphrased quote is attributed to Sun Microsystems when discussing the Java programming language.}
-
-\subsection{Core Technologies}
-
-Modern observational health research utilizes several different technologies.
-
-\textit{Data standards within observational health research.} Although other standards exist within observational health, the OMOP CDM will be discussed and focused on for the remainder of this paper.
-To give a brief history of the OMOP CDM, OMOP was convened in 2009 by the US Food and Drug Administration to address concerns related to data type, study design, and data privacy concerns of utilizing "Real World Data" \cite{ohdsi2019book} \cite{FDARealWorldEvidence}
+\textit{Data standards within observational health research.}
+Although other standards exist within observational health, the OMOP CDM will be the particular standard focused on within this paper.
+To give further background on the OMOP CDM, OMOP was convened in 2009 by the US Food and Drug Administration to address concerns related to data type, study design, and data privacy concerns of utilizing "Real World Data".
+\cite{ohdsi2019book} \cite{FDARealWorldEvidence}
 Their collective efforts resulted in the OMOP Common Data Model (OMOP CDM) to house and ingest real world data that is used in observational health research.
@@
 
-% TODO::HARD Develop this paragraph more about research software/***//***//***//***//***/
-\textit{Research software to support the analysis of observational health data.} As the OMOP CDM has become ubiquitous within observational health research, developing research software to handle and manage OMOP CDM data is necessary to meaningfully contribute within the observational health research space.
-OHDSI supports the development of open-source software tools in a centralized ecosystem called HADES (Health Analytics Data-to-Evidence Suite) for particularly managing and working with OMOP CDM formatted data.
-The HADES ecosystem has been in existence for $\approx 10$ years and has given rise to improvements in the quality and reliability of observational health research as well as motivating novel collaboration mechanisms.
-
-% TODO::MEDIUM This paragraph should be centered on the core technology of the concept of network studies/***//***//***//***//***/
-% TODO::MEDIUM Use this to address the notion of what "network studies" are/***//***//***//***//***/
-% TODO::MEDIUM Make this initially less OHDSI specific; pull from https://jacobzelko.com/05122023034151-network-macarthur/***//***//***//***//***/
+% TODO::MEDIUM This paragraph should be centered on the core technology of the concept of network studies/***//***/
+% TODO::MEDIUM Use this to address the notion of what "network studies" are/***//***/
+% TODO::MEDIUM Make this initially less OHDSI specific; pull from https://jacobzelko.com/05122023034151-network-macarthur/***//***/
 \textit{Collaborative networks and studies using observational health data.} In particular, one collaboration mechanism that OHDSI has been able to operationalize meaningfully while working with observational health data is the notion of "network studies".
@@
 The goal of OHDSI is to produce reproducible, robust, and reliable evidence by developing standardized data formats and research protocol designs to enable large-scale collaborative research and improve the transparency and scalability of observational health research.
-% TODO::MEDIUM Carry HADES discussion into the paragraph talking about research tooling/***//***//***//***//***/
+% TODO::MEDIUM Carry HADES discussion into the paragraph talking about research tooling/***//***/
 One core innovation brought about by OHDSI in the space of network science is how, given the HADES tools and procedures established by OHDSI, 
-% TODO::EASY Make this more general such that it flows into the next paragraph on research tooling better/***//***//***//***//***/
+% TODO::EASY Make this more general such that it flows into the next paragraph on research tooling better/***//***/
 a research group can develop their analysis on their own OMOP CDM instance and then bundle together their study.
-% EXPLAIN::EASY What is a study package?/***//***//***//***//***/
+% EXPLAIN::EASY What is a study package?/***//***/
 Once this packaging is done, they can pass along the study package to another site to collaborate on the same analysis but execute the study package on their unique OMOP CDM formatted dataset.
-% EXPLAIN::MEDIUM What is a collaborator network?/***//***//***//***//***/
+% EXPLAIN::MEDIUM What is a collaborator network?/***//***/
 Data confidentiality is preserved in this approach as data from individual participating sites remains exclusive to each site, with only aggregate results shared across a given collaborator network.
@@
 \cite{ohdsi2019book}
+
+% TODO::HARD Develop this paragraph more about research software/***//***/
+\textit{Research software to support the analysis of observational health data.} As the OMOP CDM has become ubiquitous within observational health research, developing research software to handle and manage OMOP CDM data is necessary to meaningfully contribute within the observational health research space.
+OHDSI supports the development of open-source software tools in a centralized ecosystem called HADES (Health Analytics Data-to-Evidence Suite) for particularly managing and working with OMOP CDM formatted data.
+The HADES ecosystem has been in existence for $\approx 10$ years and has given rise to improvements in the quality and reliability of observational health research as well as motivating novel collaboration mechanisms.
 
@@
 It leverages the strengths of the Julia language to introduce new methods into these spaces, augment existing workflows, and provide a community structure for health researchers within the global Julia ecosystem.
-% EXPLAIN::EASY What is a subecosystem?/***//***//***//***//***/
-% EXPLAIN:: How does having more subecosystems enable novel research to be done?/***//***//***//***//***/
+% EXPLAIN::EASY What is a subecosystem?/***//***/
+% EXPLAIN:: How does having more subecosystems enable novel research to be done?/***//***/
 Eventually, JuliaHealth will host a variety of subecosystems and associated resources that will enable novel research to be conducted within Julia.
@@
 It can be said that OHS had its inception in work done by Evans and Simonov with efforts such as \textit{FunOHDSI.jl} \cite{evansFunOHDSIJl2022}, \textit{OHDSICohortExpressions.jl} \cite{evansOHDSICohortExpressionsJl2023}, and \textit{FunSQL.jl} around $2021$ \cite{kirill_simonov_2023_7705325}
-% TODO::EASY Cite OHDSI Symposium poster on JuliaHealth/***//***//***//***//***/
-% TODO::HARD Register OHDSI Symposium poster to Zenodo/***//***//***//***//***//***/
-% TODO::EASY Cite JuliaCon 2023 presentation/***//***//***//***//***/
-% TODO::HARD Register JuliaCon 2023 presentation to Zenodo/***//***//***//***//***/
+% TODO::EASY Cite OHDSI Symposium poster on JuliaHealth/***//***/
+% TODO::HARD Register OHDSI Symposium poster to Zenodo/***//***//***/
+% TODO::EASY Cite JuliaCon 2023 presentation/***//***/
+% TODO::HARD Register JuliaCon 2023 presentation to Zenodo/***//***/
 Since then, Zelko et.
@@
 have spearheaded most recent developments within OHS via the creation of a variety of tools and demonstrated their effectiveness in using these tools for novel observational health research endeavors \cite{zelko2022pilot}
-% EXPLAIN::MEDIUM What does this mean?/***//***//***//***//***/
+% EXPLAIN::MEDIUM What does this mean?/***//***/
 Through such efforts, OHS is becoming further capable of executing novel observational health research.
@@
 
-% TODO::MEDIUM Add a paragraph here about what is a phenotype definition/***//***//***//***//***/
-% TODO::EASY Cite my paper here/***//***//***//***//***/
+% TODO::MEDIUM Add a paragraph here about what is a phenotype definition/***//***/
+% TODO::EASY Cite my paper here/***//***/
 
-% TODO::EASY Segue this better from phenotype definitions/***//***//***//***//***/
+% TODO::EASY Segue this better from phenotype definitions/***//***/
 One of the premier tools within the OHDSI HADES ecosystem is the ATLAS Shiny application which was developed to develop patient phenotype definitions within observational health research contexts (amongst other use cases).
@@
 
-% TODO::MEDIUM Reframe this to more say how ATLAS can be used to build cohorts/***//***//***//***//***/
-% TODO::EASY Explain how this package accomplishes similar tasks/***//***//***//***//***/
+% TODO::MEDIUM Reframe this to more say how ATLAS can be used to build cohorts/***//***/
+% TODO::EASY Explain how this package accomplishes similar tasks/***//***/
 The Julia package, OHDSICohortExpressions.jl, exists specifically to ingest these serialized expressions of computable phenotype definitions and create patient cohorts that match the given criterion in the definition.
@@
 
-% EXPLAIN::MEDIUM What is a prepared statement?/***//***//***//***//***/
-% EXPLAIN::MEDIUM What is a cohort in this context?/***//***//***//***//***/
+% EXPLAIN::MEDIUM What is a prepared statement?/***//***/
+% EXPLAIN::MEDIUM What is a cohort in this context?/***//***/
 Once these prepared statements are made, researchers can execute the generated SQL against their OMOP CDM instance to generate cohorts for future analyses.
@@
 
-% TODO::EASY Reword this sentence/***//***//***//***//***/
+% TODO::EASY Reword this sentence/***//***/
 Although OHDSICohortExpressions.jl is sufficient to translate existing phenotype definitions generated in ATLAS to executable queries, oftentimes, there is a tricky problem of how to best iterate on phenotype definitions.
 \cite{zelkoDevelopingRobustComputable2023}
-% EXPLAIN:: Better contextualize why this is important -- why not just create a new phenotype definition?/***//***//***//***//***/
+% EXPLAIN:: Better contextualize why this is important -- why not just create a new phenotype definition?/***//***/
 Often, one has to take an existing definition and either revise to a more narrow cohort or write manual code to build more niche cohorts after an initial definition is created.
-% TODO::EASY Move this to the previous section on phenotype definitions/***//***//***//***//***/
+% TODO::EASY Move this to the previous section on phenotype definitions/***//***/
 This broadly can be the notions of what we are exploring (the phenotype definition; often the responsibility of a clinical reviewer), what is possible to compute upon (the computable phenotype definition; often the responsibility of an analyst), and the goals of the overall study.
 
-% TODO::EASY Present this more as a general introduction to what tools are here for this/***//***//***//***//***/
-% TODO::EASY Talk about composition with DataFrames or Tidier/***//***//***//***//***/
+% TODO::EASY Present this more as a general introduction to what tools are here for this/***//***/
+% TODO::EASY Talk about composition with DataFrames or Tidier/***//***/
 OMOPCDMCohortCreator.jl allows one to iteratively build a patient population quickly and easily.
 One can use patient pools developed from pre-existing cohorts defined off of phenotype definitions as a starting point in exploring patients or create new patient populations quickly without the need for a well-defined phenotype definition.
-% TODO::MEDIUM Mention the ability to satisfy research requests/***//***//***//***//***/
+% TODO::MEDIUM Mention the ability to satisfy research requests/***//***/
 This can enable analysts to more quickly understand what is inside their data assets, do "quick" analysis of a particular smaller research question, and enable a more distinct separation of work between a clinical expert reviewer and analyst teams.
@@
 Although the OMOP CDM has been great for standardizing \textit{how} real world data is handled, there is a lack of consensus on \textit{what} is needed to handle these data assets.
-% TODO::EASY Make this less verbose/***//***//***//***//***/
+% TODO::EASY Make this less verbose/***//***/
 For a variety of reasons, different groups will choose different database architectures (perhaps PostgreSQL for improved permission support, DuckDB for performance, or even SQLite for ease of use) but all are subtly different enough to create headaches in seamlessly working across different research sites.
 
-% TODO::EASY Change the phrasing of this/***//***//***//***//***/
+% TODO::EASY Change the phrasing of this/***//***/
 The way that the Julia ecosystem handles this is via the creation of a common database interface called the DBInterface.
-% EXPLAIN:: What is dispatch/***//***//***//***//***/
+% EXPLAIN:: What is dispatch/***//***/
 Specific Julia database packages can implement dispatches for this interace allowing users to have a much better experience working across databases.
@@
 
-% TODO::MEDIUM Better introduce this/***//***//***//***//***/
+% TODO::MEDIUM Better introduce this/***//***/
 For that reason, DBConnector.jl was created.
-% EXPLAIN::MEDIUM Contextualize this a bit better/***//***//***//***//***/
+% EXPLAIN::MEDIUM Contextualize this a bit better/***//***/
 Although this is not an explicit JuliaHealth package, it was a pain point that was experienced throughout the process of observational health research across multiple partner sites.
@@
 
-% TODO::MEDIUM Combine with the previous paragraph and state that there are two approaches/***//***//***//***//***/
-% TODO::EASY Make explicit DBConnector's path/***//***//***//***//***/
-% TODO::EASY Make explicit the path of choosing general packages/***//***//***//***//***/
+% TODO::MEDIUM Combine with the previous paragraph and state that there are two approaches/***//***/
+% TODO::EASY Make explicit DBConnector's path/***//***/
+% TODO::EASY Make explicit the path of choosing general packages/***//***/
 Finally, if someone needs more fine-grained control over a particular database connection, they can choose to eschew this package.
@@
 
-% TODO::HARD Review if this section is actually needed/***//***//***//***//***/
+% TODO::HARD Review if this section is actually needed/***//***/
 \subsection{Ecosystem Interoperability}
@@
 
-% EXPLAIN::EASY What is this doing here?/***//***//***//***//***/
-% TODO::HARD Update DBConnector/***//***//***//***//***/
+% EXPLAIN::EASY What is this doing here?/***//***/
+% TODO::HARD Update DBConnector/***//***/
 \begin{listing}[!ht]
@@
 
-% EXPLAIN:: What is this doing here?/***//***//***//***//***/
+% EXPLAIN:: What is this doing here?/***//***/
 After having a strep throat phenotype definition defined using ATLAS, we can then execute the resulting computable phenotype definition against the \textit{Eunomia} database.
@@
 
-% EXPLAIN:: What is this section doing?/***//***//***//***//***/
+% EXPLAIN:: What is this section doing?/***//***/
 
@@
 
-% EXPLAIN::MEDIUM What is this section doing?/***//***//***//***//***/
+% EXPLAIN::MEDIUM What is this section doing?/***//***/
 
@@
 
-% TODO::HARD Create table for MIMIC III results/***//***//***//***//***/
+% TODO::HARD Create table for MIMIC III results/***//***/
 As seen in \ref{table:eunomia_prevalence}, informative results can be rapidly generated by this workflow relevant to observational health research.
@@
 
-% EXPLAIN::MEDIUM What is the overview of this section?/***//***//***//***//***/
-% EXPLAIN::MEDIUM What does each topic do?/***//***//***//***//***/
+% EXPLAIN::MEDIUM What is the overview of this section?/***//***/
+% EXPLAIN::MEDIUM What does each topic do?/***//***/
 \begin{enumerate}
@@
 
-% TODO::MEDIUM Make this more explicit about utilizing Julia's composition/***//***//***//***//***/
+% TODO::MEDIUM Make this more explicit about utilizing Julia's composition/***//***/
 An exciting emergent aspect of the JuliaHealth ecosystem that came about as a result of developing OMOPCDMCohortCreator.jl was the prospect of radical modularization and chaining of package functionalities together.
-% TODO::MEDIUM Tie into currying more directly/***//***//***//***//***/
-% TODO::MEDIUM Check if I am mentioning currying correctly/***//***//***//***//***/
+% TODO::MEDIUM Tie into currying more directly/***//***/
+% TODO::MEDIUM Check if I am mentioning currying correctly/***//***/
 We can leverage a unique property of Julia in that Julia supports partial evaluation of functions (also known as currying) to simplify repeated function calls to the same function over and over again.
-% TODO::MEDIUM Rewrite this part in light of the previous sections /***//***//***//***//***/
+% TODO::MEDIUM Rewrite this part in light of the previous sections /***//***/
 For example, taking our workflow example from the earlier section where we characterized patients by race, gender, and age group, we can do a few "tricks" to modularize these characterization functions as follows:
 
-% TODO::EASY Rework this section further/***//***//***//***//***/
+% TODO::EASY Rework this section further/***//***/
 First, we'll take our original functions and fix a connection object to each of our cofactor functions.
@@
 
-% EXPLAIN::EASY Refactor this paragraph better or put most of it into the caption for the next code block/***//***//***//***//***/
+% EXPLAIN::EASY Refactor this paragraph better or put most of it into the caption for the next code block/***//***/
 Second, as OMOPCDMCohortCreator.jl is built on top of DataFrames.jl and one has the option of creating a DataFrame with the result from every OCC function, we can use the package, Chain.jl.
@@
 In this way, future researchers using tools such as OMOPCDMCohortCreator.jl can create these more convenient abstractions.
-% TODO::EASY Move this into the future discussion section/***//***//***//***//***/
+% TODO::EASY Move this into the future discussion section/***//***/
 In fact, it could be a point of exploration in the future to build on top of OMOPCDMCohortCreator.jl even more abstracted and reusable functionalities to simplify ease of use while also enabling more readily auditable functionality.
 
-% TODO::MEDIUM Move this into its own section/***//***//***//***//***/
+% TODO::MEDIUM Move this into its own section/***//***/
 Finally, as these functions have now been fully modularized, these can now be readily distributed using Julia's distributed computing methods:
 
-% TODO::HARD Redo example; I do not feel it is good/***//***//***//***//***/
+% TODO::HARD Redo example; I do not feel it is good/***//***/
 \begin{listing}[!ht]
@@
 
-% TODO::EASY Rephrase this sentence/***//***//***//***//***/
+% TODO::EASY Rephrase this sentence/***//***/
 One crucial aspect of the OHS is that it is unavoidable to work with databases that contain patient health information.
-% TODO::EASY Rephrase this sentence better/***//***//***//***//***/
+% TODO::EASY Rephrase this sentence better/***//***/
 For better or worse, the tool that is used to best interact with these databases are SQL dialects that loosely follow consistently the ANSI SQL guidance.\cite{kelechavaSQLStandardISO2018}
-% EXPLAIN::MEDIUM Highlight this more/***//***//***//***//***/
-% EXPLAIN::MEDIUM In what scenarios might using Julia may not be possible/***//***//***//***//***/
+% EXPLAIN::MEDIUM Highlight this more/***//***/
+% EXPLAIN::MEDIUM In what scenarios might using Julia may not be possible/***//***/
 To accomodate for this, several of the core tools developed in this sub-ecosystem generate SQL that can be used in the case where Julia may not be available in a given environment housing patient health data.
-% EXPLAIN::MEDIUM Introduce this better/***//***//***//***//***/
+% EXPLAIN::MEDIUM Introduce this better/***//***/
 As an example, \textit{OMOPCDMCohortCreator.jl} functions build upon \textit{FunSQL.jl} allowing one to use a function such `GetPatientAgeGroup` without a connection object which would result in SQL.
@@
 
-% TODO::MEDIUM Combine this section with the future directions section/***//***//***//***//***/
+% TODO::MEDIUM Combine this section with the future directions section/***//***/
 \subsection{Strong Composition within JuliaHealth}
@@
 
-% TODO::EASY Rephrase this introductory paragraph/***//***//***//***//***/
+% TODO::EASY Rephrase this introductory paragraph/***//***/
 JuliaHealth is now sufficiently at a stage of maturity where potential future researchers can leverage JuliaHealth's tools to conduct further future research or build upon existing architecture to target specific needs.
@@
 
-% TODO::MEDIUM This is where to introduce the paradox of composition/***//***//***//***//***/
+% TODO::MEDIUM This is where to introduce the paradox of composition/***//***/
 Due to the strong composition that exists within the JuliaHealth and broader Julia ecosystem, solutions may already exist for solving various health informatics related research.
@@
 
-% TODO::MEDIUM Combine this paragraph with the previous paragraph/***//***//***//***//***/
-% TODO::EASY Shorten this sentiment/***//***//***//***//***/
+% TODO::MEDIUM Combine this paragraph with the previous paragraph/***//***/
+% TODO::EASY Shorten this sentiment/***//***/
 Additionally, due to the composable and modular nature of the OHS, research analyses and visualizations can be decoupled and interchanged with one another.
@@
 
-% TODO::EASY Make this more about the flexible nature of the JuliaHealth organization/***//***//***//***//***/
+% TODO::EASY Make this more about the flexible nature of the JuliaHealth organization/***//***/
 Furthermore, again exploiting the composition within the JuliaHealth ecosystem, as mentioned in the context of JuliaHealth and machine learning applications, future efforts could focus on developing fairness and bias auditing tools in patient datasets.
@@
 Although it is an ecosystem that is relatively new, its robustness allows for novel investigations to be pursued within this domain of research. 
-% TODO::MEDIUM Mention how this allows JuliaHealth to solve more problems than what is present within JuliaHealth/***//***//***//***//***/
+% TODO::MEDIUM Mention how this allows JuliaHealth to solve more problems than what is present within JuliaHealth/***//***/
 Additionally, due to the strongly compositional nature of the Julia ecosystem, these JuliaHealth tools can readily compose with tooling to support in-depth statistical analyses, mathematical modeling, and visualization capacities.
 
-% TODO::MEDIUM Rework this/***//***//***//***//***/
+% TODO::MEDIUM Rework this/***//***/
 In essence, the JuliaHealth ecosystem stands as a catalyst for transformative observational health research, poised to push the boundaries of innovation, collaboration, and standardization in the pursuit of better healthcare outcomes.

### 2024-01-17 paper.tex
Added Ergina reference to protocols in intro

@@
 Research within this space can investigate questions such as population health trends for a particular disease, health disparities within patient populations, and cost costs of care related to particular treatment procedures.
-Due to the untapped potential of this space, over the past $20$ years, there have been several developments within the observational health research community to enable more effective research to be conducted within this field (such as the creation of interoperability standards and tool creation \cite{benderHL7FHIRAgile2013; mandelSMARTFHIRStandardsbased2016; overhage2012validation}).
+Due to the untapped potential of this space, over the past $20$ years, there have been several developments within the observational health research community to enable more effective research to be conducted within this field (such as the creation of interoperability standards, protocols, and tool creation \cite{benderHL7FHIRAgile2013; mandelSMARTFHIRStandardsbased2016; overhage2012validation; ergina2013ideal}).
 
